---
title: "SWC Habitat Model Data Discovery"
author: "[Skyler Lewis](mailto:slewis@flowwest.com)"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
    toc_depth: 3
    number_sections: false
    math_method:
      engine: webtex
      url: https://latex.codecogs.com/svg.image?
  html_document:
    toc: true
    toc_depth: 3
    number_sections: false
    math_method:
      engine: webtex
      url: https://latex.codecogs.com/svg.image?
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(sf)
project_crs <- "EPSG:3310" #"EPSG:5070"
```

*Selected HUCs*

* Lower Yuba 18020107
* Upper Yuba 18020125

*Data Source*

* NHDPlusV2 dataset https://nhdplus.com/NHDPlus/NHDPlusV2_home.php
* As retrievedfrom https://www.epa.gov/waterdata/nhdplus-california-data-vector-processing-unit-18
* User Guide https://www.epa.gov/system/files/documents/2023-04/NHDPlusV2_User_Guide.pdf
* (This is different from NHDPlusHR which is generated at a higher resolution and more computationally intensive)

## Data Import

## Import HUC Watersheds

```{r}
selected_huc_8 <- c("18020107", "18020125")

# HUC-12 watersheds and higher level heirarchies
watersheds <- st_read("nhdplus/WBD_Subwatershed.shp") |> 
  janitor::clean_names() |>
  filter(huc_8 %in% selected_huc_8) |>
  st_transform(project_crs)

watersheds |> ggplot() + geom_sf()
```

## Import Supplemental Tables

```{r}
# identifiers of each flowline (COMID)
flowline_comid_reachcode_crosswalk <- 
  foreign::read.dbf("nhdplus/NHDReachCode_Comid.dbf") |> 
  as_tibble() |> 
  janitor::clean_names() |>
  mutate(huc_8 = substr(reachcode, 1, 8),
         huc_10 = substr(reachcode, 1, 10),
         huc_12 = substr(reachcode, 1, 12)) |>
  filter(huc_8 %in% selected_huc_8)

# slopes and endpoint elevations
flowline_slopes <- 
  foreign::read.dbf("nhdplus/NHDPlusAttributes/elevslope.dbf") |> 
  as_tibble() |> 
  janitor::clean_names() |>
  mutate(slope = if_else(slope==-9998, NA, slope))

# cumulative upstream area
flowline_cumarea <- 
  foreign::read.dbf("nhdplus/NHDPlusAttributes/CumulativeArea.dbf") |> 
  as_tibble() |> 
  janitor::clean_names() |>
  rename(comid = com_id)

# flow routing attributes as described in https://www.usgs.gov/national-hydrography/value-added-attributes-vaas
flowline_vaattr <- 
  foreign::read.dbf("nhdplus/NHDPlusAttributes/PlusFlowlineVAA.dbf") |> 
  as_tibble() |> 
  janitor::clean_names()

# fcode table
fcodes <- 
  foreign::read.dbf("nhdplus/NHDFCode.dbf") |> 
  as_tibble() |> 
  janitor::clean_names() |>
  rename(fcode = f_code, fcode_desc = descriptio)
```

### Import flowline geometry

```{r}
flowlines <- 
  st_read("nhdplus/Hydrography/NHDFlowline.shp") |>
  janitor::clean_names() |>
  mutate(huc_8 = substr(reachcode, 1, 8),
         huc_10 = substr(reachcode, 1, 10),
         huc_12 = substr(reachcode, 1, 12)) |>
  filter(huc_8 %in% selected_huc_8) |>
  inner_join(fcodes |> select(fcode, fcode_desc)) |>
  inner_join(flowline_slopes |> select(comid, slope, maxelevsmo, minelevsmo, slopelenkm), by = join_by(comid)) |>
  inner_join(flowline_cumarea, by = join_by(comid)) |>
  arrange(comid) |>
  st_transform(project_crs)

# plot showing slopes
flowlines |> 
  st_zm() |>
  filter(gnis_name %in% c("Yuba River", "South Yuba River", "Middle Yuba River", "North Yuba River")) |>
  ggplot() + 
  geom_sf(data=st_zm(flowlines), aes(color = slope)) +
  geom_sf(aes(color = slope), linewidth=1) + 
  scale_color_viridis_c(trans = "log")

```

### Import catchments

```{r}
# catchment associated with each flowline reach (COMID)
catchments <- 
  st_read("nhdplus/Catchment.shp") |> 
  janitor::clean_names() |>
  mutate(comid = as.numeric(featureid)) |>
  inner_join(flowlines |> st_drop_geometry() |> select(comid)) |>
  arrange(comid) |>
  st_transform(project_crs) 

# examples
ggplot() + geom_sf(data = catchments, color="orange") + 
  geom_sf(data = watersheds, color="red", fill=NA) + 
  geom_sf(data = st_zm(flowlines), color="blue") 
```

## Stream Widths?

#### Points along line

using st_line_sample or st_segmentize

```{r}
# approach with one point every 100 ft
sample_points_100ft <- 
  flowlines |> 
  filter(gnis_name %in% c("Yuba River", "South Yuba River", "Middle Yuba River", "North Yuba River")) |>
  st_line_sample(density = 1 / 100) 

# approach with just the midpoint of each segment
sample_points_midpt <- 
  flowlines |>
  filter(gnis_name %in% c("Yuba River", "South Yuba River", "Middle Yuba River", "North Yuba River")) |>
  #st_line_sample(sample = c(0, 0.25, 0.5, 0.75, 1))
  st_line_sample(sample = c(0.5))

sample_points_midpt |> ggplot() + geom_sf()
```
