---
title: "SWC Habitat Model Data Discovery"
author: "[Skyler Lewis](mailto:slewis@flowwest.com)"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
    toc_depth: 3
    number_sections: false
    math_method:
      engine: webtex
      url: https://latex.codecogs.com/svg.image?
  html_document:
    toc: true
    toc_depth: 3
    number_sections: false
    math_method:
      engine: webtex
      url: https://latex.codecogs.com/svg.image?
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(sf)
library(stars)

theme_set(theme_minimal())

#project_crs <- "EPSG:3310" # NAD83 California Teale Albers
#project_crs <- "EPSG:5070" # NAD83 CONUS Albers
project_crs <- "ESRI:102039" # NAD83 CONUS Albers USGS Version
```

*Selected HUCs*

* Lower Yuba 18020107
* Upper Yuba 18020125

*Data Source*

* NHDPlusV2 dataset https://nhdplus.com/NHDPlus/NHDPlusV2_home.php
* As retrievedfrom https://www.epa.gov/waterdata/nhdplus-california-data-vector-processing-unit-18
* User Guide https://www.epa.gov/system/files/documents/2023-04/NHDPlusV2_User_Guide.pdf
* (This is different from [NHDPlusHR](https://pubs.usgs.gov/of/2019/1096/ofr20191096.pdf) which is generated at a higher resolution and more computationally intensive)

## Data Import

## Import HUC Watersheds

```{r}
selected_huc_8 <- c("18020107", "18020125")

# HUC-12 watersheds and higher level heirarchies
watersheds <- st_read("nhdplus/WBD_Subwatershed.shp") |> 
  janitor::clean_names() |>
  filter(huc_8 %in% selected_huc_8) |>
  st_transform(project_crs)

watersheds |> ggplot() + geom_sf()
```

## Import Supplemental Tables

```{r}
# identifiers of each flowline (COMID)
flowline_comid_reachcode_crosswalk <- 
  foreign::read.dbf("nhdplus/NHDReachCode_Comid.dbf") |> 
  as_tibble() |> 
  janitor::clean_names() |>
  mutate(huc_8 = substr(reachcode, 1, 8),
         huc_10 = substr(reachcode, 1, 10),
         huc_12 = substr(reachcode, 1, 12)) |>
  filter(huc_8 %in% selected_huc_8)

# slopes and endpoint elevations
flowline_slopes <- 
  foreign::read.dbf("nhdplus/NHDPlusAttributes/elevslope.dbf") |> 
  as_tibble() |> 
  janitor::clean_names() |>
  mutate(slope = if_else(slope==-9998, NA, slope))

# cumulative upstream area
flowline_cumarea <- 
  foreign::read.dbf("nhdplus/NHDPlusAttributes/CumulativeArea.dbf") |> 
  as_tibble() |> 
  janitor::clean_names() |>
  rename(comid = com_id)

# flow routing attributes as described in https://www.usgs.gov/national-hydrography/value-added-attributes-vaas
flowline_vaattr <- 
  foreign::read.dbf("nhdplus/NHDPlusAttributes/PlusFlowlineVAA.dbf") |> 
  as_tibble() |> 
  janitor::clean_names()

# fcode table
fcodes <- 
  foreign::read.dbf("nhdplus/NHDFCode.dbf") |> 
  as_tibble() |> 
  janitor::clean_names() |>
  rename(fcode = f_code, fcode_desc = descriptio)
```

### Import flowline geometry

```{r}
flowlines <- 
  st_read("nhdplus/Hydrography/NHDFlowline.shp") |>
  janitor::clean_names() |>
  mutate(huc_8 = substr(reachcode, 1, 8),
         huc_10 = substr(reachcode, 1, 10),
         huc_12 = substr(reachcode, 1, 12)) |>
  filter(huc_8 %in% selected_huc_8) |>
  inner_join(fcodes |> select(fcode, fcode_desc)) |>
  inner_join(flowline_slopes |> select(comid, slope, maxelevsmo, minelevsmo, slopelenkm), by = join_by(comid)) |>
  inner_join(flowline_cumarea, by = join_by(comid)) |>
  arrange(comid) |>
  st_transform(project_crs)

#flowareas <- 
#  st_read("nhdplus/Hydrography/NHDArea.shp") |>
#  janitor::clean_names() |>
#  arrange(comid) #|>
#  #inner_join(flowlines |> st_drop_geometry() |> select(comid)) |>
#  #select(comid, ftype, fcode, geometry)

waterbodies <- 
  st_read("nhdplus/Hydrography/NHDWaterbody.shp") |>
  janitor::clean_names() |>
  mutate(huc_8 = substr(reachcode, 1, 8),
         huc_10 = substr(reachcode, 1, 10),
         huc_12 = substr(reachcode, 1, 12)) |>
  filter(huc_8 %in% selected_huc_8) |>
  arrange(comid) |>
  st_transform(project_crs)

# plot showing slopes
flowlines |> 
  st_zm() |>
  filter(gnis_name %in% c("Yuba River", "South Yuba River", "Middle Yuba River", "North Yuba River")) |>
  ggplot() + 
  geom_sf(data=st_zm(flowlines), aes(color = slope)) +
  geom_sf(aes(color = slope), linewidth=1) + 
  geom_sf(data=waterbodies, fill="black", color="black") +
  scale_color_viridis_c(trans = "log")

```

```{r}
# plot showing drainage area
flowlines |> 
  st_zm() |>
  filter(gnis_name %in% c("Yuba River", "South Yuba River", "Middle Yuba River", "North Yuba River")) |>
  ggplot() + 
  geom_sf(data=st_zm(flowlines), aes(color = tot_da_sq_km)) +
  geom_sf(aes(color = tot_da_sq_km), linewidth=1) + 
  geom_sf(data=waterbodies, fill="black", color="black") +
  scale_color_viridis_c(direction=-1)
```

### Import catchments

```{r}
# catchment associated with each flowline reach (COMID)
catchments <- 
  st_read("nhdplus/Catchment.shp") |> 
  janitor::clean_names() |>
  mutate(comid = as.numeric(featureid)) |>
  inner_join(flowlines |> st_drop_geometry() |> select(comid)) |>
  arrange(comid) |>
  st_transform(project_crs) 

# examples
ggplot() + geom_sf(data = catchments, color="orange") + 
  geom_sf(data = watersheds, color="red", fill=NA) + 
  geom_sf(data = st_zm(flowlines), color="blue") 
```

## Stream Widths?

#### Points along line

using st_line_sample or st_segmentize

```{r}
# approach with one point every 100 ft
sample_points_100ft <- 
  flowlines |> 
  filter(gnis_name %in% c("Yuba River", "South Yuba River", "Middle Yuba River", "North Yuba River")) |>
  st_line_sample(density = 1 / 100) 

# approach with just the midpoint of each segment
sample_points_midpt <- 
  flowlines |>
  filter(gnis_name %in% c("Yuba River", "South Yuba River", "Middle Yuba River", "North Yuba River")) |>
  #st_line_sample(sample = c(0, 0.25, 0.5, 0.75, 1))
  st_line_sample(sample = c(0.5))

sample_points_midpt |> ggplot() + geom_sf()
```

```{r}
cvpia_extents <- st_read("habitat_extents_cvpia/habitat_extents_combined_gradients_v3.shp")

cvpia_extents |> 
  st_cast("LINESTRING") |>
  mutate(geometry = st_line_sample(geometry, sample = c(0,1))) |>
  ggplot() + geom_sf()
```

## DEM 

### 10m NHDPlusHR

```{r}
dem <- read_stars("nhdplushr/dem_nhdplushr_yuba_meters_v2.tif")
# this input has already been clipped to the AOI, converted from cm to m, and crs redefined
ggplot() + geom_stars(data=dem) + coord_fixed()
```

#### export flowlines within this raster domain

```{r}
flowlines_for_vbet <- flowlines |>
  filter(gnis_name %in% c("Yuba River")) |>
  st_transform(st_crs(dem)) |>
  st_zm() |> 
  st_crop(st_bbox(dem)) |>
  smoothr::densify(5)

flowlines_for_vbet |>
  st_write("out/flowlines_for_vbet.shp", append=FALSE)

flowlines_for_vbet |> ggplot() + geom_sf(aes(color = tot_da_sq_km)) #+ geom_stars(data=dem)
```

### 1m CA NoCAL Wildfires B2 2018

Not generalizable/automatable

https://rockyweb.usgs.gov/vdelivery/Datasets/Staged/Elevation/LPC/Projects/CA_NoCAL_3DEP_Supp_Funding_2018_D18/CA_NoCAL_Wildfires_B2_2018/

https://prd-tnm.s3.amazonaws.com/index.html?prefix=StagedProducts/Elevation/OPR/Projects/CA_NoCAL_Wildfires_B2_2018

https://prd-tnm.s3.amazonaws.com/index.html?prefix=StagedProducts/Elevation/metadata/CA_NoCAL_3DEP_Supp_Funding_2018_D18/CA_NoCAL_Wildfires_B2_2018


