---
title: "Exploratory Modeling"
author: "[Skyler Lewis](mailto:slewis@flowwest.com)"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
    toc_depth: 3
    number_sections: false
    math_method:
      engine: webtex
      url: https://latex.codecogs.com/svg.image?
  html_document:
    toc: true
    toc_depth: 3
    number_sections: false
    math_method:
      engine: webtex
      url: https://latex.codecogs.com/svg.image?
---

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(sf)
library(stars)
library(tidymodels)
library(broom.mixed) # tidy output of mixed model results
library(dotwhisker) # visualize regression results

knitr::opts_chunk$set(eval=TRUE, fig.width=6.5, fig.height=4, dpi=300)

theme_set(theme_minimal())
```

```{r import}
flowlines <- readRDS("../data/flowline_geometries.Rds") |>
  st_transform("ESRI:102039")

flowline_attributes <- readRDS("../data/flowline_attributes.Rds")

flow_to_suitable_area <- readRDS("../data/fsa_combined.Rds")

# Optionally, swap out the original HSI ranges with the version using the HQT depth/velocity cutoffs
if(FALSE){
  flow_to_suitable_area <- 
    flow_to_suitable_area |>
    transmute(dataset, comid, flow_cfs, area_tot_ft2,
              area_wua_ft2 = area_wua_ft2_hqt,
              hsi_frac = hsi_frac_hqt)
}
  
# Limit to flow ranges covered by all training data, and interpolate gaps
interp_flows <- seq(300,15000,100)
if(TRUE){
  flow_to_suitable_area <- 
    flow_to_suitable_area |>
    group_by(dataset, comid) |>
    complete(flow_cfs = interp_flows) |>
    arrange(dataset, comid, flow_cfs) |>
    mutate(across(c(area_tot_ft2, area_wua_ft2, hsi_frac), function(var) zoo::na.approx(var, x = flow_cfs, na.rm=F))) |>
    filter(flow_cfs %in% interp_flows) |>
    filter(!is.na(hsi_frac))
}

train_data <- flowlines |> st_drop_geometry() |>
  left_join(flowline_attributes, by=join_by("comid"), relationship="one-to-one") |>
  inner_join(flow_to_suitable_area, by=join_by("comid"), relationship="one-to-many") |> 
  filter(hqt_gradient_class != "Valley Lowland") |>
  # eliminate reaches with near zero mean annual flow
  filter(erom_q_ma_cfs>10) |>
  glimpse()

flowlines |> st_zm() |>
  filter(comid %in% unique(flow_to_suitable_area$comid)) |>
  left_join(flowline_attributes, by=join_by("comid"), relationship="one-to-one") |>
  ggplot() + geom_sf(aes(color = hqt_gradient_class)) + theme(legend.title=element_blank())
```

## Exploration and PCA of training data

```{r pca}
df <- flowline_attributes |>
  select(
         # predictors of flow (as would be found in a regional regression)
         slope, da_area_sq_km, da_elev_mean, da_ppt_mean_mm, 
         # baseflow and peakflow statistics
         nf_bfl_dry_cfs, nf_bfl_wet_cfs, erom_q_ma_cfs, peak_q2_cfs, peak_q5_cfs, peak_q10_cfs,
         # flow and channel characteristics, hydrologically predicted
         bf_depth_m, bf_w_d_ratio, erom_v_ma_fps,
         # misc characteristics of the catchment
         da_avg_slope, da_k_erodibility, mean_ndvi,
         # misc characteristics of the locality
         loc_bfi, loc_pct_clay, loc_pct_sand, loc_permeability, loc_bedrock_depth, loc_ppt_mean_mm,
         # channel confinement characteristics
         mtpi30_min, vb_width_transect, vb_bf_w_ratio, 
         # channel sinuosity
         sinuosity,
         # aquatic
         bio_aq_rank_sw, species, species_fish
         ) |> 
  na.omit() |> 
  scale() 

df |> cor() |> corrr::rplot() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
# uniqueness: 
# df |> cor() |> abs() |> rowMeans() |> rank() 
pca <- princomp(df) 
#summary(pca)

factoextra::fviz_eig(pca, addlabels = TRUE)
factoextra::fviz_cos2(pca, choice = "var", axes = 1:2)
factoextra::fviz_pca_var(pca, axes=c(1,2), col.var = "cos2", repel = TRUE)
factoextra::fviz_pca_var(pca, axes=c(1,3), col.var = "cos2", repel = TRUE)
factoextra::fviz_pca_var(pca, axes=c(2,3), col.var = "cos2", repel = TRUE)

# flowlines |> GGally::ggpairs()
```

## Set up training data

```{r td}
td <- train_data |>
  mutate(flow_norm_cfs = flow_cfs / erom_q_ma_cfs) |> # flow as a percent of mean annual flow
  mutate(wua_per_lf = area_wua_ft2 / (reach_length_km*3280.84),
         log_wua_per_lf = log(wua_per_lf)) |> # transect-wise habitat area per linear foot
  transmute(dataset, comid, 
         # suitable habitat area normalized by reach length
         hsi_frac, wua_per_lf, log_wua_per_lf,
         # flow cfs normalized by mean annual flow
         flow_cfs, flow_norm_cfs,
         # predictors of flow (as would be found in a regional regression)
         slope, da_area_sq_km, da_elev_mean, da_ppt_mean_mm, 
         # baseflow and peakflow statistics
         nf_bfl_dry_cfs, nf_bfl_wet_cfs, erom_q_ma_cfs, peak_q2_cfs, peak_q5_cfs, peak_q10_cfs,
         # flow and channel characteristics, hydrologically predicted
         bf_depth_m, bf_w_d_ratio, erom_v_ma_fps,
         # misc characteristics of the catchment
         da_avg_slope, da_k_erodibility, mean_ndvi,
         # misc characteristics of the locality
         loc_bfi, loc_pct_clay, loc_pct_sand, loc_permeability, loc_bedrock_depth, loc_ppt_mean_mm,
         # channel confinement characteristics
         mtpi30_min, vb_width_transect, vb_bf_w_ratio, frac_leveed_longitudinal,
         # channel sinuosity
         sinuosity,
         # aquatic
         bio_aq_rank_sw, species, species_fish,
         # fixed effects
         hqt_gradient_class=droplevels(hqt_gradient_class), hyd_cls=droplevels(hyd_cls)
         ) |> 
  mutate(nf_bfl_dry_cfs_norm = nf_bfl_dry_cfs/erom_q_ma_cfs, 
         nf_bfl_wet_cfs_norm = nf_bfl_wet_cfs/erom_q_ma_cfs) |>
  drop_na() |> glimpse()
```

```{r td-ranges}
td |> 
  group_by(dataset) |>
  select_if(is.numeric) |>
  pivot_longer(cols = hsi_frac:last_col()) |>
  group_by(dataset, name) |>
  summarize(min = min(value), max=max(value))

domain <- td |> 
  select_if(is.numeric) |>
  select(-flow_cfs) |>
  pivot_longer(cols = hsi_frac:last_col()) |>
  group_by(name) |>
  summarize(min=min(value), max=max(value)) |>
  mutate(value = map2(min, max, function(x, y) list(min=x, max=y))) |>
  select(name, value) |> deframe()
```

```{r td-gridplot}
interp_flows_plot <- seq(700,5000,100)
td |> 
  # interpolate so that we are comparing like flows with like
  group_by(dataset, comid) |>
  complete(flow_cfs = interp_flows_plot) |>
  arrange(dataset, comid, flow_cfs) |>
  mutate(across(hsi_frac:last_col(), function(var) zoo::na.approx(var, x = flow_cfs))) |>
  filter(flow_cfs %in% interp_flows_plot) |>
  ungroup() |>
  # summarize by quantile of each variable
  mutate(across(slope:last_col(), function(x) factor(cut(x, breaks=5, labels=F)))) |>
  pivot_longer(cols=slope:last_col(), names_to="varname", values_to="qtile") |>
  mutate(qtile = factor(qtile, levels=c(1,2,3,4,5))) |>
  group_by(flow_cfs, varname, qtile) |> summarize(wua_per_lf=mean(wua_per_lf)) |> ungroup() |>
  ggplot() + geom_line(aes(x=flow_cfs, y=wua_per_lf, color=qtile)) + facet_wrap(~varname) + scale_color_viridis_d() + scale_y_log10() + scale_x_log10()
```

## Split into training and testing datasets
```{r td-split}
set.seed(47)
td_split <- group_initial_split(td |> mutate(strata=paste(hyd_cls, dataset)), strata=strata, group=comid)

test_comids <- testing(td_split) |> group_by(dataset, hyd_cls, comid) |> summarize() |> group_by(dataset, hyd_cls) |> filter(row_number()<=6) |> pull(comid)
test_comids
```

## Modeling

### Linear regression

```{r lm-spec}
lm_spec <- linear_reg() 
```

First set up recipes (sets of predictor and response variables), then train and predict

#### Scale-independent (%HSI versus dimensionless flow)

Scale-independent model version predicts the percent suitable habitat area using normalized flow, and without using any direct correlates of watershed size.

Technically this should not use linear regression as it is a proportion between zero and one, it should be using beta regression via the betareg package.

```{r lm-si, message=FALSE}
si_rec <- recipe(data=training(td_split), 
      formula=hsi_frac ~ 
        # flow as percent of mean annual flow
        flow_norm_cfs + 
        # channel characteristics: gradient and sinuosity
        slope + sinuosity + 
        # flow and channel characteristics, hydrologically predicted
        erom_v_ma_fps + bf_depth_m + bf_w_d_ratio + 
        # misc characteristics of the catchment
        da_k_erodibility + da_avg_slope + mean_ndvi +
        # misc characteristics of the locality
        loc_bfi + loc_pct_clay + loc_pct_sand + loc_permeability + loc_bedrock_depth + loc_ppt_mean_mm +
        # channel confinement
        mtpi30_min + vb_bf_w_ratio + frac_leveed_longitudinal +
        # baseflow as percent of mean annual flow
        nf_bfl_dry_cfs_norm + nf_bfl_wet_cfs_norm + 
        # aquatic
        bio_aq_rank_sw + species + species_fish + 
        # fixed effects
        hqt_gradient_class + hyd_cls
        ) |>
  step_log(all_numeric_predictors(), -mtpi30_min, -mean_ndvi, -bio_aq_rank_sw, -vb_bf_w_ratio, -frac_leveed_longitudinal) |>
  # let effect of flow vary with gradient
  # step_interact(terms = ~ flow_norm_cfs:slope) #|>
  step_dummy(hqt_gradient_class) |>
  step_dummy(hyd_cls) |>
  # PCA before interactions with flow
  # step_pca(all_numeric_predictors(), -flow_norm_cfs, num_comp = 4) |>
  # try interacting with all numeric predictors
  step_interact(terms = ~ flow_norm_cfs:all_predictors()) |>
  # let all effects vary with class
  #step_interact(terms = ~hqt_gradient_class:all_numeric_predictors())
  step_naomit(all_predictors()) |>
  step_zv(all_predictors()) |>
  step_normalize(all_numeric_predictors()) #, -flow_cfs)

lm_si <- workflow() |>
  add_recipe(si_rec) |>
  add_model(lm_spec) |>
  fit(data=training(td_split))

lm_si |> glance()
lm_si |> tidy()
lm_si$fit$fit |> dotwhisker::dwplot()

lm_si_res <- testing(td_split) |>
  mutate(hsi_frac_pred = predict(lm_si, testing(td_split))[[".pred"]]) |>
  select(comid, hsi_frac, hsi_frac_pred, flow_norm_cfs) |>
  # crude constraint to between 0 and 1; better to use betareg
  mutate(hsi_frac_pred=pmax(pmin(hsi_frac_pred,1),0))

lm_si_res |>
  arrange(comid, flow_norm_cfs) |>
  ggplot() + geom_path(aes(x=hsi_frac, y=hsi_frac_pred, color=flow_norm_cfs, group=comid), linewidth=2) + 
  coord_fixed() + geom_abline() + scale_color_viridis_c(trans="log10", name="Frac. of MAF") + 
  ggtitle("Scale-independent model, linear regression") +
  xlab("HSI (Frac. of Channel Area) (Actual)") + ylab("HSI (Frac. of Channel Area) (Predicted)")

lm_si_res |> 
  left_join(select(flowline_attributes, comid, erom_q_ma_cfs)) |>
  mutate(flow_cfs = flow_norm_cfs * erom_q_ma_cfs) |>
  filter(flow_cfs %in% c(300,500,1000,1500,3000,5000,10000,15000)) |>
  ggplot() + geom_point(aes(x=hsi_frac, y=hsi_frac_pred, color=flow_cfs)) + 
  facet_wrap(~flow_cfs, ncol = 4) +
  coord_fixed() + geom_abline() + scale_color_viridis_c(name = "Flow (cfs)") + 
  ggtitle("Scale-independent model, linear regression") +
  xlab("HSI (Frac. of Channel Area) (Actual)") + ylab("HSI (Frac. of Channel Area) (Predicted)")

lm_si_res |> filter(comid %in% test_comids) |>
  left_join(transmute(flowline_attributes, comid, erom_q_ma_cfs)) |> #, chan_width_ft=chan_width_m/0.3048)) |>
  ggplot(aes(x=flow_norm_cfs*erom_q_ma_cfs, group=1)) + facet_wrap(~comid) + 
  geom_line(aes(y=hsi_frac_pred*100, color="predicted", group=1)) + geom_line(aes(y=hsi_frac*100, color="actual", group=1)) + 
  xlab("Flow (cfs)") + ylab("Suitable Habitat Area (% of Total Area)")
```

```{r lm-si-coeffs, fig.width=6.5, fig.height=6.5, dpi=300}
lm_si |>
  tidy() |>
  mutate(ixn_class = if_else(str_detect(term, "flow_norm_cfs_x_"), "term_flow_norm_cfs_x_", "term") |> 
                     factor(levels = c("term", "term_flow_norm_cfs_x_"), labels = c("Constant Effect", "Interaction with Flow (cfs)")),
         term_root = str_replace(term, "flow_norm_cfs_x_", ""),
         coeff_dir = if_else(estimate>0, "Positive", "Negative"),
         ci_lower = estimate - qt(0.975, df = glance(lm_si)$df) * std.error,
         ci_upper = estimate + qt(0.975, df = glance(lm_si)$df) * std.error,
         estimate_abs = abs(estimate),
         ci_lower_abs = ci_lower * if_else(coeff_dir=="Negative",-1,1),
         ci_upper_abs = ci_upper * if_else(coeff_dir=="Negative",-1,1),
         is_significant = sign(ci_lower)==sign(ci_upper)
         ) |>
  filter(term!="flow_norm_cfs" & term!="(Intercept)") |>
  ggplot(aes(y = term_root, color = coeff_dir)) +
  facet_wrap(~ixn_class, ncol = 2) + 
  geom_errorbarh(aes(xmin = ci_lower_abs, xmax=ci_upper_abs)) + #pmax(0,ci_lower_abs), xmax = pmax(0,ci_upper_abs))) +
  geom_point(aes(x = estimate_abs, fill = is_significant), shape=21) +
  theme(legend.position="top",
        panel.grid.minor = element_blank(),
        panel.border = element_rect(fill = NA, color = "black"), 
        legend.direction = "vertical") +
  scale_y_discrete(limits=rev) + 
  scale_fill_manual(name = "Significant (95%)", values = c("FALSE" = "white", "TRUE" = "black")) +
  scale_color_manual(name = "Direction of Effect", values = c("Negative" = "goldenrod", "Positive" = "darkcyan")) +
  xlab("Coefficient Magnitude") + ylab("")
``` 

#### Scale-dependent (WUA-per-linear-ft versus flow)

Scale-dependent model version predicts the WUA per linear foot, using absolute flow values as well as watershed scale parameters.

* interact flow and mean annual flow; flow and area/elev/ppt; to add scale dependence to the flow effect

```{r lm-sd, message=FALSE}
sd_rec <- recipe(data=training(td_split), 
      formula = log_wua_per_lf ~ 
        # flow (cfs)
        flow_cfs + 
        # predictors of flow (catchment area, elevation, and MAP) -- attributes for gradient and upstream drainage area are interrelated
        slope + da_area_sq_km + da_elev_mean + da_ppt_mean_mm +
        # baseflow/peak flow statistics
        nf_bfl_dry_cfs + nf_bfl_wet_cfs + erom_q_ma_cfs + #log(peak_q2_cfs)  + log(peak_q5_cfs) + log(peak_q10_cfs) +
        # flow and channel characteristics, hydrologically predicted
        erom_v_ma_fps + bf_depth_m + bf_w_d_ratio + 
        # misc characteristics of the catchment
        da_k_erodibility + da_avg_slope + mean_ndvi +
        # misc characteristics of the locality
        loc_bfi + loc_pct_clay + loc_pct_sand + loc_permeability + loc_bedrock_depth + loc_ppt_mean_mm +
        # channel confinement characteristics
        mtpi30_min + vb_bf_w_ratio + frac_leveed_longitudinal + vb_width_transect +
        # channel sinuosity
        sinuosity + 
        # aquatic
        bio_aq_rank_sw + species + species_fish + 
        # fixed effects
        hqt_gradient_class + hyd_cls
        ) |>
  step_log(all_numeric_predictors(), -mtpi30_min, -mean_ndvi, -vb_bf_w_ratio, -frac_leveed_longitudinal) |>
  #step_interact(terms = ~ flow_cfs:erom_q_ma_cfs) |>
  step_interact(terms = ~ slope:da_area_sq_km) |>
  #step_interact(terms = ~ flow_cfs:da_area_sq_km) |>
  #step_interact(terms = ~ flow_cfs:da_elev_mean) |>
  #step_interact(terms = ~ flow_cfs:da_ppt_mean_mm) #|>
  # let all effects vary with class
  #step_interact(terms = ~hqt_gradient_class:all_numeric_predictors())
  step_dummy(hqt_gradient_class) |>
  step_dummy(hyd_cls) |>
  # PCA before interaction with flow 
  # step_pca(all_numeric_predictors(), -flow_cfs, num_comp = 4) |>
  # intract all terms with flow
  step_interact(terms = ~flow_cfs:all_predictors()) |>
  step_naomit(all_predictors()) |>
  step_zv(all_predictors()) |>
  step_normalize(all_numeric_predictors())#, -flow_cfs)

lm_sd <- workflow() |>
  add_recipe(sd_rec) |>
  add_model(lm_spec) |>
  fit(data=training(td_split))

lm_sd |> glance()
lm_sd |> tidy()
lm_sd$fit$fit |> dotwhisker::dwplot()

lm_sd_res <-
testing(td_split) |>
  mutate(log_wua_per_lf_pred = predict(lm_sd, testing(td_split))[[".pred"]]) |>
  transmute(comid, wua_per_lf=exp(log_wua_per_lf), wua_per_lf_pred=exp(log_wua_per_lf_pred), flow_cfs) 

lm_sd_res |>
  arrange(comid, flow_cfs) |>
  ggplot() + geom_path(aes(x=wua_per_lf, y=wua_per_lf_pred, color=flow_cfs, group=comid), linewidth=2) + 
  scale_y_log10() + scale_x_log10() +  coord_fixed() + geom_abline() + scale_color_viridis_c(name="Flow (cfs)") + 
  ggtitle("Scale-dependent model, linear regression") +
  xlab("WUA per LF (Actual)") + ylab("WUA per LF (Predicted)")

lm_sd_res |>
  filter(flow_cfs %in% c(300,500,1000,1500,3000,5000,10000,15000)) |>
  ggplot() + geom_point(aes(x=wua_per_lf, y=wua_per_lf_pred, color=flow_cfs)) + 
  facet_wrap(~flow_cfs, ncol = 4) +
  scale_y_log10() + scale_x_log10() +  coord_fixed() + geom_abline() + scale_color_viridis_c(name="Flow (cfs)") + 
  ggtitle("Scale-dependent model, linear regression") +
  xlab("WUA per LF (Actual)") + ylab("WUA per LF (Predicted)")

lm_sd_res |> filter(comid %in% test_comids) |>
  ggplot(aes(x=flow_cfs, group=1)) + facet_wrap(~comid) + 
  geom_line(aes(y=wua_per_lf_pred, color="predicted", group=1)) + geom_line(aes(y=wua_per_lf, color="actual", group=1)) + 
  scale_y_log10() + xlab("Flow (cfs)") + ylab("Suitable Habitat Area per LF Channel (ft)")

```

```{r lm-sd-coeffs, fig.width=6.5, fig.height=6.5, dpi=300}
lm_sd |>
  tidy() |>
  mutate(ixn_class = if_else(str_detect(term, "flow_cfs_x_"), "term_flow_cfs_x_", "term") |> 
                     factor(levels = c("term", "term_flow_cfs_x_"), labels = c("Constant Effect", "Interaction with Flow (cfs)")),
         term_root = str_replace(term, "flow_cfs_x_", ""),
         coeff_dir = if_else(estimate>0, "Positive", "Negative"),
         ci_lower = estimate - qt(0.975, df = glance(lm_sd)$df) * std.error,
         ci_upper = estimate + qt(0.975, df = glance(lm_sd)$df) * std.error,
         estimate_abs = abs(estimate),
         ci_lower_abs = ci_lower * if_else(coeff_dir=="Negative",-1,1),
         ci_upper_abs = ci_upper * if_else(coeff_dir=="Negative",-1,1),
         is_significant = sign(ci_lower)==sign(ci_upper)
         ) |>
  filter(term!="flow_cfs" & term!="(Intercept)") |>
  ggplot(aes(y = term_root, color = coeff_dir)) +
  facet_wrap(~ixn_class, ncol = 2) + 
  geom_errorbarh(aes(xmin = ci_lower_abs, xmax=ci_upper_abs)) + #pmax(0,ci_lower_abs), xmax = pmax(0,ci_upper_abs))) +
  geom_point(aes(x = estimate_abs, fill = is_significant), shape=21) +
  theme(legend.position="top",
        panel.grid.minor = element_blank(),
        panel.border = element_rect(fill = NA, color = "black"), 
        legend.direction = "vertical") +
  scale_y_discrete(limits=rev) + 
  scale_fill_manual(name = "Significant (95%)", values = c("FALSE" = "white", "TRUE" = "black")) +
  scale_color_manual(name = "Direction of Effect", values = c("Negative" = "goldenrod", "Positive" = "darkcyan")) +
  xlab("Coefficient Magnitude") + ylab("")
``` 

### Linear regression with regularized (lasso) feature selection 

#### Scale-independent (%HSI versus dimensionless flow)

```{r lasso-si, message=FALSE}
drop_vars <- workflow() |>
  add_recipe(si_rec) |>
  add_model(linear_reg(penalty = 0.005, mixture = 1) |> set_engine("glmnet")) |>
  fit(data=training(td_split)) |>
  tidy() |> 
  filter(estimate==0) |> 
  pull(term)

lasso_si <- workflow() |>
  add_recipe(si_rec |> step_rm(all_of(drop_vars))) |>
  add_model(lm_spec) |>
  fit(data=training(td_split))

lasso_si |> glance()
lasso_si |> tidy()
lasso_si$fit$fit |> dotwhisker::dwplot()

lasso_si_res <- 
testing(td_split) |>
  mutate(hsi_frac_pred = predict(lasso_si, testing(td_split))[[".pred"]]) |>
  select(comid, hsi_frac, hsi_frac_pred, flow_norm_cfs) |>
  # crude constraint to between 0 and 1; better to use betareg
  mutate(hsi_frac_pred=pmax(pmin(hsi_frac_pred,1),0))

lasso_si_res |>
  arrange(comid, flow_norm_cfs) |>
  ggplot() + geom_path(aes(x=hsi_frac, y=hsi_frac_pred, color=flow_norm_cfs, group=comid), linewidth=2) + 
  coord_fixed() + geom_abline() + scale_color_viridis_c(trans="log10", name="Frac. of MAF") + 
  ggtitle("Scale-independent model, linear regression w/ lasso feature selection") +
  xlab("HSI (Frac. of Channel Area) (Actual)") + ylab("HSI (Frac. of Channel Area) (Predicted)")

lasso_si_res |> 
  left_join(select(flowline_attributes, comid, erom_q_ma_cfs)) |>
  mutate(flow_cfs = flow_norm_cfs * erom_q_ma_cfs) |>
  filter(flow_cfs %in% c(300,500,1000,1500,3000,5000,10000,15000)) |>
  ggplot() + geom_point(aes(x=hsi_frac, y=hsi_frac_pred, color=flow_cfs)) + 
  facet_wrap(~flow_cfs, ncol = 4) +
  coord_fixed() + geom_abline() + scale_color_viridis_c(name = "Flow (cfs)") + 
  ggtitle("Scale-independent model, linear regression w/ lasso feature selection") +
  xlab("HSI (Frac. of Channel Area) (Actual)") + ylab("HSI (Frac. of Channel Area) (Predicted)")

lasso_si_res |> filter(comid %in% test_comids) |>
  left_join(transmute(flowline_attributes, comid, erom_q_ma_cfs)) |> #, chan_width_ft=chan_width_m/0.3048)) |>
  ggplot(aes(x=flow_norm_cfs*erom_q_ma_cfs, group=1)) + facet_wrap(~comid) + 
  geom_line(aes(y=hsi_frac_pred*100, color="predicted", group=1)) + geom_line(aes(y=hsi_frac*100, color="actual", group=1)) + 
  xlab("Flow (cfs)") + ylab("Suitable Habitat Area (% of Total Area)")
```

```{r lasso-si-coeffs, fig.width=6.5, fig.height=6.5, dpi=300}
lasso_si |>
  tidy() |>
  mutate(ixn_class = if_else(str_detect(term, "flow_norm_cfs_x_"), "term_flow_norm_cfs_x_", "term") |> 
                     factor(levels = c("term", "term_flow_norm_cfs_x_"), labels = c("Constant Effect", "Interaction with Flow (cfs)")),
         term_root = str_replace(term, "flow_norm_cfs_x_", ""),
         coeff_dir = if_else(estimate>0, "Positive", "Negative"),
         ci_lower = estimate - qt(0.975, df = glance(lasso_si)$df) * std.error,
         ci_upper = estimate + qt(0.975, df = glance(lasso_si)$df) * std.error,
         estimate_abs = abs(estimate),
         ci_lower_abs = ci_lower * if_else(coeff_dir=="Negative",-1,1),
         ci_upper_abs = ci_upper * if_else(coeff_dir=="Negative",-1,1),
         is_significant = sign(ci_lower)==sign(ci_upper)
         ) |>
  filter(term!="flow_norm_cfs" & term!="(Intercept)") |>
  ggplot(aes(y = term_root, color = coeff_dir)) +
  facet_wrap(~ixn_class, ncol = 2) + 
  geom_errorbarh(aes(xmin = ci_lower_abs, xmax=ci_upper_abs)) + #pmax(0,ci_lower_abs), xmax = pmax(0,ci_upper_abs))) +
  geom_point(aes(x = estimate_abs, fill = is_significant), shape=21) +
  theme(legend.position="top",
        panel.grid.minor = element_blank(),
        panel.border = element_rect(fill = NA, color = "black"), 
        legend.direction = "vertical") +
  scale_y_discrete(limits=rev) + 
  scale_fill_manual(name = "Significant (95%)", values = c("FALSE" = "white", "TRUE" = "black")) +
  scale_color_manual(name = "Direction of Effect", values = c("Negative" = "goldenrod", "Positive" = "darkcyan")) +
  xlab("Coefficient Magnitude") + ylab("")
``` 

#### Scale-dependent (WUA-per-linear-ft versus flow)

```{r lasso-sd, message=FALSE}
drop_vars <- workflow() |>
  add_recipe(sd_rec) |>
  add_model(linear_reg(penalty = 0.005, mixture = 1) |> set_engine("glmnet")) |>
  fit(data=training(td_split)) |>
  tidy() |> 
  filter(estimate==0) |> 
  pull(term)

lasso_sd <- workflow() |>
  add_recipe(sd_rec |>
               step_rm(all_of(drop_vars))) |>
  add_model(lm_spec) |>
  fit(data=training(td_split))

lasso_sd |> glance()
lasso_sd |> tidy()
lasso_sd$fit$fit |> dotwhisker::dwplot()

lasso_sd_res <- 
testing(td_split) |>
  mutate(log_wua_per_lf_pred = predict(lasso_sd, testing(td_split))[[".pred"]]) |>
  transmute(comid, wua_per_lf=exp(log_wua_per_lf), wua_per_lf_pred=exp(log_wua_per_lf_pred), flow_cfs) 

lasso_sd_res |>
  arrange(comid, flow_cfs) |>
  ggplot() + geom_path(aes(x=wua_per_lf, y=wua_per_lf_pred, color=flow_cfs, group=comid), linewidth=2) + 
  scale_y_log10() + scale_x_log10() +  coord_fixed() + geom_abline() + scale_color_viridis_c(name="Flow (cfs)")   + 
  ggtitle("Scale-dependent model, linear regression w/ lasso feature selection") +
  xlab("WUA per LF (Actual)") + ylab("WUA per LF (Predicted)")

lasso_sd_res |>
  filter(flow_cfs %in% c(300,500,1000,1500,3000,5000,10000,15000)) |>
  ggplot() + geom_point(aes(x=wua_per_lf, y=wua_per_lf_pred, color=flow_cfs)) + 
  facet_wrap(~flow_cfs, ncol = 4) +
  scale_y_log10() + scale_x_log10() +  coord_fixed() + geom_abline() + scale_color_viridis_c(name="Flow (cfs)") + 
  ggtitle("Scale-dependent model, linear regression w/ lasso feature selection") +
  xlab("WUA per LF (Actual)") + ylab("WUA per LF (Predicted)")

lasso_sd_res |> filter(comid %in% test_comids) |>
  ggplot(aes(x=flow_cfs, group=1)) + facet_wrap(~comid) + 
  geom_line(aes(y=wua_per_lf_pred, color="predicted", group=1)) + geom_line(aes(y=wua_per_lf, color="actual", group=1)) + 
  scale_y_log10() + xlab("Flow (cfs)") + ylab("Suitable Habitat Area per LF Channel (ft)")
```
```{r lasso-sd-coeffs, fig.width=6.5, fig.height=6.5, dpi=300}
lasso_sd |>
  tidy() |>
  mutate(ixn_class = if_else(str_detect(term, "flow_cfs_x_"), "term_flow_cfs_x_", "term") |> 
                     factor(levels = c("term", "term_flow_cfs_x_"), labels = c("Constant Effect", "Interaction with Flow (cfs)")),
         term_root = str_replace(term, "flow_cfs_x_", ""),
         coeff_dir = if_else(estimate>0, "Positive", "Negative"),
         ci_lower = estimate - qt(0.975, df = glance(lasso_sd)$df) * std.error,
         ci_upper = estimate + qt(0.975, df = glance(lasso_sd)$df) * std.error,
         estimate_abs = abs(estimate),
         ci_lower_abs = ci_lower * if_else(coeff_dir=="Negative",-1,1),
         ci_upper_abs = ci_upper * if_else(coeff_dir=="Negative",-1,1),
         is_significant = sign(ci_lower)==sign(ci_upper)
         ) |>
  filter(term!="flow_cfs" & term!="(Intercept)") |>
  ggplot(aes(y = term_root, color = coeff_dir)) +
  facet_wrap(~ixn_class, ncol = 2) + 
  geom_errorbarh(aes(xmin = ci_lower_abs, xmax=ci_upper_abs)) + #pmax(0,ci_lower_abs), xmax = pmax(0,ci_upper_abs))) +
  geom_point(aes(x = estimate_abs, fill = is_significant), shape=21) +
  theme(legend.position="top",
        panel.grid.minor = element_blank(),
        panel.border = element_rect(fill = NA, color = "black"), 
        legend.direction = "vertical") +
  scale_y_discrete(limits=rev) + 
  scale_fill_manual(name = "Significant (95%)", values = c("FALSE" = "white", "TRUE" = "black")) +
  scale_color_manual(name = "Direction of Effect", values = c("Negative" = "goldenrod", "Positive" = "darkcyan")) +
  xlab("Coefficient Magnitude") + ylab("")
``` 

### Random Forest Regrssion

```{r rfr-spec}
#install.packages("ranger")
#rfr_spec <- rand_forest(mode = "regression", trees = 2^12)
```

#### Scale-independent (%HSI versus dimensionless flow)

```{r rfr-si-tune, eval=FALSE} 
# set up cross validation
folds <- vfold_cv(training(td_split), v = 10, strata = dataset)
val_metrics <- metric_set(rmse, rsq, ccc)

# tuned random forest: how many trees to use?
rfr_spec <- rand_forest(mode = "regression", trees = tune())
rfr_grid <- tibble(trees = 2^seq(0, 12, 1))

rfr_si_tune <- 
  workflow() |>
  add_recipe(si_rec) |>
  add_model(rfr_spec) |>
  tune_grid(resamples = folds,
            grid = rfr_grid,
            metrics = val_metrics)
rfr_si_tune |>
  collect_metrics() |>
  ggplot(aes(x = trees, y = mean, color = .metric)) +
  geom_errorbar(aes(ymin = mean - std_err,
                    ymax = mean + std_err),
                width = 0.05) +
  geom_line() +
  facet_wrap(. ~ .metric, ncol = 1, scales="free") +
  theme(legend.position = "none") + scale_x_log10()
```

```{r rfr-si, message=FALSE}
rfr_spec <- rand_forest(mode = "regression", trees = 2^8)

rfr_si <- workflow() |>
  add_recipe(si_rec) |>
  add_model(rfr_spec) |>
  fit(data=training(td_split))

rfr_si$fit$fit |> print()

rfr_si_res <- 
testing(td_split) |>
  mutate(hsi_frac_pred = predict(rfr_si, testing(td_split))[[".pred"]]) |>
  select(comid, hsi_frac, hsi_frac_pred, flow_norm_cfs) |>
  # crude constraint to between 0 and 1; better to use betareg
  mutate(hsi_frac_pred=pmax(pmin(hsi_frac_pred,1),0))

rfr_si_res |>
  arrange(comid, flow_norm_cfs) |>
  ggplot() + geom_path(aes(x=hsi_frac, y=hsi_frac_pred, color=flow_norm_cfs, group=comid), linewidth=2) + 
  coord_fixed() + geom_abline() + scale_color_viridis_c(trans="log10", name="Frac. of MAF") + 
  ggtitle("Scale-independent model, random forest regression") +
  xlab("HSI (Frac. of Channel Area) (Actual)") + ylab("HSI (Frac. of Channel Area) (Predicted)")

rfr_si_res |> 
  left_join(select(flowline_attributes, comid, erom_q_ma_cfs)) |>
  mutate(flow_cfs = flow_norm_cfs * erom_q_ma_cfs) |>
  filter(flow_cfs %in% c(300,500,1000,1500,3000,5000,10000,15000)) |>
  ggplot() + geom_point(aes(x=hsi_frac, y=hsi_frac_pred, color=flow_cfs)) + 
  facet_wrap(~flow_cfs, ncol = 4) +
  coord_fixed() + geom_abline() + scale_color_viridis_c(name = "Flow (cfs)") + 
  ggtitle("Scale-independent model, random forest regression") +
  xlab("HSI (Frac. of Channel Area) (Actual)") + ylab("HSI (Frac. of Channel Area) (Predicted)")

rfr_si_res |> filter(comid %in% test_comids) |>
  left_join(transmute(flowline_attributes, comid, erom_q_ma_cfs)) |> #, chan_width_ft=chan_width_m/0.3048)) |>
  ggplot(aes(x=flow_norm_cfs*erom_q_ma_cfs, group=1)) + facet_wrap(~comid) + 
  geom_line(aes(y=hsi_frac_pred*100, color="predicted", group=1)) + geom_line(aes(y=hsi_frac*100, color="actual", group=1)) + 
  xlab("Flow (cfs)") + ylab("Suitable Habitat Area (% of Total Area)")
```

#### Scale-dependent (WUA-per-linear-ft versus flow)

```{r rfr-sd-tune, eval=FALSE}
rfr_spec <- rand_forest(mode = "regression", trees = tune())
rfr_grid <- tibble(trees = 2^seq(0, 12, 1))

rfr_sd_tune <- 
  workflow() |>
  add_recipe(sd_rec) |>
  add_model(rfr_spec) |>
  tune_grid(resamples = folds,
            grid = rfr_grid,
            metrics = val_metrics)
rfr_sd_tune |>
  collect_metrics() |>
  ggplot(aes(x = trees, y = mean, color = .metric)) +
  geom_errorbar(aes(ymin = mean - std_err,
                    ymax = mean + std_err),
                width = 0.05) +
  geom_line() +
  facet_wrap(. ~ .metric, ncol = 1, scales="free") +
  theme(legend.position = "none") + scale_x_log10()
```

```{r rfr-sd, message=FALSE}
rfr_spec <- rand_forest(mode = "regression", trees = 2^8)

rfr_sd <- workflow() |>
  add_recipe(sd_rec) |>
  add_model(rfr_spec) |>
  fit(data=training(td_split))

rfr_sd$fit$fit |> print()

rfr_sd_res <-
testing(td_split) |>
  mutate(log_wua_per_lf_pred = predict(rfr_sd, testing(td_split))[[".pred"]]) |>
  transmute(comid, wua_per_lf=exp(log_wua_per_lf), wua_per_lf_pred=exp(log_wua_per_lf_pred), flow_cfs) 

rfr_sd_res |>
  arrange(comid, flow_cfs) |>
  ggplot() + geom_path(aes(x=wua_per_lf, y=wua_per_lf_pred, color=flow_cfs, group=comid), linewidth=2) + 
  scale_y_log10() + scale_x_log10() +  coord_fixed() + geom_abline() + scale_color_viridis_c(name="Flow (cfs)")   + 
  ggtitle("Scale-dependent model, random forest regression") +
  xlab("WUA per LF (Actual)") + ylab("WUA per LF (Predicted)")

rfr_sd_res |>
  filter(flow_cfs %in% c(300,500,1000,1500,3000,5000,10000,15000)) |>
  ggplot() + geom_point(aes(x=wua_per_lf, y=wua_per_lf_pred, color=flow_cfs)) + 
  facet_wrap(~flow_cfs, ncol = 4) +
  scale_y_log10() + scale_x_log10() +  coord_fixed() + geom_abline() + scale_color_viridis_c(name="Flow (cfs)") + 
  ggtitle("Scale-dependent model, random forest regression") +
  xlab("WUA per LF (Actual)") + ylab("WUA per LF (Predicted)")

rfr_sd_res |> filter(comid %in% test_comids) |>
  ggplot(aes(x=flow_cfs, group=1)) + facet_wrap(~comid) + 
  geom_line(aes(y=wua_per_lf_pred, color="predicted", group=1)) + geom_line(aes(y=wua_per_lf, color="actual", group=1)) + 
  scale_y_log10() + xlab("Flow (cfs)") + ylab("Suitable Habitat Area per LF Channel (ft)") 
```

## Random Forest model prediction output

```{r prediction-data}
# assemble prediction dataset
pd <- flowline_attributes |> 
  # filter out Valley Lowland areas
  filter(hqt_gradient_class != "Valley Lowland") |>
  # filter for just the areas that are hydrologically similar to the training data
  filter(hyd_cls %in% c("High-volume snowmelt and rain",
                        "Low-volume snowmelt and rain")) |>
  # select the variables that are used in the model and drop NAs
  mutate(nf_bfl_dry_cfs_norm = coalesce(nf_bfl_dry_cfs/erom_q_ma_cfs,0), # normalized baseflow values
         nf_bfl_wet_cfs_norm = coalesce(nf_bfl_wet_cfs/erom_q_ma_cfs,0)) |>
  select(comid, any_of(c(sd_rec$var_info$variable, si_rec$var_info$variable))) |> drop_na() |>
  # create series of flow prediction points
  expand_grid(flow_cfs = c(0,signif(100*2^seq(-2,7,0.5),2))) |>
  mutate(flow_cfs = if_else(flow_cfs>0, flow_cfs, erom_q_ma_cfs)) |> # also evaluate at mean annual flow
  mutate(flow_norm_cfs = coalesce(flow_cfs/erom_q_ma_cfs,0)) |> # flow as a percent of mean annual flow
  arrange(comid, flow_cfs) |>
  glimpse()

# # get min/max ranges from the training data
# ranges <- td |> select(all_of(names(pd))) |> 
#   reframe(across(everything(), function(x) (range(x)))) 
# # filter the prediction data for just those falling within the training data
# pd_idx <- pd |> 
#   select(comid, all_of(names(ranges))) |>
#   mutate(across(-comid, function(x) x>ranges[[cur_column()]][[1]] & x<=ranges[[cur_column()]][[2]])) |>
#   filter(if_all(-comid)) |>
#   pull(comid)
# # update pd to exclude the out-of-range data
# pd <- pd |> filter(comid %in% pd_idx) |> glimpse()
```

```{r prediction-output}
sd_pred <- 
  sd_rec |> 
  prep(training(td_split)) |> 
  bake(pd) |> 
  bind_cols(pd |> transmute(comid, flow_cfs_orig=flow_cfs)) |>
  drop_na() %>% # need to use magrittr pipe for this purpose
  mutate(log_wua_per_lf_pred = predict(rfr_sd$fit$fit, new_data=.)[[".pred"]]) |>
  transmute(comid, flow_cfs=flow_cfs_orig, log_wua_per_lf_pred) |>
  mutate(wua_per_lf_pred = exp(log_wua_per_lf_pred)) |>
  glimpse()

si_pred <- 
  si_rec |> 
  prep(training(td_split)) |> 
  bake(pd) |> 
  bind_cols(pd |> transmute(comid, flow_norm_cfs_orig=flow_norm_cfs)) |>
  drop_na() %>% # need to use magrittr pipe for this purpose
  mutate(hsi_frac_pred = predict(rfr_si$fit$fit, new_data=.)[[".pred"]]) |>
  transmute(comid, flow_norm_cfs=flow_norm_cfs_orig, hsi_frac_pred) |>
  left_join(flowline_attributes |> transmute(comid, bf_width_ft = bf_width_m/0.3048), by=join_by(comid)) |>
  mutate(wua_per_lf_pred = hsi_frac_pred * bf_width_ft) |>
  glimpse()
```

Example plot of habitat area per linear foot at 800 cfs (scale-dependent model)

```{r map-output-sd, fig.height=6.5}
flowlines |> st_zm() |>
  inner_join(sd_pred |> filter(round(flow_cfs)==800), by=join_by(comid)) |>
  ggplot() + geom_sf(aes(color = wua_per_lf_pred)) + 
  scale_color_viridis_c(trans="log", breaks = scales::breaks_log(6))
```

Example plot of percent habitat area at mean annual flow (scale-independent model)

```{r map-output-si, fig.height=6.5}
flowlines |> st_zm() |>
  inner_join(si_pred |> filter(flow_norm_cfs==1), by=join_by(comid)) |>
  ggplot() + geom_sf(aes(color = hsi_frac_pred)) + 
  scale_color_viridis_c(breaks = scales::breaks_pretty(5))
```

## Predictions summarized by DSMHabitat reach

```{r dsmhabitat-prep}
mainstems_comid <- 
  st_read("/vsizip/rearing_spatial_data/nhdplusv2_comid_habitat_xw.shp.zip", as_tibble=T) |>
  janitor::clean_names() |>
  st_zm() |>
  st_transform(st_crs(flowlines)) |>
  mutate(length_ft = st_length(geometry) |> units::set_units("ft") |> units::drop_units()) |>
  filter(habitat=="rearing") |>
  left_join(flowline_attributes |> select(comid, hqt_gradient_class), by=join_by(comid)) |>
  filter(!(river %in% c("Sacramento River", "San Joaquin River")))

mainstems <-
  mainstems_comid |>
  group_by(river, hqt_gradient_class) |>
  summarize() 

mainstems_comid |> 
  ggplot() + 
  geom_sf(aes(color=river, linetype=hqt_gradient_class)) + 
  theme(legend.key.height = unit(12, "point"))
```

```{r dsmhabitat-join}
watersheds <- mainstems |> pull(river) |> unique()
watershed_name <- tolower(gsub(pattern = "-| ", replacement = "_", x = watersheds))
watershed_rda_name <- paste(watershed_name, "floodplain", sep = "_")

dsm_habitat_floodplain <- map_df(watershed_rda_name, function(watershed) {
  df <- as.data.frame(do.call(`::`, list(pkg = "DSMhabitat", name = watershed)))
}) |> 
  rename(river = watershed) |>
  mutate(FR_floodplain_m2 = FR_floodplain_acres * 4046.86,
         FR_floodplain_m2_suitable = DSMhabitat::apply_suitability(FR_floodplain_m2),
         FR_floodplain_acres_suitable = FR_floodplain_m2_suitable / 4046.86)

dsm_habitat_instream <- map_df(paste(watershed_name, "instream", sep = "_"), 
                               possibly(function(watershed) {
                                 df <- as.data.frame(do.call(`::`, list(pkg = "DSMhabitat", name = watershed)))
                                 }, otherwise = NULL)) |> 
  rename(river = watershed) 

dsm_flows <- bind_rows(dsm_habitat_floodplain, dsm_habitat_instream) |>
  group_by(river, flow_cfs) |>
  summarize() |>
  ungroup() |>
  arrange(river, flow_cfs)

pd_dsmhabitat <- flowline_attributes |> 
  # select the variables that are used in the model and drop NAs
  select(comid, any_of(sd_rec$var_info$variable)) |> 
  inner_join(select(mainstems_comid, comid, river)) |>
  # left join dsm_flows to bring in the exact flows used in DSMhabitat:
  # left_join(dsm_flows, by=join_by(river), relationship="many-to-many") |>
  # or use expand_grid to bring in the modeled flows instead:
  expand_grid(flow_cfs = interp_flows) |>
  arrange(comid, flow_cfs) |>
  filter(comid %in% mainstems_comid$comid) |>
  drop_na() #

mainstems_comid |> 
  st_zm() |> 
  filter(comid %in% pd_dsmhabitat$comid) |>
  ggplot() + 
  geom_sf(aes(color=river)) + 
  theme(legend.key.height = unit(12, "point"))
```

```{r dsmhabitat-val}
pd_dsmhabitat_pred <- 
  sd_rec |> 
  prep(training(td_split)) |> 
  bake(pd_dsmhabitat) |> 
  bind_cols(pd_dsmhabitat |> transmute(comid, flow_cfs_orig=flow_cfs)) |>
  drop_na() %>% # need to use magrittr pipe for this purpose
  mutate(log_wua_per_lf_pred = predict(rfr_sd$fit$fit, new_data=.)[[".pred"]]) |>
  transmute(comid, flow_cfs=flow_cfs_orig, log_wua_per_lf_pred) |>
  mutate(wua_per_lf_pred = exp(log_wua_per_lf_pred)) |>
  select(comid, 
         flow_cfs, wua_per_lf_pred)

pd_rf_by_mainstem <-
  mainstems_comid |> 
  st_drop_geometry() |>
  inner_join(pd_dsmhabitat_pred, by=join_by(comid), relationship="many-to-many") |>
  group_by(river, flow_cfs) |>
  summarize(tot_wua_per_lf_pred = sum(coalesce(wua_per_lf_pred,0)),
            tot_length_ft = sum(coalesce(length_ft,0))) |>
  mutate(tot_wua_ft2 = tot_wua_per_lf_pred * tot_length_ft,
         tot_wua_ac = tot_wua_ft2 / 43560,
         avg_wua_ft2_per_lf = tot_wua_ft2 / tot_length_ft,
         avg_wua_ft2_per_1000ft = 1000 * avg_wua_ft2_per_lf) 

pd_rf_by_mainstem |>
  ggplot() + 
  geom_line(aes(x = flow_cfs, y = tot_wua_ac, color = river)) + #, color=species, linetype=habitat)) + 
  facet_wrap(~river, scales="free_y") + 
  scale_x_log10(labels=scales::comma_format(), expand=c(0,0)) + 
  scale_y_continuous() + 
  theme(legend.position="none", 
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  xlab("Flow (cfs)") + ylab("Total WUA (ac)")

pd_rf_by_mainstem |>
  ggplot() + 
  geom_line(aes(x = flow_cfs, y = avg_wua_ft2_per_lf, color = river)) + #, color=species, linetype=habitat)) + 
  facet_wrap(~river, scales="free_y") + 
  scale_x_log10(labels=scales::comma_format(), expand=c(0,0)) + 
  scale_y_continuous() + 
  theme(legend.position="none", 
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  xlab("Flow (cfs)") + ylab("Suitable Habitat Area (ft2 per linear ft)")
```

```{r dsmhabitat-watershed-comp-floodplain}
pd_rf_by_mainstem |> 
  inner_join(dsm_habitat_floodplain, by=join_by(river==river, flow_cfs==flow_cfs)) |> 
  #filter(river == "American River") |> 
  ggplot() +
  geom_line(aes(x = flow_cfs, y = tot_wua_ac, color='modeled')) + 
  geom_line(aes(x = flow_cfs, y = FR_floodplain_acres_suitable, color = 'DSMhabitat')) +  
  facet_wrap(~river) + #, scales="free_y") + 
  scale_x_log10(labels=scales::comma_format(), expand=c(0,0)) + 
  scale_y_log10(labels=scales::comma_format(), expand=c(0,0)) + 
  theme(legend.position="top", legend.title=element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  xlab("Flow (cfs)") + ylab("Total WUA (ac)")
```

```{r dsmhabitat-watershed-comp-instream}
pd_rf_by_mainstem |> 
  inner_join(dsm_habitat_instream, by=join_by(river==river, flow_cfs==flow_cfs)) |> 
  ggplot() +
  geom_line(aes(x = flow_cfs, y = avg_wua_ft2_per_1000ft, color='modeled')) + 
  geom_line(aes(x = flow_cfs, y = FR_juv_wua, color = 'DSMhabitat')) +  
  facet_wrap(~river) + #, scales="free_y") + 
  scale_x_log10(labels=scales::comma_format(), expand=c(0,0)) + 
  scale_y_log10(labels=scales::comma_format(), expand=c(0,0)) + 
  theme(legend.position="top", legend.title=element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  xlab("Flow (cfs)") + ylab("Suitable Habitat Area (ft2 per linear ft)")
```

===

```{r}
knitr::knit_exit()
```

## TESTS

### Cross Validation

```{r}
# test cross validation on the linear model
lm_sd |> fit_resamples(resamples=folds, metrics=val_metrics) |> collect_metrics()

# run the same cross validation against all models
enframe(list("lm"=lm_sd, "lasso"=lasso_sd, "rfr"=rfr_sd), value="model") |> 
  mutate(result = map(model, function(x) {x |> 
      fit_resamples(resamples=folds, metrics=metric_set(rsq)) |> 
      collect_metrics()
    })) |> unnest(result)

```

### Check for overfit

```{r}
# check to make sure not overfitting

# rfr_sd |> predict(training(td_split)) |> bind_cols(training(td_split)) |> val_metrics(truth = wua_per_lf, estimate=.pred)
# rfr_sd |> predict(testing(td_split)) |> bind_cols(testing(td_split)) |> val_metrics(truth = wua_per_lf, estimate=.pred)
bind_rows(list(training=bind_cols(predict(rfr_sd, training(td_split)), training(td_split)), 
               testing=bind_cols(predict(rfr_sd, testing(td_split)), testing(td_split)))
          , .id="id") |>
  group_by(id)  |> 
  val_metrics(truth = wua_per_lf, estimate=.pred)

bind_rows(list(training=bind_cols(predict(rfr_si, training(td_split)), training(td_split)), 
               testing=bind_cols(predict(rfr_si, testing(td_split)), testing(td_split)))
          , .id="id") |>
  group_by(id)  |> 
  val_metrics(truth = hsi_frac, estimate=.pred)
```

```{r}
```

### Series interpolated (panel) regression

```{r}
# interpolate series so there is a value fo every 100 flows
td_series <- td |>
  group_by(comid) |>
  arrange(comid, flow_cfs) |>
  complete(flow_cfs = seq(from = min(flow_cfs), to = max(flow_cfs), by = 100)) |>
  mutate(across(hsi_frac:last_col(), function(var) zoo::na.approx(var, x = flow_cfs))) |>
  filter(flow_cfs%%100==0) |> ungroup()

sd_series_rec <- recipe(data=td_series, 
      formula = log_wua_per_lf ~ 
        # flow (cfs)
        flow_cfs + 
        # predictors of flow (catchment area, elevation, and MAP) -- attributes for gradient and upstream drainage area are interrelated
        slope + da_area_sq_km + da_elev_mean + da_ppt_mean_mm +
        # baseflow/peak flow statistics
        nf_bfl_dry_cfs + nf_bfl_wet_cfs + erom_q_ma_cfs + #log(peak_q2_cfs)  + log(peak_q5_cfs) + log(peak_q10_cfs) +
        # flow and channel characteristics, hydrologically predicted
        erom_v_ma_fps + bf_depth_m + bf_w_d_ratio + 
        # misc characteristics of the catchment
        da_k_erodibility + da_avg_slope + mean_ndvi +
        # misc characteristics of the locality
        loc_bfi + loc_pct_clay + loc_pct_sand + loc_permeability + loc_bedrock_depth + loc_ppt_mean_mm +
        # channel confinement characteristics
        mtpi30_min + #log(vb_width_transect) + log(vb_bf_w_ratio) +
        # channel sinuosity
        sinuosity
        ) |>
  step_log(all_numeric_predictors(), -mtpi30_min) |>
  step_interact(terms = ~ flow_cfs:erom_q_ma_cfs) |>
  step_interact(terms = ~ slope:da_area_sq_km) |>
  step_interact(terms = ~ flow_cfs:da_area_sq_km) |>
  step_interact(terms = ~ flow_cfs:da_elev_mean) |>
  step_interact(terms = ~ flow_cfs:da_ppt_mean_mm) |>
  step_lag(flow_cfs) |>
  step_interact(terms = ~ flow_cfs:lag_1_flow_cfs) 
  #step_interact(terms = ~ lag_1_flow_cfs:all_numeric_predictors())

lm_sd_series <- workflow() |>
  add_recipe(sd_series_rec) |>
  add_model(lm_spec) |>
  fit(data=td_series)

lm_sd_series |> glance()
lm_sd_series |> tidy()

td_series |>
  mutate(log_wua_per_lf_pred = predict(lm_sd_series, td_series)[[".pred"]]) |>
  transmute(wua_per_lf=exp(log_wua_per_lf), wua_per_lf_pred=exp(log_wua_per_lf_pred), flow_cfs) |>
  ggplot() + geom_point(aes(x=wua_per_lf, y=wua_per_lf_pred, color=flow_cfs)) + 
  scale_y_log10() + scale_x_log10() +  coord_fixed() + geom_abline() + scale_color_viridis_c() + 
  ggtitle("Series interpolated scale-dependent model, linear regression")
```

### Use regression to predict FSA curve shape

* model flow-to-suitable area curve shape with exponential function
* use spatial predictors to predict the value of this exponent

```{r}
exponents <- 
  td |> 
  group_by(comid) |>
  group_modify(~ broom::tidy(lm(log(wua_per_lf) ~ 0 + log(flow_cfs), data = .x))) 
                                    
td_exp <- 
  flowlines |> st_drop_geometry() |>
  left_join(readRDS("../data/flowline_attributes.Rds"), by=join_by("comid"), relationship="one-to-one") |>
  inner_join(exponents |> select(comid, fsa_exponent=estimate), by=join_by(comid)) |>
  select(comid, fsa_exponent,
       # predictors of flow (as would be found in a regional regression)
       slope, da_area_sq_km, da_elev_mean, da_ppt_mean_mm, 
       # baseflow and peakflow statistics
       nf_bfl_dry_cfs, nf_bfl_wet_cfs, erom_q_ma_cfs, peak_q2_cfs, peak_q5_cfs, peak_q10_cfs,
       # flow and channel characteristics, hydrologically predicted
       bf_depth_m, bf_w_d_ratio, erom_v_ma_fps,
       # misc characteristics of the catchment
       da_avg_slope, da_k_erodibility, mean_ndvi,
       # misc characteristics of the locality
       loc_bfi, loc_pct_clay, loc_pct_sand, loc_permeability, loc_bedrock_depth, loc_ppt_mean_mm,
       # channel confinement characteristics
       mtpi30_min, #vb_width_transect, vb_bf_w_ratio, 
       # channel sinuosity
       sinuosity,
       )

```



