---
title: "Terrain-Derived Predictors"
author: "[Skyler Lewis](mailto:slewis@flowwest.com)"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
    toc_depth: 3
    number_sections: false
    math_method:
      engine: webtex
      url: https://latex.codecogs.com/svg.image?
  html_document:
    toc: true
    toc_depth: 3
    number_sections: false
    math_method:
      engine: webtex
      url: https://latex.codecogs.com/svg.image?
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(sf)
library(stars)

theme_set(theme_minimal())

#project_crs <- "EPSG:3310" # NAD83 California Teale Albers
#project_crs <- "EPSG:5070" # NAD83 CONUS Albers
project_crs <- "ESRI:102039" # NAD83 CONUS Albers USGS Version

library(googledrive)
drive_file_by_id <- function(id=character(), dir="temp", vsizip=FALSE) {
  d <- googledrive::as_dribble(googledrive::as_id(id))
  p <- file.path(dir, d$name)
  if(!file.exists(p)){
    googledrive::drive_download(file = d, path = p)
  }
  if(vsizip){
    return(file.path("/vsizip",p))
  } else {
    return(p)
  }
}
```

```{r imports}
selected_huc_8 <- c("18020107", "18020125")

flowlines <- readRDS("../data/flowline_geometries.Rds") |>
  left_join(readRDS("../data/flowline_attributes.Rds")) |>
#  filter(huc_8 %in% selected_huc_8) |>
  st_transform(project_crs) |>
  st_zm()

catchments <- readRDS("../data/catchments.Rds")
# for now, this is just the catchments exported for the selected hucs
aoi <- catchments |> summarize() |> st_union()
```

## Load DEM

### Data Source: 10m NHDPlusHR DEM

This input has already been clipped to the AOI, converted from cm to m, and crs redefined

Future improvement: script should pull straight from the source geotiff and apply these changes programatically

```{r import-dem, eval=FALSE}
dem <- read_stars("nhdplushr/dem_nhdplushr_yuba_meters_v2.tif") |>
  rename(elev = dem_nhdplushr_yuba_meters_v2.tif)

ggplot() + geom_stars(data=dem) + coord_fixed()
```

Example calculating slope (can use `stars` or `terra`)

```{r calc-slope, eval=FALSE}
slope <- dem |> starsExtra::slope()

#slope <- dem |> terra::rast() |> terra::slope() |> st_as_stars()

ggplot() + geom_stars(data=slope) + coord_fixed()
```

### Data Source: 30m NHDPlusV2 DEM

```{r import-dem30}
dir.create("temp/NEDSnapshot18b", recursive = TRUE)
drive_file_by_id("1mBCDj4vd8DvJecm99FF6_kX5YKfg78hr") |>
  archive::archive_extract(dir = "temp/NEDSnapshot18b")

dem30 <- stars::read_stars("temp/NEDSnapshot18b/elev_cm")

dem_crs <- st_crs(dem30)

dem30 |> mutate(elev_m = elev_cm/100) |> select(elev_m) |> plot()
```

## Channel Confinement via mTPI

mTPI as a measure of topographic confinement

Souce: Theobald DM, Harrison-Atlas D, Monahan WB, Albano CM. 2015. Ecologically-relevant maps of landforms and physiographic diversity for climate adaptation planning. PLOS ONE. DOI: 10.1371/journal.pone.0143619

```{r calc-mtpi, eval=FALSE}
dem_rast <- terra::rast(dem) 
cell_size <- 10 # mean(terra::res(dem_rast))
tpi_90 <- dem_rast - terra::focal(dem_rast, w=90/cell_size, fun="mean")
tpi_270 <- dem_rast - terra::focal(dem_rast, w=270/cell_size, fun="mean")
tpi_810 <- dem_rast - terra::focal(dem_rast, w=810/cell_size, fun="mean")
#tpi_2430 <- dem_rast - terra::focal(dem_rast, w=2430/cell_size, fun="mean")
mtpi <- terra::app(c(tpi_90, tpi_270, tpi_810), mean)
mtpi |> st_as_stars() |> plot()
```

```{r intersect-mtpi, eval=FALSE}
flow_buffered <- terra::vect(flowlines |> filter(huc_8 %in% selected_huc_8)) |> terra::buffer(width=50)
vec_mtpi_min <- terra::zonal(mtpi, flow_buffered, "min") |> rename(mtpi_min = mean)

mtpi_comid <- flowlines |> filter(huc_8 %in% selected_huc_8) |>
  select(comid) |>
  mutate(vec_mtpi_min) |>
  st_drop_geometry() |> 
  select(comid, mtpi_min) |> 
  filter(!is.nan(mtpi_min))

flowlines |> filter(huc_8 %in% selected_huc_8) |> 
  inner_join(mtpi_comid) |>
  st_zm() |>
  ggplot() + 
  geom_sf(aes(color = mtpi_min), linewidth=1) + 
  scale_color_viridis_c(direction=-1)
```

Version converting to function and using the basin-wide 30m DEM

```{r calc-mtpi-func}
calc_mtpi <- function(dem, cell_size, neighborhoods=c(90,270,810,2430)) {
  is_stars <- "stars" %in% class(dem)
  if(is_stars){
    dem <- terra::rast(dem)
  }
  tpis <- terra::rast(lapply(neighborhoods, function(x) dem - terra::focal(dem, w=x/cell_size, fun="mean", na.rm=TRUE, na.policy="omit")))
  mtpi <- terra::app(tpis, mean)
  if(is_stars) {
    mtpi <- st_as_stars(mtpi)
  }
  return(mtpi)
}
```

```{r calc-mtpi-30m}
cell_size <- 30

if(!file.exists("../data/attr_mtpi.Rds")){
  flow_buffered <- terra::vect(flowlines) |> terra::buffer(width=cell_size/2)
  
  #aoi_buffer <- aoi |> st_buffer(2430/2) # big enough to not create edge effects
  #dem30_rast <- dem30 |> st_crop(aoi_buffer) |> terra::rast()
  dem30_rast <- dem30 |> terra::rast()
  
  dem30_rast_mtpi <- (calc_mtpi(dem30_rast, cell_size, c(90,270,810, 2430)) / 100) |> terra::app(round) |> terra::as.int() # convert to m, 16-bit integer
  dem30_rast_mtpi |> terra::writeRaster("temp/rast_mtpi.tif")
  #dem30_rast_mtpi |> st_as_stars() |> plot()
  
  vec_mtpi30_min <- terra::zonal(dem30_rast_mtpi, flow_buffered, "min") |> rename(mtpi30_min = lyr.1)
  
  flowlines_mtpi <- flowlines |> select(comid) |> mutate(vec_mtpi30_min) 
  #flowlines_mtpi |> st_zm() |> select(mtpi30_min) |> plot()
  flowlines_mtpi_attr <- flowlines_mtpi |> st_drop_geometry() |> filter(!is.nan(mtpi30_min))
  flowlines_mtpi_attr |> saveRDS("../data/attr_mtpi.Rds")
  
} else {
  flowlines_mtpi_attr <- readRDS("../data/attr_mtpi.Rds")
  flowlines_mtpi <- flowlines |> left_join(flowlines_mtpi_attr)
}

```

### Valley Bottom Width via Slope Cutoff Method

Method described in https://watermanagement.ucdavis.edu/download_file/view_inline/144

```{r vbw-functions}
valley_bottom <- function(catchment, flowline, dem) {
  dem |> 
    st_crop(catchment) |> 
    starsExtra::slope() |>
    mutate(valley_bottom = if_else(slope<units::set_units(atan(0.25)*180/pi,"degrees"), 1, NA)) |>
    select(valley_bottom) |>
    st_as_sf(merge=TRUE) |>
    st_filter(st_zm(flowline)) #|> st_as_sfc()
}

valley_bottom_by_comid <- function(x, dem) {
  c <- catchments |> filter(comid==x)
  f <- flowlines |> filter(comid==x)
  valley_bottom(c, f, dem)
}

valley_width_naive <- function(geom_valley_bottom) {
   geom_valley_bottom |> 
    terra::vect() |> 
    terra::width()
}

perpendicular_transect <- function(pair=st_multipoint(), length=numeric()) {
  x1 <- pair[1]
  x2 <- pair[2]
  y1 <- pair[3]
  y2 <- pair[4]
  midpoint_x <- (x1+x2)/2
  midpoint_y <- (y1+y2)/2
  orig_bearing <- atan2((y2-y1), (x2-x1))
  perp_bearing <- orig_bearing + pi/2
  radius <- length/2
  perp_x1 <- midpoint_x + radius*cos(perp_bearing)
  perp_y1 <- midpoint_y + radius*sin(perp_bearing)
  perp_x2 <- midpoint_x - radius*cos(perp_bearing)
  perp_y2 <- midpoint_y - radius*sin(perp_bearing)
  return(st_linestring(c(st_point(c(perp_x1,perp_y1)), 
                         st_point(c(perp_x2, perp_y2)))))
}

perpendicular_transects <- function(ls=st_linestring(), length=numeric()) {
  if ("sf" %in% class(ls)){
    ls <- st_as_sfc(ls, crs=st_crs(ls))
  }
  if ("sfc" %in% class(ls)){
    m <- t(st_zm(ls[[1]]))
  } 
  if ("sfg" %in% class(ls)){
    m <- t(ls)
  }
  point_pairs <- lapply(seq(1, length(m)-3, 2), function(i) c(st_point(c(m[i], m[i+1])), st_point(c(m[i+2], m[i+3]))))
  perp_lines <- lapply(point_pairs, function(pair) perpendicular_transect(pair, length))
  return(st_sfc(perp_lines, crs=st_crs(ls)))
}

valley_width_via_transects <- function(geom_valley_bottom, flowline, max_width) {
  transects <- perpendicular_transects(flowline, length=max_width)
  if("sfg" %in% class(geom_valley_bottom)) {
    geom_valley_bottom <- st_sfc(geom_valley_bottom, crs=st_crs(transects))
  }
  clipped <- st_intersection(transects, geom_valley_bottom)
  return(mean(st_length(clipped)))
}
```

```{r vbw-example1}
flowline <- flowlines |> filter(comid==8062593) |> st_zm()
catchment <- catchments |> filter(comid==8062593) |> st_transform(st_crs(flowline))
ggplot() + 
  #geom_stars(data=catchment_confinement(catchment, flowline, dem)) + 
  geom_sf(data=valley_bottom(catchment, flowline, dem)) +
  geom_sf(data=st_zm(flowline))
valley_width_naive(valley_bottom(catchment, flowline, dem))
valley_polygon <- valley_bottom(catchment, flowline, dem)
valley_width_via_transects(valley_bottom(catchment, flowline, dem), flowline, max_width=1000)
ggplot() + 
  geom_sf(data=catchment) +
  geom_sf(data=valley_polygon, fill="lightblue") +
  geom_sf(data=flowline, color="blue") +
  geom_sf(data=perpendicular_transects(flowline, length=1000) |> st_intersection(valley_polygon), color="red") 
```

```{r vbw-example2}
dem30_m <- st_as_stars(dem30_rast/100) 

flowline <- flowlines |> filter(comid==8062593) |> st_zm()
catchment <- catchments |> filter(comid==8062593) |> st_transform(st_crs(flowline))
ggplot() + 
  #geom_stars(data=catchment_confinement(catchment, flowline, dem)) + 
  geom_sf(data=valley_bottom(catchment, flowline, dem30_m)) +
  geom_sf(data=st_zm(flowline))
valley_width_naive(valley_bottom(catchment, flowline, dem30_m))
valley_polygon <- valley_bottom(catchment, flowline, dem30_m)
valley_width_via_transects(valley_bottom(catchment, flowline, dem30_m), flowline, max_width=1000)
ggplot() + 
  geom_sf(data=catchment) +
  geom_sf(data=valley_polygon, fill="lightblue") +
  geom_sf(data=flowline, color="blue") +
  geom_sf(data=perpendicular_transects(flowline, length=1000) |> st_intersection(valley_polygon), color="red") 
```

```{r vbw-batch-old}
#vb1 <- flowlines |> 
#  filter(huc_8 %in% selected_huc_8) |> 
#  #filter(comid==8062593) |> # comment out the filter to process all
#  mutate(flowline = st_set_crs(st_zm(geometry),project_crs)) |>
#  st_drop_geometry() |>
#  mutate(valley_bottom_geoms = st_as_sfc(bind_rows(map(comid, possibly(function(x) valley_bottom_by_comid(x, dem), otherwise=NA))), crs=project_crs)) |>
#  mutate(vb_width_simple = map_dbl(valley_bottom_geoms, possibly(function(x) valley_width_naive(x), otherwise=NA))) |> 
#  mutate(vb_width_transect = map2_dbl(valley_bottom_geoms, flowline, function(x, y) valley_width_via_transects(x, y, 1000)),
#         vb_width_transect = if_else(!is.nan(vb_width_transect),vb_width_transect,NA)) |> 
#  select(comid, flowline, valley_bottom_geoms, vb_width_simple, vb_width_transect) 
#
#vb1 |> saveRDS("../data/attr_vb1.Rds")
```

```{r vbw-batch}
vb1 <- 
  flowlines |> 
  filter(huc_8 %in% selected_huc_8) |> #|> head(10) |>
  #filter(comid==8062593 | (comid>8062580 & comid<8062599)) |> # comment out the filter to process all
  mutate(flowline = st_set_crs(st_zm(geometry),project_crs)) |>
  st_drop_geometry() |>
  #mutate(valley_bottom_geoms = st_as_sfc(bind_rows(map(comid, possibly(function(x) valley_bottom_by_comid(x, dem30_m), otherwise=NA))), crs=project_crs))# |>
  mutate(valley_bottom_geoms = st_sfc(map(comid, possibly(function(x) st_union(valley_bottom_by_comid(x, dem30_m)$geometry)[[1]], otherwise=st_polygon())), crs=dem_crs)) |>
  mutate(vb_width_simple = map_dbl(valley_bottom_geoms, possibly(function(x) valley_width_naive(x), otherwise=NA))) |> 
  mutate(vb_width_transect = map2_dbl(valley_bottom_geoms, flowline, function(x, y) valley_width_via_transects(x, y, 1000)),
         vb_width_transect = if_else(!is.nan(vb_width_transect),vb_width_transect,0)) |> 
  select(comid, flowline, valley_bottom_geoms, vb_width_simple, vb_width_transect) 

vb1 |> saveRDS("../data/attr_vb1.Rds")
```

```{r vbw-batch-plot}
vb1 |> 
  filter(!is.na(vb_width_transect)) |> 
  st_set_geometry("flowline") |> 
  ggplot() +
  geom_sf(data=vb1$valley_bottom_geoms, color=NA, fill="orange") + 
  geom_sf(aes(color=vb_width_transect)) + 
  scale_color_viridis_c(direction=-1, trans="log10")
```

### Valley Bottom Width via VBET

Extract flowlines to use in vbet

```{r vbet-flowline-export}
flowlines_for_vbet <- flowlines |>
  filter(gnis_name %in% c("Yuba River")) |>
  st_transform(st_crs(dem)) |>
  st_zm() |> 
  st_crop(st_bbox(dem)) |>
  smoothr::densify(5)
```

===

```{r}
knitr::knit_exit()
```

## NDVI

image server? https://gis.apfo.usda.gov/arcgis/rest/services/NAIP/USDA_CONUS_PRIME/ImageServer

2020 seamlines: https://www.fpacbc.usda.gov/geo/customer-service/naip-quarter-quad-shapefiles/index.html

2016 JP2s: https://www.sciencebase.gov/catalog/item/51355312e4b0e1603e4fed62

* example URL: "https://prd-tnm.s3.amazonaws.com/StagedProducts/NAIP/ca_2016/38121/m_3812136_ne_10_h_20160620_20161004.jp2"

* 2016 metadata: https://thor-f5.er.usgs.gov/ngtoc/metadata/waf/orthoimagery/naip/ca_2016/

* footprints generated from earth explorer: ca_naip_2016.zip

```{r ndvi-download-naip, message=FALSE, warning=FALSE}
#naip_seam <- read_sf("/vsizip/temp/ca_naip20seam.zip") |>
#  st_filter(st_transform(aoi, "EPSG:4326"))

#naip_qq |>
#  select(file_name) |>
#  mutate(jp2_name = paste0("https://prd-tnm.s3.amazonaws.com/StagedProducts/NAIP/ca_2016/",
#                           str_sub(apfoname, start=1, end=5),
#                           str_sub()))

#naip_2020_qq_footprints <- read_sf("/vsizip/naip/ca_naip20qq.zip") |>
#  st_filter(st_transform(aoi, "EPSG:4326")) |>
#  janitor::clean_names()

# Get spatial index and filenames of NAIP tiles

naip_2016_qq_filenames_raw <- 
  RCurl::getURL("https://thor-f5.er.usgs.gov/ngtoc/metadata/waf/orthoimagery/naip/ca_2016/", ftp.use.epsv = FALSE, dirlistonly = TRUE) |>
  rvest::read_html() |> 
  xml2::xml_find_all("//a") |> 
  xml2::xml_attrs("href") |>
  as.vector()

naip_2016_qq_filenames <- tibble(xml_filename = naip_2016_qq_filenames_raw[-(1:5)]) |> 
  unnest(xml_filename) |>
  mutate(xml_filename = str_replace(xml_filename, ".xml", "")) |>
  separate_wider_delim(xml_filename, "_", names = c("prefix", "apfoname", "quadrant", "zone", "suffix", "src_img_date", "ver_date")) |>
  mutate(naip_entit = str_to_upper(paste(prefix, apfoname, quadrant, zone, suffix, src_img_date, sep="_")),
         filename = paste(paste(prefix, apfoname, quadrant, zone, suffix, src_img_date, ver_date, sep="_"), "jp2", sep="."),
         download = paste("https://prd-tnm.s3.amazonaws.com/StagedProducts/NAIP/ca_2016", str_sub(apfoname, start=1, end=5), filename, sep="/"))

naip_2016_qq_footprints <- 
  read_sf("/vsizip/naip/ca_naip_2016.zip") |>
  janitor::clean_names() |>
  left_join(naip_2016_qq_filenames)

# function to download NAIP tiles

options(timeout = max(3600, getOption("timeout")))

download_naip_tiles <- function(aoi, dir="temp/naip/tiles") {
  dir.create(dir, recursive = TRUE)
  tile_urls <- 
    naip_2016_qq_footprints |>
    st_filter(st_transform(aoi, "EPSG:4326")) |>
    st_drop_geometry() |>
    select(naip_entit, download) |>
    deframe()
  local_files <- map_vec(
    names(tile_urls),
    possibly(function(x) {
      f = paste0(dir, "/", x , ".jp2")
      if(!file.exists(f)) download.file(url = tile_urls[x], destfile = f, mode = "wb", quiet=T)
      cat(x, " ")
      return(f)}, otherwise=NA))
  return(local_files)
}

# download NAIP tiles and mosaic
naip_tiles <- download_naip_tiles(aoi)
```

```{r ndvi-calc-veg-pct, message=FALSE, warning=FALSE}
buffers <-
  flowlines |> 
  st_zm() |>
  #  mutate(channel_buffer = map2(geometry, bf_width_m, function(g, d) st_buffer(g, dist = d/2))) |>
  filter(is.finite(bf_width_m)) |>
  mutate(channel_buffer = st_buffer(geometry, dist = bf_width_m/2)) |>
  st_drop_geometry() |>
  select(comid, channel_buffer) |>
  left_join(vb1 |> select(comid, valley_bottom_geoms)) |>
  mutate(valley_bottom = map2(channel_buffer, valley_bottom_geoms, 
                              function(cb, vb) st_union(cb, vb, by_feature = FALSE)) |> 
                         st_sfc(crs=st_crs(flowlines)) |>
                         st_cast("MULTIPOLYGON")) |>
  select(-valley_bottom_geoms)

### TERRA VERSION (try stars instead, too big to read into memory)
# naip_mosaic <- naip_tiles |> terra::vrt(filename="temp/naip/naip.vrt", overwrite=TRUE)
# naip_ndvi <- (naip_mosaic["naip_4"] - naip_mosaic["naip_1"]) / (naip_mosaic["naip_4"] + naip_mosaic["naip_1"])
# ndvi_threshold <- 0.25
# naip_ndvi_cls <- 
#   naip_ndvi |> terra::classify(matrix(c(-1,ndvi_threshold,0,
#                                          ndvi_threshold,1,1), ncol=3, byrow=T), 
#                                include.lowest=T)
# ndvi_pct <-
#   buffers |>
#   st_drop_geometry() |>
#   select(comid) |>
#   mutate(veg_pct_cb = terra::zonal(naip_ndvi_cls, buffers$channel_buffer, "mean") |> rename(veg_pct_cb = mean),
#          veg_pct_vb = terra::zonal(naip_ndvi_cls, buffers$valley_bottom, "mean") |> rename(veg_pct_vb = mean))
# ndvi_pct |> saveRDS("./data/attr_veg_pct.Rds")

### STARS VERSION: better because of lazy evaluation
ndvi_threshold <- 0.25 # above which a pixel is considered vegetation
naip_mosaic <- 
  st_mosaic(naip_tiles, file_ext=".vrt") |> 
  read_stars(proxy=TRUE) #along="band"

# define functions
rf_ndvi = function(x1, x2, x3, x4) (x4 - x1)/(x4 + x1 + 1)
rf_rveg = function(ndvi) ndvi >= ndvi_threshold
get_veg_cls <- function(geom) {
  rveg <- naip_mosaic |>
    st_crop(st_bbox(geom)) |>
    st_apply(c("x", "y"), rf_ndvi, keep=T) |>
    st_apply(c("x", "y"), rf_rveg, keep=T) 
}
get_veg_pct <- function(geom) {
  agg <- geom |> 
    get_veg_cls() |> 
    aggregate(geom, mean, na.rm=T)
  #starsExtra::extract2(st_as_sfc(geom), mean, na.rm = TRUE)
  return(agg[[1]])
}

# test function with a random flowline buffer
test_geom <- flowlines |> st_zm() |> head(1) |> st_buffer(dist=100) |> st_transform(st_crs(naip_mosaic)) 
geom |> get_veg_cls() |> plot()
geom |> get_veg_pct()

# # batch quick test
# flowlines |> 
#   st_zm() |> 
#   head(3) |> 
#   st_buffer(dist=100) |> 
#   st_transform(st_crs(naip_mosaic)) |>
#   mutate(veg_pct = map_dbl(geometry, function(x) get_veg_pct(st_sfc(x, crs=st_crs(naip_mosaic)))))

# batch actual
veg_pcts <- buffers |>
  #head(3) |> # commment out to do all
  mutate(channel_buffer = st_transform(channel_buffer, st_crs(naip_mosaic)),
         valley_bottom = st_transform(valley_bottom, st_crs(naip_mosaic))) |>
  mutate(veg_pct_cb = map_dbl(channel_buffer, function(x) get_veg_pct(st_sfc(x, crs=st_crs(naip_mosaic)))),
         veg_pct_vb = map_dbl(valley_bottom, function(x) get_veg_pct(st_sfc(x, crs=st_crs(naip_mosaic))))) |>
  select(comid, veg_pct_cb, veg_pct_vb)
veg_pcts |> saveRDS("./data/attr_veg_pct.Rds")
```

## Channel Widths

Potential approach:

* Collect some random sample channel widths at transects
* Fit a curve of measured channel width versus NHD+ bankfull width
* Transform the NHD+ bankfull width into a "better" estimate using the model

Stratified sample of flowlines with perpendicular transects for manually sampling channel width:

```{r}
# # approach with one point every 100 ft
# sample_points_100ft <- 
#   flowlines |> 
#   filter(gnis_name %in% c("Yuba River", "South Yuba River", "Middle Yuba River", "North Yuba River")) |>
#   st_line_sample(density = 1 / 100) 
# 
# # approach with just the midpoint of each segment
# sample_points_midpt <- 
#   flowlines |>
#   filter(gnis_name %in% c("Yuba River", "South Yuba River", "Middle Yuba River", "North Yuba River")) |>
#   #st_line_sample(sample = c(0, 0.25, 0.5, 0.75, 1))
#   st_line_sample(sample = c(0.5))
# 
# sample_points_midpt |> ggplot() + geom_sf()

# approach with perpendicular transects on every segment

transects <- 
  flowlines |> 
  filter(huc_8 %in% selected_huc_8) |> 
  st_zm() |>
  mutate(geometry = map(geometry, function(x) st_union(perpendicular_transects(x, length=1000))[[1]]) |> st_sfc(crs=st_crs(flowlines))) 

#transects <- 
#  flowlines |> 
#  filter(huc_8 %in% selected_huc_8) |> #head(100) |>
#  st_zm() |>
#  mutate(transect = map(geometry, function(x) perpendicular_transects(x, length=1000))) |>
#  unnest(transect) |>
#  st_drop_geometry() 
# transects |> 
#   st_set_geometry(transects$transect)

set.seed(255) 

sample_transects <- 
  transects |> 
  filter(ftype_desc=="Stream/River" & !is.na(stream_order)) |>
  group_by(stream_order) |>
  slice_sample(n = 20) |>
  st_as_sf(crs=st_crs(flowlines)) |>
  st_cast("MULTILINESTRING")

sample_transects |>
  ggplot() + 
  geom_sf(data=st_zm(filter(flowlines, ftype_desc=="Stream/River" & !is.na(stream_order))),
          aes(color=as_factor(stream_order)), alpha=0.3) + 
  geom_sf(aes(color = as_factor(stream_order))) + 
  theme(legend.position = "top")

sample_transects |> 
  select(comid, gnis_name, stream_order) |>
  st_write("out/sample_transects.shp", append=FALSE)

```












