---
title: "Statistical Modeling to Predict Flow-to-Suitable-Area Curves"
author: "[Skyler Lewis](mailto:slewis@flowwest.com)"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
    toc_depth: 3
    number_sections: false
    math_method:
      engine: webtex
      url: https://latex.codecogs.com/svg.image?
---

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(sf)
library(stars)
library(tidymodels)
library(broom.mixed) # tidy output of mixed model results
library(dotwhisker) # visualize regression results

knitr::opts_chunk$set(eval=TRUE, fig.width=6.5, fig.height=6.5, dpi=300)

theme_set(theme_minimal())

ihs <- trans_new("ihs", 
                 transform = function(x) asinh(x), 
                 inverse = function(y) sinh(y), 
                 breaks = function(i) scales::breaks_log(n=5, base=10)(pmax(i,0.01)), 
                 #minor_breaks = scales::minor_breaks_n(n = 0),
                 domain=c(0, Inf),
                 format = scales::label_comma())

oom_range <- function(from, to) {
  from_mag <- floor(log10(if (from > 0) from else 1))
  to_mag <- ceiling(log10(to))
  magnitudes <- seq(from_mag, to_mag + 1, 1)
  expanded <- as.vector(t(10^seq(magnitudes) %o% seq(1,9,1)))
  result <- expanded[which(expanded>=signif(from,1) & expanded<=signif(to,1))]
  return(if (from == 0) c(0, result) else result)
}

hex_color_blend <- function(x, y){
  hex_r <- as.character.hexmode((strtoi(substr(x,2,3), base=16L) + strtoi(substr(y,2,3), base=16L)) / 2)
  hex_g <- as.character.hexmode((strtoi(substr(x,4,5), base=16L) + strtoi(substr(y,4,5), base=16L)) / 2)
  hex_b <- as.character.hexmode((strtoi(substr(x,6,7), base=16L) + strtoi(substr(y,6,7), base=16L)) / 2)
  return(paste0("#", hex_r, hex_g, hex_b))
}

palette_dsmhabitat_comparison <- 
  c("habistat prediction" = "#ffae34", 
    "DSMhabitat instream" = "#6388b4",
    "DSMhabitat floodplain" = "#8cc2ca")

palette_hqt_gradient_class <- 
  c("Valley Lowland"  = "#6388b4", #hex_color_blend("#6388b4","#55ad89"),
    "Valley Foothill" = "#55ad89", #hex_color_blend("#55ad89","#c3bc3f"),
    "Bedrock"         = "#bb7693") #hex_color_blend("#c3bc3f","#ffae34"))
```

## Import and Preprocess Training Data

```{r import, message=FALSE, warning=FALSE}
# flowline geometries
flowlines <- readRDS(here::here("data-raw", "results", "flowline_geometries_proj.Rds"))
flowlines_gcs <- readRDS(here::here("data-raw", "results", "flowline_geometries.Rds"))

# predictor variables by ComID
flowline_attributes <- readRDS(here::here("data-raw", "results", "flowline_attributes.Rds")) |>
  mutate(da_scalar = (da_area_sq_km*247.1053815) * (da_ppt_mean_mm*0.0393701) / 1E6) |># 1 million acre-inches
  mutate(length_ft_nhd = reach_length_km * 1000 / 0.3048) 
# TODO: Clean up import of these attributes; ensure UTF-8 encoding
```

```{r fsa-import}
# response variables by ComID and Flow 
#flow_to_suitable_area <- readRDS(here::here("data-raw", "results", "fsa_combined.Rds"))
fsa_combined <- 
  bind_rows(.id = "dataset",
    # VECTOR SRH-2D MODELS
    "Lower Yuba River" = 
      readRDS(here::here("data-raw", "results", "fsa_yuba.Rds")) |> 
      select(-reach),
    "Stanislaus River" = 
      readRDS(here::here("data-raw", "results", "fsa_stan.Rds")),
    # RASTER HEC-RAS 2D MODELS
    "Deer Creek" = 
      readRDS(here::here("data-raw", "results", "fsa_deer.Rds")),
    "Tuolumne River (Basso-La Grange)" = 
      readRDS(here::here("data-raw", "results", "fsa_basso.Rds")),
    ) |>
  rename(area_tot_ft2 = area_tot, area_wua_ft2 = area_wua, hsi_frac = area_pct)

fsa_combined |> saveRDS(here::here("data-raw", "results", "fsa_combined.Rds"))
```

```{r import-step2, fig.width=4, fig.height=6.5, dpi=300}
# Limit to a specified flow range and interpolate gaps
interp_flows <- seq(300,15000,100)

#interp_flows <- seq(300,40000,100)

# Use this alternative to bring in flows alll the way up to 85000 cfs
#interp_flows <- c(seq(300,950,50), seq(1000,9500,500), seq(10000,85000,5000))

flow_to_suitable_area <- 
  fsa_combined |>
  group_by(dataset, comid, length_ft) |>
  complete(flow_cfs = interp_flows) |>
  arrange(dataset, comid, flow_cfs) |>
  mutate(across(c(area_tot_ft2, area_wua_ft2, hsi_frac, ind_per_lf, wua_per_lf), 
                function(var) zoo::na.approx(var, x = flow_cfs, na.rm=F))) |>
  filter(flow_cfs %in% interp_flows) |>
  filter(!is.na(hsi_frac)) |>
  ungroup()

train_data <- flowlines |> st_drop_geometry() |>
  left_join(flowline_attributes, by=join_by("comid"), relationship="one-to-one") |>
  inner_join(flow_to_suitable_area, by=join_by("comid"), relationship="one-to-many") |>
  mutate(case_wt = hardhat::importance_weights(length_ft)) |>
  #filter(hqt_gradient_class != "Valley Lowland") |>
  # # eliminate reaches with near zero mean annual flow
  # filter(erom_q_ma_cfs>10) |>
  glimpse()

train_flowlines <- flowlines_gcs |>
  filter(comid %in% unique(flow_to_suitable_area$comid)) |>
  left_join(flowline_attributes |> select(comid, gnis_name, hqt_gradient_class), 
            by=join_by("comid"), 
            relationship="one-to-one") 

training_reach_labels <- train_flowlines |>
  group_by(gnis_name) |>
  summarize() |>
  st_centroid() |>
  filter(!is.na(gnis_name)) %>%
  mutate(coord_x = st_coordinates(geometry)[,1],
         coord_y = st_coordinates(geometry)[,2])

hqt_poly <- readRDS(here::here("data-raw", "source", "hqt", "hqt_valley_lowland.Rds"))
central_valley <- st_read(file.path("/vsizip",here::here("data-raw", "source", "hqt", "Alluvial_Bnd.shp.zip"))) |> st_transform("+proj=longlat +datum=NAD83")

ggplot() + 
  geom_sf(data=central_valley, aes(fill="Valley Foothill"), color=NA) +
  geom_sf(data=hqt_poly, aes(fill="Valley Lowland"), color=NA) +
  geom_sf(data = flowlines_gcs |> 
            inner_join(flowline_attributes |> 
                         filter(da_area_sq_km>100) |>
                         filter(substr(reachcode,1, 4) %in% c("1802", "1803", "1804"))), color="black", alpha=0.1, linewidth=0.5) +
  geom_sf(data = train_flowlines, aes(color = hqt_gradient_class), linewidth=1) + 
  geom_text(data = training_reach_labels, aes(x=coord_x, y=coord_y, label=gnis_name), vjust=-1, size=9/.pt) + 
  theme(legend.title=element_blank(), legend.position = "top", axis.title = element_blank(), legend.box="vertical") +
  scale_color_manual(name="Gradient Class", 
                     values = palette_hqt_gradient_class) +
  scale_fill_manual(name="Gradient Class",
                    values = c("Valley Foothill" = hex_color_blend("#55ad89","#ffffff"), 
                               "Valley Lowland" = hex_color_blend("#6388b4", "#ffffff"))) 
```

### Exploration and PCA

```{r pca}
df <- flowline_attributes |>
  select(
         # predictors of flow (as would be found in a regional regression)
         slope, da_area_sq_km, da_elev_mean, da_ppt_mean_mm, 
         # baseflow and peakflow statistics
         nf_bfl_dry_cfs, nf_bfl_wet_cfs, erom_q_ma_cfs, peak_q2_cfs, peak_q5_cfs, peak_q10_cfs,
         # flow and channel characteristics, hydrologically predicted
         bf_depth_m, bf_w_d_ratio, erom_v_ma_fps,
         # misc characteristics of the catchment
         da_avg_slope, da_k_erodibility, mean_ndvi,
         # misc characteristics of the locality
         loc_bfi, loc_pct_clay, loc_pct_sand, loc_permeability, loc_bedrock_depth, loc_ppt_mean_mm,
         # channel confinement characteristics
         mtpi30_min, vb_width_transect, vb_bf_w_ratio, 
         # channel sinuosity
         sinuosity,
         # aquatic
         # bio_aq_rank_sw, species, species_fish
         ) |> 
  na.omit() |> 
  scale() 

df |> cor() |> corrr::rplot() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
# uniqueness: 
# df |> cor() |> abs() |> rowMeans() |> rank() 
pca <- princomp(df) 
#summary(pca)

factoextra::fviz_eig(pca, addlabels = TRUE)
factoextra::fviz_cos2(pca, choice = "var", axes = 1:2)
factoextra::fviz_pca_var(pca, axes=c(1,2), col.var = "cos2", repel = TRUE)
factoextra::fviz_pca_var(pca, axes=c(1,3), col.var = "cos2", repel = TRUE)
factoextra::fviz_pca_var(pca, axes=c(2,3), col.var = "cos2", repel = TRUE)

# flowlines |> GGally::ggpairs()
```

## Training Data Setup

```{r td}
# pick variable to use for normalization
# norm_var <- sym("erom_q_ma_cfs")
#norm_var <- sym("da_area_sq_km")
norm_var <- sym("da_scalar")

filter_variable_ranges <- function(data) {
  data |>
    filter(da_area_sq_km > 100) # 1000 may have better model performance, 100 encompasses all streams we want to predict
}

td <- train_data |>
  mutate(flow_norm_cfs = flow_cfs / !!norm_var) |> # flow as a percent of mean annual flow
  mutate(wua_per_lf = area_wua_ft2 / length_ft,
         #log_wua_per_lf = log(wua_per_lf), # transect-wise habitat area per linear foot
         ihs_wua_per_lf = asinh(wua_per_lf), # inverse hyperbolic sine as alternative to log that can handle zeros
         wua_per_lf_norm = wua_per_lf / !!norm_var,
         #log_wua_per_lf_norm = log(wua_per_lf_norm), # transect-wise habitat area per linear foot
         ihs_wua_per_lf_norm = asinh(wua_per_lf_norm), # transect-wise habitat area per linear foot
         tot_area_per_lf = area_tot_ft2 / length_ft,
         #log_tot_area_per_lf = log(tot_area_per_lf),
         ihs_tot_area_per_lf = asinh(tot_area_per_lf)
         ) |> 
  transmute(dataset, comid, length_ft, case_wt,
         # suitable habitat area normalized by reach length
         hsi_frac, wua_per_lf, ihs_wua_per_lf, wua_per_lf_norm, ihs_wua_per_lf_norm, tot_area_per_lf, ihs_tot_area_per_lf,
         # flow cfs normalized by mean annual flow
         flow_cfs, flow_norm_cfs,
         # predictors of flow (as would be found in a regional regression)
         slope, da_area_sq_km, da_elev_mean, da_ppt_mean_mm, 
         # baseflow and peakflow statistics
         # nf_bfl_dry_cfs, nf_bfl_wet_cfs, erom_q_ma_cfs, peak_q2_cfs, peak_q5_cfs, peak_q10_cfs,
         # flow and channel characteristics, hydrologically predicted
         bf_depth_m, bf_w_d_ratio, # erom_v_ma_fps,
         # misc characteristics of the catchment
         da_avg_slope, da_k_erodibility, mean_ndvi,
         # misc characteristics of the locality
         loc_bfi, loc_pct_clay, loc_pct_sand, loc_permeability, loc_bedrock_depth, loc_ppt_mean_mm,
         # channel confinement characteristics
         mtpi30_min, vb_width_transect, vb_bf_w_ratio, frac_leveed_longitudinal,
         # channel sinuosity
         sinuosity,
         # aquatic
         # bio_aq_rank_sw, species, species_fish,
         # fixed effects
         hqt_gradient_class=droplevels(hqt_gradient_class), hyd_cls=droplevels(hyd_cls),
         # scalar for normalizing
         da_scalar
         ) |> 
  #mutate(nf_bfl_dry_cfs_norm = nf_bfl_dry_cfs/!!norm_var, 
  #       nf_bfl_wet_cfs_norm = nf_bfl_wet_cfs/!!norm_var) |>
  drop_na() |> 
  filter_variable_ranges() |>
  glimpse()
```

```{r td-summary, fig.height=6}
td |> ggplot() +
  geom_line(aes(x = flow_cfs, y = wua_per_lf, group=comid), color="lightgray") + 
  geom_smooth(aes(x = flow_cfs, y = wua_per_lf, color=hqt_gradient_class), se=F, method="gam") + 
  scale_x_log10(labels=scales::label_comma()) + scale_y_log10(labels=scales::label_comma()) +
  ggtitle("Summary of Training Data") + ylab("WUA (ft per linear ft)") + xlab("Flow (cfs)") + 
  facet_wrap(~dataset, ncol=1) + theme(legend.position = "top")
```

Calculate the ranges of the predictor variables

```{r td-ranges}
# td |> 
#   group_by(dataset) |>
#   select_if(is.numeric) |>
#   pivot_longer(cols = hsi_frac:last_col()) |>
#   group_by(dataset, name) |>
#   summarize(min = min(value), max=max(value))

domain <- td |> 
  select_if(is.numeric) |>
  pivot_longer(cols = hsi_frac:last_col()) |>
  group_by(name) |>
  summarize(min=min(value), max=max(value)) |>
  mutate(value = map2(min, max, function(x, y) list(min=x, max=y))) |>
  select(name, value) |> deframe()

within_domain <- function(x, name) {
  if (name %in% names(domain)) {
    res <- x >= domain[[name]][["min"]] & x <= domain[[name]][["max"]]
    message(paste(name, "matches", sum(res), "of", length(x), "rows"))
    return(res)
  }
  else {
    return(TRUE)
  }
}

filter_within_domain <- function(data, .cols = everything()) {
  out <- data |>
    filter((if_all({{.cols}}, function(x) within_domain(x, cur_column()))))
  message(paste("all conditions matched by", nrow(out), "of", nrow(data), "rows"))
  return(out)
}

# example usage
# pd |> nrow()
# pd |> filter_within_domain(c(flow_cfs, da_area_sq_km, da_elev_mean)) |> nrow()
```

```{r td-flow-predictor-curves, fig.width=6.5, fig.height=6.5, dpi=300}
td |> 
  select(wua_per_lf, slope:sinuosity) |>
  pivot_longer(cols = everything()) |>
  ggplot(aes(x = value)) + 
  geom_histogram(aes(y = after_stat(count / sum(count)))) +
  facet_wrap(~name, scales = "free_x") +
  scale_x_continuous(trans = ihs) + theme(panel.grid.minor = element_blank())

identify_outliers <- function(x, qlo = 0.1, qhi = 0.9) {
  lo <- quantile(x, qlo)
  hi <- quantile(x, qhi)
  return(x < lo | x > hi)
}
# td |> filter((if_all(slope:sinuosity, function(x) !identify_outliers(x, qlo=0.1, qhi=0.9)))) 
  
clip_outliers <- function(x, qlo = 0.1, qhi = 0.9) {
  lo <- quantile(x, qlo)
  hi <- quantile(x, qhi)
  clip_lo <- if_else(x > lo, x, lo)
  clip_hi <- if_else(clip_lo < hi, clip_lo, hi)
  return(clip_hi)
}
# td |> mutate(across(slope:sinuosity, function(x) clip_outliers(x, qlo=0.1, qhi=0.9)))
  
n_breaks <- 10
td |>
  select(comid, wua_per_lf, flow_cfs, slope:sinuosity) |>
  mutate(across(slope:sinuosity, function(x) cut(x, breaks=n_breaks, labels=seq(1,n_breaks)/n_breaks))) |>
  pivot_longer(cols = slope:sinuosity) |>
  ggplot(aes(x = flow_cfs, y = wua_per_lf)) +
  facet_wrap(~name) + geom_smooth(aes(color = value), se=F) +
  scale_color_viridis_d(name = "Percentile of predictor variable", option = "cividis", direction=-1) +
  theme(legend.position = "top") + guides(color=guide_legend(nrow=2, byrow=TRUE)) +
  scale_x_log10() + scale_y_log10() + theme(panel.grid.minor = element_blank())
```

Split into training and testing datasets

```{r td-split}
set.seed(47)
td_split <- group_initial_split(td |> mutate(strata=paste(hyd_cls, dataset)), strata=strata, group=comid)

# select some example reaches to use for validation
test_comids <- testing(td_split) |> group_by(dataset, hyd_cls, comid) |> summarize() |> group_by(dataset, hyd_cls) |> filter(row_number()<=11) |> pull(comid) 
```

## Model Training

### Scale Dependent Model: WUA-per-linear-ft versus flow

Scale-dependent model version predicts the WUA per linear foot, using absolute flow values as well as watershed scale parameters.

Straightforward but highly sensitive to the ranges of flow data available in training datasets.

```{r recipe-sd, message=FALSE}
sd_rec <- recipe(data=training(td_split), 
      formula = ihs_wua_per_lf + case_wt ~ 
        # flow (cfs)
        flow_cfs + 
        # predictors of flow (catchment area, elevation, and MAP) -- attributes for gradient and upstream drainage area are interrelated
        slope + da_area_sq_km + da_elev_mean + da_ppt_mean_mm +
        # baseflow/peak flow statistics
        # nf_bfl_dry_cfs + nf_bfl_wet_cfs + erom_q_ma_cfs + #log(peak_q2_cfs)  + log(peak_q5_cfs) + log(peak_q10_cfs) +
        # flow and channel characteristics, hydrologically predicted
        #erom_v_ma_fps + 
        bf_depth_m + bf_w_d_ratio + 
        # misc characteristics of the catchment
        da_k_erodibility + da_avg_slope + mean_ndvi +
        # misc characteristics of the locality
        loc_bfi + loc_pct_clay + loc_pct_sand + loc_permeability + loc_bedrock_depth + loc_ppt_mean_mm +
        # channel confinement characteristics
        mtpi30_min + vb_bf_w_ratio + frac_leveed_longitudinal + vb_width_transect +
        # channel sinuosity
        sinuosity #+ 
        # aquatic
        # bio_aq_rank_sw + species + species_fish + 
        # fixed effects
        # hqt_gradient_class + hyd_cls
        ) |>
 #step_log(all_numeric_predictors(), -mtpi30_min, -mean_ndvi, -vb_bf_w_ratio, -frac_leveed_longitudinal) |>
  step_mutate_at(all_numeric_predictors(), fn = asinh) |>
  step_interact(terms = ~ slope:da_area_sq_km) |>
  #step_dummy(hqt_gradient_class) |>
  #step_dummy(hyd_cls) |>
  step_interact(terms = ~flow_cfs:all_predictors()) |>
  step_naomit(all_predictors()) |>
  step_zv(all_predictors()) |>
  step_normalize(all_numeric_predictors()) 
```

#### Linear Regression

```{r lm-sd}
lm_spec <- linear_reg() 

lm_sd <- workflow() |>
  add_recipe(sd_rec) |>
  add_model(lm_spec) |>
  add_case_weights(case_wt) |>
  fit(data=training(td_split))

lm_sd |> glance()
lm_sd |> tidy()

lm_sd_res <- testing(td_split) |>
  transmute(comid, flow_cfs,
            # actuals
            ihs_wua_per_lf, 
            wua_per_lf = sinh(ihs_wua_per_lf),
            # predicted
            ihs_wua_per_lf_pred = predict(lm_sd, testing(td_split))[[".pred"]],
            wua_per_lf_pred=sinh(ihs_wua_per_lf_pred))

lm_sd_res |>
  arrange(comid, flow_cfs) |>
  ggplot() + geom_path(aes(x=wua_per_lf, y=wua_per_lf_pred, color=flow_cfs, group=comid), linewidth=2) + 
  #scale_y_log10(limits=c(1,NA)) + scale_x_log10(limits=c(1,NA)) + 
  scale_y_continuous(trans=ihs, limits=c(1,NA), labels=scales::label_comma()) + 
  scale_x_continuous(trans=ihs, limits=c(1,NA), labels=scales::label_comma()) +
  coord_fixed() + geom_abline() + scale_color_viridis_c(name="Flow (cfs)") + 
  ggtitle("Scale-dependent model, linear regression") +
  xlab("Suitable Habitat Area (ft2) per LF (Actual)") + ylab("Suitable Habitat Area (ft2) per LF (Predicted)")

lm_sd_res |>
  filter(flow_cfs %in% c(300,500,1000,1500,3000,5000,10000,15000)) |>
  ggplot() + geom_point(aes(x=wua_per_lf, y=wua_per_lf_pred, color=flow_cfs)) + 
  facet_wrap(~flow_cfs, ncol = 4) +
  #scale_y_log10(limits=c(1,NA)) + scale_x_log10(limits=c(1,NA)) + 
  scale_y_continuous(trans=ihs, limits=c(1,NA), labels=scales::label_comma()) + 
  scale_x_continuous(trans=ihs, limits=c(1,NA), labels=scales::label_comma()) +
  coord_fixed() + geom_abline() + scale_color_viridis_c(name="Flow (cfs)") + 
  ggtitle("Scale-dependent model, linear regression") +
  xlab("Suitable Habitat Area (ft2) per LF (Actual)") + ylab("Suitable Habitat Area (ft2) per LF (Predicted)")

lm_sd_res |> filter(comid %in% test_comids) |>
  ggplot(aes(x=flow_cfs, group=1)) + facet_wrap(~comid) + 
  geom_line(aes(y=wua_per_lf_pred, color="predicted", group=1)) + geom_line(aes(y=wua_per_lf, color="actual", group=1)) + 
  scale_y_continuous(trans = ihs, limits = c(0, NA)) +
  xlab("Flow (cfs)") + ylab("Suitable Habitat Area per LF Channel (ft)")  +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), panel.grid.minor=element_blank()) +
  scale_x_continuous(labels = scales::label_comma())

```

```{r lm-sd-coeffs, fig.width=6.5, fig.height=6.5, dpi=300}
lm_sd |>
  tidy() |>
  mutate(ixn_class = if_else(str_detect(term, "flow_cfs_x_"), "term_flow_cfs_x_", "term") |> 
                     factor(levels = c("term", "term_flow_cfs_x_"), labels = c("Constant Effect", "Interaction with Flow (cfs)")),
         term_root = str_replace(term, "flow_cfs_x_", ""),
         coeff_dir = if_else(estimate>0, "Positive", "Negative"),
         ci_lower = estimate - qt(0.975, df = glance(lm_sd)$df) * std.error,
         ci_upper = estimate + qt(0.975, df = glance(lm_sd)$df) * std.error,
         estimate_abs = abs(estimate),
         ci_lower_abs = ci_lower * if_else(coeff_dir=="Negative",-1,1),
         ci_upper_abs = ci_upper * if_else(coeff_dir=="Negative",-1,1),
         is_significant = sign(ci_lower)==sign(ci_upper)
         ) |>
  filter(term!="flow_cfs" & term!="(Intercept)") |>
  ggplot(aes(y = term_root, color = coeff_dir)) +
  facet_wrap(~ixn_class, ncol = 2, scales = "free_x") + 
  geom_errorbarh(aes(xmin = ci_lower_abs, xmax=ci_upper_abs)) + #pmax(0,ci_lower_abs), xmax = pmax(0,ci_upper_abs))) +
  geom_point(aes(x = estimate_abs, fill = is_significant), shape=21) +
  theme(legend.position="top",
        panel.grid.minor = element_blank(),
        panel.border = element_rect(fill = NA, color = "black"), 
        legend.direction = "vertical") +
  scale_y_discrete(limits=rev) + 
  scale_fill_manual(name = "Significant (95%)", values = c("FALSE" = "white", "TRUE" = "black")) +
  scale_color_manual(name = "Direction of Effect", values = c("Negative" = "goldenrod", "Positive" = "darkcyan")) +
  xlab("Coefficient Magnitude") + ylab("")
```

#### Random Forest Regression

```{r tuning1, eval=FALSE, include=FALSE}
tune_rf <- function(mtry, trees, min_n) {
  
  spec <- rand_forest(mode = "regression", mtry = mtry, trees = trees, min_n = min_n) |>
    set_engine("ranger", num.threads = 4)
  
  wf <- workflow() |>
    add_recipe(sd_rec) |>
    add_model(spec) |>
    fit(data=training(td_split))
  
  residuals <- sd_rec |> 
    prep(training(td_split)) |> 
    bake(td) |>
    bind_cols(td |> transmute(comid, 
                              wua_per_lf,
                              flow_cfs_orig = flow_cfs,
                              train_test = if_else(comid %in% training(td_split)$comid, "train", "test"))) |>
    drop_na() %>% # need to use magrittr pipe for this purpose
    mutate(ihs_wua_per_lf_pred = predict(rfr_sd$fit$fit, new_data=.)[[".pred"]]) |>
    transmute(comid, 
              wua_per_lf,
              train_test,
              flow_cfs = flow_cfs_orig, 
              ihs_wua_per_lf_pred,
              wua_per_lf_pred = sinh(ihs_wua_per_lf_pred),
              resid = wua_per_lf_pred - wua_per_lf,
              sqerr = resid^2)
  
  residuals |> 
    group_by(train_test) |>
    summarize(mres = mean(resid),
              rmse = sqrt(mean(sqerr))) |>
    pivot_wider(names_from = train_test, values_from = c(mres, rmse))

}

library(future) # parallel processing backend
library(future.apply) # parallelized versions of lapply, sapply, ...
library(furrr) # parallelized versions of purrr functions: map, map2, pmap,...
plan(multisession, workers = 8) #plan(sequential)

#tune_grid <- expand_grid(mtry = c(5, 10, 20), 
#                         min_n = c(5, 10, 20), 
#                         trees = 2^seq(1, 9, 1))
#tune_grid <- grid_regular(finalize(mtry(), sd_rec |> prep(td) |> bake(td)),
#             trees(), min_n(), levels=5)
tune_res <- tune_grid |>
  mutate(result = future_pmap(list(mtry, trees, min_n), tune_rf)) |>
  unnest_wider(result)

tune_res |>
  ggplot(aes(x = trees)) + 
  geom_smooth(aes(y = rmse_train, linetype = "rmse", color = "train")) + 
  geom_smooth(aes(y = rmse_test, linetype = "rmse", color = "test")) + 
  geom_smooth(aes(y = mres_train, linetype = "mean resid", color = "train")) + 
  geom_smooth(aes(y = mres_test, linetype = "mean resid", color = "test")) + 
  scale_y_continuous(name = "RMSE", sec.axis = sec_axis(trans = ~., name = "Mean Residual")) +
  xlab("Number of Trees")

```

```{r tuning2, eval=FALSE, include=FALSE}

folds <- vfold_cv(training(td_split), v = 10, strata = dataset)
val_metrics <- metric_set(rmse, rsq, ccc)

cvtune_spec <- rand_forest(mode = "regression", mtry = tune(), trees = tune(), min_n = tune()) |>
   set_engine("ranger", num.threads = 4)
 
cvtune_res <- workflow() |>
   add_recipe(sd_rec) |>
   add_model(cvtune_spec) |>
   tune_grid(resamples = folds,
             grid = tune_grid,
             metrics = val_metrics)
 
cvtune_res |>
   collect_metrics() |>
   filter(.metric == "rmse") |>
   ggplot(aes(x = trees, y = mean, color = as_factor(min_n), linetype = as_factor(mtry), group = paste(min_n, mtry))) +
   #geom_errorbar(aes(ymin = mean - std_err,
   #                  ymax = mean + std_err),
   #              width = 0.05) +
   geom_line() +
   #facet_wrap(. ~ mtry, ncol = 1) +
   scale_x_log10()

```

```{r rfr-sd, message=FALSE}
mtry <- floor(length(sd_rec$var_info$variable[which(sd_rec$var_info$role=="predictor")])/3)
rfr_spec <- rand_forest(mode = "regression", trees = 2^8, mtry = mtry, min_n = 10) |> #, min_n = 10) |>
  #set_engine("ranger", num.threads = 4)#, sample.fraction = .8, importance = "impurity")
  set_engine("ranger", num.threads = 4)#, sample.fraction = .8, importance = "impurity")

rfr_sd <- workflow() |>
  add_recipe(sd_rec) |>
  add_model(rfr_spec) |>
  add_case_weights(case_wt) |>
  fit(data=training(td_split))

rfr_sd$fit$fit |> print()

rfr_sd_res <- testing(td_split) |>
  transmute(comid, flow_cfs,
            # actuals
            ihs_wua_per_lf, 
            wua_per_lf = sinh(ihs_wua_per_lf),
            # predicted
            ihs_wua_per_lf_pred = predict(rfr_sd, testing(td_split))[[".pred"]],
            wua_per_lf_pred=sinh(ihs_wua_per_lf_pred))

rfr_sd_res |>
  arrange(comid, flow_cfs) |>
  ggplot() + geom_path(aes(x=wua_per_lf, y=wua_per_lf_pred, color=flow_cfs, group=comid), linewidth=2) + 
  #scale_y_log10(limits=c(1,NA)) + scale_x_log10(limits=c(1,NA)) + 
  scale_y_continuous(trans=ihs, limits=c(1,NA), labels=scales::label_comma()) + 
  scale_x_continuous(trans=ihs, limits=c(1,NA), labels=scales::label_comma()) +
  coord_fixed() + geom_abline() + scale_color_viridis_c(name="Flow (cfs)")   + 
  ggtitle("Scale-dependent model, random forest regression") +
  xlab("Suitable Habitat Area (ft2) per LF (Actual)") + ylab("Suitable Habitat Area (ft2) per LF (Predicted)")

rfr_sd_res |>
  filter(flow_cfs %in% c(300,500,1000,1500,3000,5000,10000,15000)) |>
  ggplot() + geom_point(aes(x=wua_per_lf, y=wua_per_lf_pred, color=flow_cfs)) + 
  facet_wrap(~flow_cfs, ncol = 4) +
  #scale_y_log10(limits=c(1,NA)) + scale_x_log10(limits=c(1,NA)) + 
  scale_y_continuous(trans=ihs, limits=c(1,NA), labels=scales::label_comma()) + 
  scale_x_continuous(trans=ihs, limits=c(1,NA), labels=scales::label_comma()) +
  coord_fixed() + geom_abline() + scale_color_viridis_c(name="Flow (cfs)") + 
  ggtitle("Scale-dependent model, random forest regression") +
  xlab("Suitable Habitat Area (ft2) per LF (Actual)") + ylab("Suitable Habitat Area (ft2) per LF (Predicted)")

rfr_sd_res |> filter(comid %in% test_comids) |>
  ggplot(aes(x=flow_cfs, group=1)) + facet_wrap(~comid) + 
  geom_line(aes(y=wua_per_lf_pred, color="predicted", group=1)) + geom_line(aes(y=wua_per_lf, color="actual", group=1)) + 
  scale_y_continuous(trans = ihs, limits = c(0, NA)) + 
  xlab("Flow (cfs)") + ylab("Suitable Habitat Area per LF Channel (ft)")  +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), panel.grid.minor=element_blank()) +
  scale_x_continuous(labels = scales::label_comma())

residuals <- sd_rec |> 
  prep(training(td_split)) |> 
  bake(td) |>
  bind_cols(td |> transmute(comid, 
                            wua_per_lf,
                            flow_cfs_orig = flow_cfs,
                            train_test = if_else(comid %in% training(td_split)$comid, "Train", "Test"))) |>
  drop_na() %>% # need to use magrittr pipe for this purpose
  mutate(ihs_wua_per_lf_pred = predict(rfr_sd$fit$fit, new_data=.)[[".pred"]]) |>
  transmute(comid, 
            wua_per_lf,
            train_test,
            flow_cfs = flow_cfs_orig, 
            ihs_wua_per_lf_pred,
            wua_per_lf_pred = sinh(ihs_wua_per_lf_pred),
            resid = wua_per_lf_pred - wua_per_lf)
residuals |> 
  group_by(train_test) |>
  summarize(mean(resid)) |> 
  deframe()
residuals |>
  ggplot(aes(x = flow_cfs)) + 
  geom_line(aes(y = wua_per_lf_pred - wua_per_lf, group = comid, color = train_test)) +
  scale_x_log10()

```

### Scale Dependent Model 2: TOTAL Inundated Area per linear ft versus flow:

To generate the inundation area estimates that are needed to interpret the scale-independent %HSI model outputs

```{r recipe-sd2, message=FALSE}
sd2_rec <- recipe(data=training(td_split), 
      formula = ihs_tot_area_per_lf + case_wt ~ 
        # flow (cfs)
        flow_cfs + 
        # predictors of flow (catchment area, elevation, and MAP) -- attributes for gradient and upstream drainage area are interrelated
        slope + da_area_sq_km + da_elev_mean + da_ppt_mean_mm +
        # baseflow/peak flow statistics
        #nf_bfl_dry_cfs + nf_bfl_wet_cfs + erom_q_ma_cfs + #log(peak_q2_cfs)  + log(peak_q5_cfs) + log(peak_q10_cfs) +
        # flow and channel characteristics, hydrologically predicted
        #erom_v_ma_fps + 
        bf_depth_m + bf_w_d_ratio + 
        # misc characteristics of the catchment
        da_k_erodibility + da_avg_slope + 
        # misc characteristics of the locality
        loc_bfi + loc_pct_clay + loc_pct_sand + loc_permeability + loc_bedrock_depth + loc_ppt_mean_mm +
        # channel confinement characteristics
        mtpi30_min + vb_bf_w_ratio + frac_leveed_longitudinal + vb_width_transect +
        # channel sinuosity
        sinuosity #+ 
        # fixed effects
        # hqt_gradient_class + hyd_cls
        ) |>
  #step_log(all_numeric_predictors(), -mtpi30_min, -vb_bf_w_ratio, -frac_leveed_longitudinal) |>
  step_mutate_at(all_numeric_predictors(), fn = asinh) |>
  step_interact(terms = ~ slope:da_area_sq_km) |>
  # step_dummy(hqt_gradient_class) |>
  # step_dummy(hyd_cls) |>
  #step_spline_nonnegative(flow_cfs) |>
  step_interact(terms = ~flow_cfs:all_predictors()) |>
  step_naomit(all_predictors()) |>
  step_zv(all_predictors()) |>
  step_normalize(all_numeric_predictors())
```

#### Linear Regression

```{r lm-sd2}
lm_spec <- linear_reg() 

lm_sd2 <- workflow() |>
  add_recipe(sd2_rec) |>
  add_model(lm_spec) |>
  add_case_weights(case_wt) |>
  fit(data=training(td_split))

lm_sd2 |> glance()
lm_sd2 |> tidy()

lm_sd2_res <- testing(td_split) |>
  transmute(comid, flow_cfs,
            # actuals
            ihs_tot_area_per_lf, 
            tot_area_per_lf = sinh(ihs_tot_area_per_lf),
            # predicted
            ihs_tot_area_per_lf_pred = predict(lm_sd2, testing(td_split))[[".pred"]],
            tot_area_per_lf_pred=sinh(ihs_tot_area_per_lf_pred))

lm_sd2_res |>
  arrange(comid, flow_cfs) |>
  ggplot() + geom_path(aes(x=tot_area_per_lf, y=tot_area_per_lf_pred, color=flow_cfs, group=comid), linewidth=2) + 
  scale_y_log10() + scale_x_log10() +  
  coord_fixed() + geom_abline() + scale_color_viridis_c(name="Flow (cfs)") + 
  ggtitle("Scale-dependent model, linear regression") +
  xlab("Total Inundated Area per LF (Actual)") + ylab("Total Inundated Area per LF (Predicted)")

lm_sd2_res |>
  filter(flow_cfs %in% c(300,500,1000,1500,3000,5000,10000,15000)) |>
  ggplot() + geom_point(aes(x=tot_area_per_lf, y=tot_area_per_lf_pred, color=flow_cfs)) + 
  facet_wrap(~flow_cfs, ncol = 4) +
  scale_y_log10() + scale_x_log10() +  coord_fixed() + geom_abline() + scale_color_viridis_c(name="Flow (cfs)") + 
  ggtitle("Scale-dependent model, linear regression") +
  xlab("Total Inundated Area per LF (Actual)") + ylab("Total Inundated Area per LF (Predicted)")

lm_sd2_res |> filter(comid %in% test_comids) |>
  ggplot(aes(x=flow_cfs, group=1)) + facet_wrap(~comid) + 
  geom_line(aes(y=tot_area_per_lf_pred, color="predicted", group=1)) + geom_line(aes(y=tot_area_per_lf, color="actual", group=1)) + 
  scale_y_continuous(trans = ihs) + 
  xlab("Flow (cfs)") + ylab("Total Inundated Area per LF Channel (ft)")  +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), panel.grid.minor=element_blank()) +
  scale_x_continuous(labels = scales::label_comma())

```

```{r lm-sd2-coeffs, fig.width=6.5, fig.height=6.5, dpi=300}
lm_sd2 |>
  tidy() |>
  mutate(ixn_class = if_else(str_detect(term, "flow_cfs_x_"), "term_flow_cfs_x_", "term") |> 
                     factor(levels = c("term", "term_flow_cfs_x_"), labels = c("Constant Effect", "Interaction with Flow (cfs)")),
         term_root = str_replace(term, "flow_cfs_x_", ""),
         coeff_dir = if_else(estimate>0, "Positive", "Negative"),
         ci_lower = estimate - qt(0.975, df = glance(lm_sd)$df) * std.error,
         ci_upper = estimate + qt(0.975, df = glance(lm_sd)$df) * std.error,
         estimate_abs = abs(estimate),
         ci_lower_abs = ci_lower * if_else(coeff_dir=="Negative",-1,1),
         ci_upper_abs = ci_upper * if_else(coeff_dir=="Negative",-1,1),
         is_significant = sign(ci_lower)==sign(ci_upper)
         ) |>
  filter(term!="flow_cfs" & term!="(Intercept)") |>
  ggplot(aes(y = term_root, color = coeff_dir)) +
  facet_wrap(~ixn_class, ncol = 2, scales = "free_x") + 
  geom_errorbarh(aes(xmin = ci_lower_abs, xmax=ci_upper_abs)) + #pmax(0,ci_lower_abs), xmax = pmax(0,ci_upper_abs))) +
  geom_point(aes(x = estimate_abs, fill = is_significant), shape=21) +
  theme(legend.position="top",
        panel.grid.minor = element_blank(),
        panel.border = element_rect(fill = NA, color = "black"), 
        legend.direction = "vertical") +
  scale_y_discrete(limits=rev) + 
  scale_fill_manual(name = "Significant (95%)", values = c("FALSE" = "white", "TRUE" = "black")) +
  scale_color_manual(name = "Direction of Effect", values = c("Negative" = "goldenrod", "Positive" = "darkcyan")) +
  xlab("Coefficient Magnitude") + ylab("")
```

#### Random Forest Regression

```{r rfr-sd2, message=FALSE}
mtry <- floor(length(sd2_rec$var_info$variable[which(sd2_rec$var_info$role=="predictor")])/3)
rfr_spec <- rand_forest(mode = "regression", trees = 2^8, mtry = mtry, min_n = 10) |> #, min_n = 10) |>
    set_engine("ranger", num.threads = 4)

rfr_sd2 <- workflow() |>
  add_recipe(sd2_rec) |>
  add_model(rfr_spec) |>
  add_case_weights(case_wt) |>
  fit(data=training(td_split))

rfr_sd2$fit$fit |> print()

rfr_sd2_res <- testing(td_split) |>
  transmute(comid, flow_cfs,
            # actuals
            ihs_tot_area_per_lf, 
            tot_area_per_lf = sinh(ihs_tot_area_per_lf),
            # predicted
            ihs_tot_area_per_lf_pred = predict(rfr_sd2, testing(td_split))[[".pred"]],
            tot_area_per_lf_pred=sinh(ihs_tot_area_per_lf_pred))

rfr_sd2_res |>
  arrange(comid, flow_cfs) |>
  ggplot() + geom_path(aes(x=tot_area_per_lf, y=tot_area_per_lf_pred, color=flow_cfs, group=comid), linewidth=2) + 
  scale_y_log10() + scale_x_log10() +  coord_fixed() + geom_abline() + scale_color_viridis_c(name="Flow (cfs)")   + 
  ggtitle("Scale-dependent model, random forest regression") +
  xlab("Total Inundated Area per LF (Actual)") + ylab("Total Inundated Area per LF (Predicted)")

rfr_sd2_res |>
  filter(flow_cfs %in% c(300,500,1000,1500,3000,5000,10000,15000)) |>
  ggplot() + geom_point(aes(x=tot_area_per_lf, y=tot_area_per_lf_pred, color=flow_cfs)) + 
  facet_wrap(~flow_cfs, ncol = 4) +
  scale_y_log10() + scale_x_log10() +  coord_fixed() + geom_abline() + scale_color_viridis_c(name="Flow (cfs)") + 
  ggtitle("Scale-dependent model, random forest regression") +
  xlab("Total Inundated Area per LF (Actual)") + ylab("Total Inundated Area per LF (Predicted)")

rfr_sd2_res |> filter(comid %in% test_comids) |>
  ggplot(aes(x=flow_cfs, group=1)) + facet_wrap(~comid) + 
  geom_line(aes(y=tot_area_per_lf_pred, color="predicted", group=1)) + geom_line(aes(y=tot_area_per_lf, color="actual", group=1)) + 
  scale_y_continuous(trans = ihs) +
  xlab("Flow (cfs)") + ylab("Total Inundated Area per LF Channel (ft)")  +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), panel.grid.minor=element_blank()) +
  scale_x_continuous(labels = scales::label_comma())
```

### Scale Independent Model: %HSI versus flow (normalized by MAF)

Scale-independent model version predicts the percent suitable habitat area using normalized flow, and without using any direct correlates of watershed size.

To actually apply these results requires an estimate of total inundated area, which can be then multiplied with the predicted %HSI.

Technically this should not use linear regression as it is a proportion between zero and one, it should more properly be using beta regression via the betareg package.

```{r recipe-si, message=FALSE}
si_rec <- recipe(data=training(td_split), 
      formula=hsi_frac + case_wt ~ 
        # flow as percent of mean annual flow
        flow_norm_cfs + 
        # channel characteristics: gradient and sinuosity
        slope + sinuosity + 
        # flow and channel characteristics, hydrologically predicted
        # erom_v_ma_fps + 
        bf_depth_m + bf_w_d_ratio + 
        # misc characteristics of the catchment
        da_k_erodibility + da_avg_slope + mean_ndvi +
        # misc characteristics of the locality
        loc_bfi + loc_pct_clay + loc_pct_sand + loc_permeability + loc_bedrock_depth + loc_ppt_mean_mm +
        # channel confinement
        mtpi30_min + vb_bf_w_ratio + frac_leveed_longitudinal #+
        # baseflow as percent of mean annual flow
        # nf_bfl_dry_cfs_norm + nf_bfl_wet_cfs_norm #+ 
        # aquatic
        # bio_aq_rank_sw + species + species_fish + 
        # fixed effects
        #hqt_gradient_class + hyd_cls
        ) |>
  #step_log(all_numeric_predictors(), -mtpi30_min, -mean_ndvi, -bio_aq_rank_sw, -vb_bf_w_ratio, -frac_leveed_longitudinal) |>
  step_mutate_at(all_numeric_predictors(), fn = asinh) |>
  #step_dummy(hqt_gradient_class) |>
  #step_dummy(hyd_cls) |>
  step_interact(terms = ~ flow_norm_cfs:all_predictors()) |>
  step_naomit(all_predictors()) |>
  step_zv(all_predictors()) |>
  step_normalize(all_numeric_predictors()) 
```

#### Linear Regression

```{r lm-si, message=FALSE}
lm_spec <- linear_reg() 

lm_si <- workflow() |>
  add_recipe(si_rec) |>
  add_model(lm_spec) |>
  add_case_weights(case_wt) |>
  fit(data=training(td_split))

lm_si |> glance()
lm_si |> tidy()

lm_si_res <- testing(td_split) |>
  transmute(comid, flow_cfs, flow_norm_cfs, !!norm_var, hsi_frac,
            hsi_frac_pred = predict(lm_si, testing(td_split))[[".pred"]]) |>
  # crude constraint to between 0 and 1; better to use betareg
  mutate(hsi_frac_pred=pmax(pmin(hsi_frac_pred,1),0))

lm_si_res |>
  arrange(comid, flow_norm_cfs) |>
  ggplot() + geom_path(aes(x=hsi_frac, y=hsi_frac_pred, color=flow_norm_cfs, group=comid), linewidth=2) + 
  coord_fixed() + geom_abline() + scale_color_viridis_c(trans="log10", name="Frac. of MAF") + 
  ggtitle("Scale-independent model, linear regression") +
  xlab("HSI (Frac. of Channel Area) (Actual)") + ylab("HSI (Frac. of Channel Area) (Predicted)")

lm_si_res |> 
  filter(flow_cfs %in% c(300,500,1000,1500,3000,5000,10000,15000)) |>
  ggplot() + geom_point(aes(x=hsi_frac, y=hsi_frac_pred, color=flow_cfs)) + 
  facet_wrap(~flow_cfs, ncol = 4) +
  coord_fixed() + geom_abline() + scale_color_viridis_c(name = "Flow (cfs)") + 
  ggtitle("Scale-independent model, linear regression") +
  xlab("HSI (Frac. of Channel Area) (Actual)") + ylab("HSI (Frac. of Channel Area) (Predicted)")

lm_si_res |> filter(comid %in% test_comids) |>
  ggplot(aes(x=flow_cfs, group=1)) + facet_wrap(~comid) + 
  geom_line(aes(y=hsi_frac_pred*100, color="predicted", group=1)) + geom_line(aes(y=hsi_frac*100, color="actual", group=1)) + 
  xlab("Flow (cfs)") + ylab("Suitable Habitat Area (% of Total Area)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), panel.grid.minor=element_blank()) +
  scale_x_continuous(labels = scales::label_comma())
```

```{r lm-si-coeffs, fig.width=6.5, fig.height=6.5, dpi=300}
lm_si |>
  tidy() |>
  mutate(ixn_class = if_else(str_detect(term, "flow_norm_cfs_x_"), "term_flow_norm_cfs_x_", "term") |> 
                     factor(levels = c("term", "term_flow_norm_cfs_x_"), labels = c("Constant Effect", "Interaction with Normalized Flow")),
         term_root = str_replace(term, "flow_norm_cfs_x_", ""),
         coeff_dir = if_else(estimate>0, "Positive", "Negative"),
         ci_lower = estimate - qt(0.975, df = glance(lm_si)$df) * std.error,
         ci_upper = estimate + qt(0.975, df = glance(lm_si)$df) * std.error,
         estimate_abs = abs(estimate),
         ci_lower_abs = ci_lower * if_else(coeff_dir=="Negative",-1,1),
         ci_upper_abs = ci_upper * if_else(coeff_dir=="Negative",-1,1),
         is_significant = sign(ci_lower)==sign(ci_upper)
         ) |>
  filter(term!="flow_norm_cfs" & term!="(Intercept)") |>
  ggplot(aes(y = term_root, color = coeff_dir)) +
  facet_wrap(~ixn_class, ncol = 2, scales = "free_x") + 
  geom_errorbarh(aes(xmin = ci_lower_abs, xmax=ci_upper_abs)) + #pmax(0,ci_lower_abs), xmax = pmax(0,ci_upper_abs))) +
  geom_point(aes(x = estimate_abs, fill = is_significant), shape=21) +
  theme(legend.position="top",
        panel.grid.minor = element_blank(),
        panel.border = element_rect(fill = NA, color = "black"), 
        legend.direction = "vertical") +
  scale_y_discrete(limits=rev) + 
  scale_fill_manual(name = "Significant (95%)", values = c("FALSE" = "white", "TRUE" = "black")) +
  scale_color_manual(name = "Direction of Effect", values = c("Negative" = "goldenrod", "Positive" = "darkcyan")) +
  xlab("Coefficient Magnitude") + ylab("")
```

#### Random Forest Regression

```{r rfr-si-tune, eval=FALSE}
# set up cross validation
folds <- vfold_cv(training(td_split), v = 10, strata = dataset)
val_metrics <- metric_set(rmse, rsq, ccc)

# tuned random forest: how many trees to use?
rfr_spec <- rand_forest(mode = "regression", trees = tune())
rfr_grid <- tibble(trees = 2^seq(0, 12, 1))

rfr_si_tune <- 
  workflow() |>
  add_recipe(si_rec) |>
  add_model(rfr_spec) |>
  tune_grid(resamples = folds,
            grid = rfr_grid,
            metrics = val_metrics)
rfr_si_tune |>
  collect_metrics() |>
  ggplot(aes(x = trees, y = mean, color = .metric)) +
  geom_errorbar(aes(ymin = mean - std_err,
                    ymax = mean + std_err),
                width = 0.05) +
  geom_line() +
  facet_wrap(. ~ .metric, ncol = 1, scales="free") +
  theme(legend.position = "none") + scale_x_log10()
```

```{r rfr-si, message=FALSE}
mtry <- floor(length(si_rec$var_info$variable[which(si_rec$var_info$role=="predictor")])/3)
rfr_spec <- rand_forest(mode = "regression", trees = 2^8, mtry = mtry, min_n = 10) |> #, min_n = 10) |>
    set_engine("ranger", num.threads = 4)

rfr_si <- workflow() |>
  add_recipe(si_rec) |>
  add_model(rfr_spec) |>
  add_case_weights(case_wt) |>
  fit(data=training(td_split))

rfr_si$fit$fit |> print()

rfr_si_res <- testing(td_split) |>
  transmute(comid, flow_cfs, flow_norm_cfs, !!norm_var, hsi_frac,
            hsi_frac_pred = predict(rfr_si, testing(td_split))[[".pred"]]) |>
  # crude constraint to between 0 and 1; better to use betareg
  mutate(hsi_frac_pred=pmax(pmin(hsi_frac_pred,1),0))

rfr_si_res |>
  arrange(comid, flow_norm_cfs) |>
  ggplot() + geom_path(aes(x=hsi_frac, y=hsi_frac_pred, color=flow_norm_cfs, group=comid), linewidth=2) + 
  coord_fixed() + geom_abline() + scale_color_viridis_c(trans="log10", name="Frac. of MAF") + 
  ggtitle("Scale-independent model, random forest regression") +
  xlab("HSI (Frac. of Channel Area) (Actual)") + ylab("HSI (Frac. of Channel Area) (Predicted)")

rfr_si_res |> 
  filter(flow_cfs %in% c(300,500,1000,1500,3000,5000,10000,15000)) |>
  ggplot() + geom_point(aes(x=hsi_frac, y=hsi_frac_pred, color=flow_cfs)) + 
  facet_wrap(~flow_cfs, ncol = 4) +
  coord_fixed() + geom_abline() + scale_color_viridis_c(name = "Flow (cfs)") + 
  ggtitle("Scale-independent model, random forest regression") +
  xlab("HSI (Frac. of Channel Area) (Actual)") + ylab("HSI (Frac. of Channel Area) (Predicted)")

rfr_si_res |> filter(comid %in% test_comids) |>
  ggplot(aes(x=flow_cfs, group=1)) + facet_wrap(~comid) + 
  geom_line(aes(y=hsi_frac_pred*100, color="predicted", group=1)) + geom_line(aes(y=hsi_frac*100, color="actual", group=1)) + 
  xlab("Flow (cfs)") + ylab("Suitable Habitat Area (% of Total Area)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), panel.grid.minor=element_blank()) +
  scale_x_continuous(labels = scales::label_comma())
```

### Scale Independent Model 2: WUA-per-linear-ft (normalized by `r norm_var`) versus flow (normalized by `r norm_var`)

A potential compromise between the previous two models that scales the results by `r norm_var`, then unscales them after prediction.

```{r recipe-si2, message=FALSE}
si2_rec <- recipe(data=training(td_split), 
      formula=ihs_wua_per_lf_norm + case_wt ~ 
        # flow as percent of mean annual flow
        flow_norm_cfs + 
        # channel characteristics: gradient and sinuosity
        slope + sinuosity + 
        # flow and channel characteristics, hydrologically predicted
        #erom_v_ma_fps + bf_depth_m + 
        bf_w_d_ratio + 
        # misc characteristics of the catchment
        da_k_erodibility + da_avg_slope + mean_ndvi +
        # misc characteristics of the locality
        loc_bfi + loc_pct_clay + loc_pct_sand + loc_permeability + loc_bedrock_depth + loc_ppt_mean_mm +
        # channel confinement
        mtpi30_min + vb_bf_w_ratio + frac_leveed_longitudinal #+
        # baseflow as percent of mean annual flow
        # nf_bfl_dry_cfs_norm + nf_bfl_wet_cfs_norm #+ 
        # aquatic
        # bio_aq_rank_sw + species + species_fish + 
        # fixed effects
        # hqt_gradient_class + hyd_cls
        ) |>
  #step_log(all_numeric_predictors(), -mtpi30_min, -mean_ndvi, -bio_aq_rank_sw, -vb_bf_w_ratio, -frac_leveed_longitudinal) |>
  step_mutate_at(all_numeric_predictors(), fn = asinh) |>
  # step_dummy(hqt_gradient_class) |>
  # step_dummy(hyd_cls) |>
  step_interact(terms = ~ flow_norm_cfs:all_predictors()) |>
  step_naomit(all_predictors()) |>
  step_zv(all_predictors()) |>
  step_normalize(all_numeric_predictors())
```

#### Linear Regression

```{r lm-si2, message=FALSE}
lm_spec <- linear_reg() 

lm_si2 <- workflow() |>
  add_recipe(si2_rec) |>
  add_model(lm_spec) |>
  add_case_weights(case_wt) |>
  fit(data=training(td_split))

lm_si2 |> glance()
lm_si2 |> tidy()

lm_si2_res <- testing(td_split) |>
  transmute(comid, flow_norm_cfs, flow_cfs, !!norm_var,
            # actuals
            ihs_wua_per_lf_norm,
            wua_per_lf_norm = sinh(ihs_wua_per_lf_norm), 
            wua_per_lf = wua_per_lf_norm * !!norm_var,
            # predicted
            ihs_wua_per_lf_norm_pred = predict(lm_si2, testing(td_split))[[".pred"]],
            wua_per_lf_norm_pred = sinh(ihs_wua_per_lf_norm_pred),
            wua_per_lf_pred = wua_per_lf_norm_pred * !!norm_var)

lm_si2_res |>
  arrange(comid, flow_norm_cfs) |>
  ggplot() + geom_path(aes(x=wua_per_lf, y=wua_per_lf_pred, color=flow_cfs, group=comid), linewidth=2) + 
  coord_fixed() + geom_abline() + scale_color_viridis_c(trans="log10", name="Frac. of MAF") + 
  ggtitle("Scale-independent model, linear regression") +
  xlab("Suitable Habitat Area (ft2) per LF (Actual)") + ylab("Suitable Habitat Area (ft2) per LF (Predicted)")

lm_si2_res |> 
  filter(flow_cfs %in% c(300,500,1000,1500,3000,5000,10000,15000)) |>
  ggplot() + geom_point(aes(x=wua_per_lf, y=wua_per_lf_pred, color=flow_cfs)) + 
  facet_wrap(~flow_cfs, ncol = 4) +
  coord_fixed() + geom_abline() + scale_color_viridis_c(name = "Flow (cfs)") + 
  ggtitle("Scale-independent model, linear regression") +
  xlab("Suitable Habitat Area (ft2) per LF (Actual)") + ylab("Suitable Habitat Area (ft2) per LF (Predicted)")

lm_si2_res |> filter(comid %in% test_comids) |>
  ggplot(aes(x=flow_cfs, group=1)) + facet_wrap(~comid) + 
  geom_line(aes(y=wua_per_lf_pred, color="predicted", group=1)) + 
  geom_line(aes(y=wua_per_lf, color="actual", group=1)) + 
  scale_y_continuous(trans = ihs, limits = c(0, NA)) + 
  xlab("Flow (cfs)") + ylab("Suitable Habitat Area (ft2) per LF") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), panel.grid.minor=element_blank()) +
  scale_x_continuous(labels = scales::label_comma())
```

```{r lm-si2-coeffs, fig.width=6.5, fig.height=6.5, dpi=300}
lm_si2 |>
  tidy() |>
  mutate(ixn_class = if_else(str_detect(term, "flow_norm_cfs_x_"), "term_flow_norm_cfs_x_", "term") |> 
                     factor(levels = c("term", "term_flow_norm_cfs_x_"), labels = c("Constant Effect", "Interaction with Normalized Flow")),
         term_root = str_replace(term, "flow_norm_cfs_x_", ""),
         coeff_dir = if_else(estimate>0, "Positive", "Negative"),
         ci_lower = estimate - qt(0.975, df = glance(lm_si)$df) * std.error,
         ci_upper = estimate + qt(0.975, df = glance(lm_si)$df) * std.error,
         estimate_abs = abs(estimate),
         ci_lower_abs = ci_lower * if_else(coeff_dir=="Negative",-1,1),
         ci_upper_abs = ci_upper * if_else(coeff_dir=="Negative",-1,1),
         is_significant = sign(ci_lower)==sign(ci_upper)
         ) |>
  filter(term!="flow_norm_cfs" & term!="(Intercept)") |>
  ggplot(aes(y = term_root, color = coeff_dir)) +
  facet_wrap(~ixn_class, ncol = 2, scales = "free_x") + 
  geom_errorbarh(aes(xmin = ci_lower_abs, xmax=ci_upper_abs)) + #pmax(0,ci_lower_abs), xmax = pmax(0,ci_upper_abs))) +
  geom_point(aes(x = estimate_abs, fill = is_significant), shape=21) +
  theme(legend.position="top",
        panel.grid.minor = element_blank(),
        panel.border = element_rect(fill = NA, color = "black"), 
        legend.direction = "vertical") +
  scale_y_discrete(limits=rev) + 
  scale_fill_manual(name = "Significant (95%)", values = c("FALSE" = "white", "TRUE" = "black")) +
  scale_color_manual(name = "Direction of Effect", values = c("Negative" = "goldenrod", "Positive" = "darkcyan")) +
  xlab("Coefficient Magnitude") + ylab("")
```

#### Random Forest Regression

```{r rfr-si2, message=FALSE}
mtry <- floor(length(si2_rec$var_info$variable[which(si2_rec$var_info$role=="predictor")])/3)
rfr_spec <- rand_forest(mode = "regression", trees = 2^8, mtry = mtry, min_n = 10) |> #, min_n = 10) |>
    set_engine("ranger", num.threads = 4)

rfr_si2 <- workflow() |>
  add_recipe(si2_rec) |>
  add_model(rfr_spec) |>
  add_case_weights(case_wt) |>
  fit(data=training(td_split))

rfr_si2$fit$fit |> print()

rfr_si2_res <- testing(td_split) |>
  transmute(comid, flow_norm_cfs, flow_cfs, !!norm_var,
            # actuals
            ihs_wua_per_lf_norm,
            wua_per_lf_norm = sinh(ihs_wua_per_lf_norm), 
            wua_per_lf = wua_per_lf_norm * !!norm_var,
            # predicted
            ihs_wua_per_lf_norm_pred = predict(rfr_si2, testing(td_split))[[".pred"]],
            wua_per_lf_norm_pred = sinh(ihs_wua_per_lf_norm_pred),
            wua_per_lf_pred = wua_per_lf_norm_pred * !!norm_var)

rfr_si2_res |>
  arrange(comid, flow_norm_cfs) |>
  ggplot() + geom_path(aes(x=wua_per_lf, y=wua_per_lf_pred, color=flow_cfs, group=comid), linewidth=2) + 
  coord_fixed() + geom_abline() + scale_color_viridis_c(trans="log10", name="Frac. of MAF") + 
  ggtitle("Scale-independent model 2, random forest regression") +
  xlab("Suitable Habitat Area (ft2) per LF (Actual)") + ylab("Suitable Habitat Area (ft2) per LF (Predicted)")

rfr_si2_res |> 
  filter(flow_cfs %in% c(300,500,1000,1500,3000,5000,10000,15000)) |>
  ggplot() + geom_point(aes(x=wua_per_lf, y=wua_per_lf_pred, color=flow_cfs)) + 
  facet_wrap(~flow_cfs, ncol = 4) +
  coord_fixed() + geom_abline() + scale_color_viridis_c(name = "Flow (cfs)") + 
  ggtitle("Scale-independent model 2, random forest regression") +
  xlab("Suitable Habitat Area ft2) per LF (Actual)") + ylab("Suitable Habitat Area ft2) per LF (Predicted)")

rfr_si2_res |> filter(comid %in% test_comids) |>
  ggplot(aes(x=flow_cfs, group=1)) + facet_wrap(~comid) + 
  geom_line(aes(y=wua_per_lf_pred, color="predicted", group=1)) + 
  geom_line(aes(y=wua_per_lf, color="actual", group=1)) + 
  scale_y_continuous(trans = ihs, limits = c(0, NA)) + 
  xlab("Flow (cfs)") + ylab("Suitable Habitat Area (ft2) per LF")  +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), panel.grid.minor=element_blank()) +
  scale_x_continuous(labels = scales::label_comma())
```

## Prediction and Validation

```{r prediction-data, fig.width=6.5, fig.height=6.5, dpi=300}
# assemble prediction dataset
pd_attr <- flowline_attributes |> 
  # select just major streams
  filter_variable_ranges() |> #stream_level<=4 & !is.na(gnis_name) 
      filter(substr(reachcode,1, 4) %in% c("1802", "1803", "1804")) |> # Central Valley basin only
  # # filter out Valley Lowland areas
  # filter(hqt_gradient_class != "Valley Lowland") |>
  # # filter for just the areas that are hydrologically similar to the training data
  # filter(hyd_cls %in% c("High-volume snowmelt and rain",
  #                       "Low-volume snowmelt and rain")) |>
  # select the variables that are used in the model and drop NAs
  mutate(nf_bfl_dry_cfs_norm = coalesce(nf_bfl_dry_cfs / !!norm_var, 0), # normalized baseflow values
         nf_bfl_wet_cfs_norm = coalesce(nf_bfl_wet_cfs / !!norm_var, 0)) |>
  select(comid, any_of(c(sd_rec$var_info$variable, si_rec$var_info$variable)), !!norm_var, hyd_cls, hqt_gradient_class) |> 
  drop_na() 
  
pd <- pd_attr |>
  # create series of flow prediction points
  #expand_grid(flow_cfs = interp_flows) |>
  expand_grid(flow_cfs = c(oom_range(100, 10000), 15000)) |>
  # expand_grid(flow_cfs = c(0,signif(100*2^seq(-2,7,0.5),2))) |>
  #expand_grid(flow_cfs = c(0,100,250,500,1000,2500,5000,10000)) |>
  #mutate(flow_cfs = if_else(flow_cfs>0, flow_cfs, erom_q_ma_cfs)) |> # also evaluate at mean annual flow
  mutate(flow_norm_cfs = coalesce(flow_cfs / !!norm_var, 0)) |> # flow as a percent of mean annual flow
  arrange(comid, flow_cfs) |>
  glimpse()

flowlines_gcs |>
  inner_join(pd_attr, by=join_by(comid)) |>
  ggplot() + geom_sf(aes(color = da_area_sq_km)) + 
  scale_color_viridis_c(name = "Drainage area (km2)", trans="sqrt", breaks = scales::breaks_log(6), direction = -1)

flowlines_gcs |>
  inner_join(pd_attr, by=join_by(comid)) |>
  ggplot() + geom_sf(aes(color = hyd_cls)) +
  scale_color_manual(name = "eFlows hydrologic class",
                     values = c(
                       "Snowmelt" = "#F9F863",
                       "High-volume snowmelt and rain" = "#55BA36",
                       "Low-volume snowmelt and rain" = "#B4E749",
                       "Winter storms" = "#0011A6",
                       "Groundwater" = "#183F11",
                       "Perennial groundwater and rain" = "#2C6CD9",
                       "Flashy, ephemeral rain" = "#A2ACF9",
                       "Rain and seasonal groundwater" = "#89E0F8",
                       "High elevation, low precipitation" = "#D9D245"))
```

### One-step model, scale-dependent: WUA/LF vs flow

```{r prediction-output-sd, fig.width=6.5, fig.height=6.5, dpi=300}
sd_pred <- 
  sd_rec |> 
  prep(training(td_split)) |> 
  bake(pd) |># |> filter_within_domain(flow_cfs)) |> 
  bind_cols(pd |> #filter_within_domain(flow_cfs) |>
              transmute(comid, flow_cfs_orig = flow_cfs)) |>
  drop_na() %>% # need to use magrittr pipe for this purpose
  mutate(ihs_wua_per_lf_pred = predict(rfr_sd$fit$fit, new_data=.)[[".pred"]]) |>
  transmute(comid, 
            flow_cfs = flow_cfs_orig, 
            ihs_wua_per_lf_pred,
            wua_per_lf_pred = sinh(ihs_wua_per_lf_pred)) |>
  glimpse()

flowlines_gcs |>
  inner_join(sd_pred |> filter(round(flow_cfs)==800), by=join_by(comid)) |>
  mutate(wua_per_lf_pred = if_else(wua_per_lf_pred <1, 1, wua_per_lf_pred)) |> # for visualization only
  ggplot() + geom_sf(aes(color = wua_per_lf_pred)) + 
  scale_color_viridis_c(trans="log", breaks = scales::breaks_log(6), direction=-1)

n_breaks <- 10
sd_pred |>
  inner_join(flowline_attributes, by=join_by(comid), relationship="many-to-one") |>
  select(comid, wua_per_lf_pred, any_of(sd_rec$var_info$variable[which(sd_rec$var_info$role=="predictor")])) |>
  mutate(across(-(1:3), function(x) cut(x, breaks=n_breaks, labels=seq(1,n_breaks)/n_breaks))) |>
  pivot_longer(cols = -(1:3)) |>
  ggplot(aes(x = flow_cfs, y = wua_per_lf_pred)) +
  facet_wrap(~name) + geom_smooth(aes(color = value), se=F) +
  scale_color_viridis_d(name = "Percentile of predictor variable", option = "cividis", direction=-1) +
  theme(legend.position = "top") + guides(color=guide_legend(nrow=2, byrow=TRUE)) +
  scale_x_log10() + scale_y_log10() + theme(panel.grid.minor = element_blank())
```

### One-step model, scale-independent: normalized WUA/LF vs normalized flow

```{r prediction-output-si2, fig.width=6.5, fig.height=6.5, dpi=300}
si2_pred <- 
  si2_rec |> 
  prep(training(td_split)) |> 
  bake(pd) |> 
  # bring in unbaked variables
  bind_cols(pd |> transmute(comid, 
                            flow_cfs_orig = flow_cfs,
                            flow_norm_cfs_orig = flow_norm_cfs, 
                            "{norm_var}_orig" := !!norm_var)) |>
  drop_na() %>% # need to use magrittr pipe for this purpose
  mutate(ihs_wua_per_lf_norm_pred = predict(rfr_si2$fit$fit, new_data=.)[[".pred"]]) |>
  transmute(comid, 
            flow_cfs = flow_cfs_orig, 
            flow_norm_cfs = flow_norm_cfs_orig, 
            "{norm_var}" := !!sym(glue::glue("{norm_var}_orig")),
            ihs_wua_per_lf_norm_pred) |>
  mutate(wua_per_lf_norm_pred = sinh(ihs_wua_per_lf_norm_pred),
         wua_per_lf_pred = wua_per_lf_norm_pred * !!norm_var) |>
  glimpse()

flowlines_gcs |>
  inner_join(si2_pred |> filter(round(flow_cfs)==800), by=join_by(comid)) |>
  mutate(wua_per_lf_pred = if_else(wua_per_lf_pred <1, 1, wua_per_lf_pred)) |> # for visualization only
  ggplot() + geom_sf(aes(color = wua_per_lf_pred)) + 
  scale_color_viridis_c(trans="log", breaks = scales::breaks_log(6), direction=-1)

n_breaks <- 10
si2_pred |>
  inner_join(flowline_attributes, by=join_by(comid), relationship="many-to-one") |>
  select(comid, wua_per_lf_pred, flow_cfs, any_of(si2_rec$var_info$variable[which(si2_rec$var_info$role=="predictor")]), -flow_norm_cfs) |>
  mutate(across(-(1:3), function(x) cut(x, breaks=n_breaks, labels=seq(1,n_breaks)/n_breaks))) |>
  pivot_longer(cols = -(1:3)) |>
  ggplot(aes(x = flow_cfs, y = wua_per_lf_pred)) +
  facet_wrap(~name) + geom_smooth(aes(color = value), se=F) +
  scale_color_viridis_d(name = "Percentile of predictor variable", option = "cividis", direction=-1) +
  theme(legend.position = "top") + guides(color=guide_legend(nrow=2, byrow=TRUE)) +
  scale_x_log10() + scale_y_log10() + theme(panel.grid.minor = element_blank())
```

### Two-step model: (%HSI vs normalized flow) * (Total Area/LF vs flow)

```{r prediction-output-si, fig.width=6.5, fig.height=6.5, dpi=300}
sd2_pred <- 
  sd2_rec |> 
  prep(training(td_split)) |> 
  bake(pd) |> 
  bind_cols(pd |> transmute(comid, 
                            flow_cfs_orig = flow_cfs)) |>
  drop_na() %>% # need to use magrittr pipe for this purpose
  mutate(ihs_tot_area_per_lf_pred = predict(rfr_sd2$fit$fit, new_data=.)[[".pred"]]) |>
  transmute(comid, 
            flow_cfs = flow_cfs_orig, 
            ihs_tot_area_per_lf_pred,
            tot_area_per_lf_pred = sinh(ihs_tot_area_per_lf_pred)) |>
  glimpse()

sd2si_pred <- 
  si_rec |> 
  prep(training(td_split)) |> 
  bake(pd) |> 
  bind_cols(pd |> transmute(comid, 
                            flow_cfs_orig = flow_cfs,
                            flow_norm_cfs_orig = flow_norm_cfs,
                            "{norm_var}_orig" := !!norm_var)) |>
  drop_na() %>% # need to use magrittr pipe for this purpose
  mutate(hsi_frac_pred = predict(rfr_si$fit$fit, new_data=.)[[".pred"]]) |>
  transmute(comid, 
            flow_cfs = flow_cfs_orig, 
            flow_norm_cfs = flow_norm_cfs_orig, 
            hsi_frac_pred) |>
  #left_join(flowline_attributes |> transmute(comid, bf_width_ft = bf_width_m/0.3048), by=join_by(comid)) |>
  #mutate(wua_per_lf_pred = hsi_frac_pred * bf_width_ft) |>
  left_join(sd2_pred, by=join_by(comid, flow_cfs)) |>
  mutate(wua_per_lf_pred = hsi_frac_pred * tot_area_per_lf_pred) |>
  glimpse()

flowlines_gcs |>
  inner_join(sd2si_pred |> filter(round(flow_cfs)==800), by=join_by(comid)) |>
  mutate(wua_per_lf_pred = if_else(wua_per_lf_pred <1, 1, wua_per_lf_pred)) |> # for visualization only
  ggplot() + geom_sf(aes(color = wua_per_lf_pred)) + 
  scale_color_viridis_c(trans="log", breaks = scales::breaks_log(6), direction=-1)
```

## Predictions summarized by DSMHabitat reach

```{r prediction-flows}
# summary of typical flow ranges by DSMhabitat stream
prediction_flows <- readRDS(here::here("data-raw", "results", "watershed_flow_summary.Rds")) |>
  # make the numbers cleaner by rounding to nearest 100 cfs or 2 sig figs max
  mutate(across(-watershed, function(x) signif(round(x, -2), 2))) |>
  # establish ranges
  mutate(flow_cfs = map2(min_cfs, max_cfs, function(x, y) oom_range(x, y))) |>
  select(watershed, flow_cfs) |>
  unnest(flow_cfs) |>
  #instead of filtering here, filter later
  #filter(flow_cfs >= min(interp_flows) & flow_cfs <= max(interp_flows)) |>
  glimpse()

#prediction_flows |> group_by(watershed) |> summarize(max_flow = max(flow_cfs)) |> arrange(-max_flow) 
```

```{r dsmhabitat-prep, fig.width=6.5, fig.height=6.5, dpi=300}
mainstems_comid <- 
  st_read(file.path("/vsizip", here::here("data-raw", "source", "rearing_spatial_data", "nhdplusv2_comid_habitat_xw.shp.zip")), as_tibble=T) |>
  janitor::clean_names() |>
  st_zm() |>
  st_transform(st_crs(flowlines)) |>
  mutate(length_ft = st_length(geometry) |> units::set_units("ft") |> units::drop_units()) |>
  filter(habitat=="rearing") |>
  left_join(flowline_attributes |> select(comid, hqt_gradient_class), by=join_by(comid)) |>
  filter(!(river %in% c("Sacramento River", "San Joaquin River")))

mainstems <-
  mainstems_comid |>
  group_by(river, hqt_gradient_class) |>
  summarize() 

mainstems_comid |> 
  ggplot() + 
  geom_sf(aes(group=river, color=hqt_gradient_class)) + 
  theme(legend.key.height = unit(12, "point"))
```

```{r dsmhabitat-join, fig.width=6.5, fig.height=6.5, dpi=300}
watersheds <- mainstems |> pull(river) |> unique()
watershed_name <- tolower(gsub(pattern = "-| ", replacement = "_", x = watersheds))
watershed_rda_name <- paste(watershed_name, "floodplain", sep = "_")

dsm_habitat_floodplain <- map_df(watershed_rda_name, function(watershed) {
  df <- as.data.frame(do.call(`::`, list(pkg = "DSMhabitat", name = watershed)))
}) |> 
  transmute(river = watershed,
            flow_cfs,
            FR_floodplain_m2 = FR_floodplain_acres * 4046.86,
            FR_floodplain_m2_suitable = DSMhabitat::apply_suitability(FR_floodplain_m2),
            FR_floodplain_acres_suitable = FR_floodplain_m2_suitable / 4046.86)

dsm_habitat_instream <- map_df(paste(watershed_name, "instream", sep = "_"), 
                               possibly(function(watershed) {
                                 df <- as.data.frame(do.call(`::`, list(pkg = "DSMhabitat", name = watershed)))
                                 }, otherwise = NULL)) |> 
  transmute(river = watershed,
            flow_cfs,
            FR_juv_wua) 

dsm_flows <- bind_rows(dsm_habitat_floodplain, dsm_habitat_instream) |>
  group_by(river, flow_cfs) |>
  summarize() |>
  ungroup() |>
  arrange(river, flow_cfs)

dsm_flow_ranges <- 
  dsm_flows |> 
  group_by(river) |> 
  summarize(min_flow_cfs = min(flow_cfs), max_flow_cfs = max(flow_cfs))

mainstems_comid |> 
  st_zm() |> 
  filter(comid %in% mainstems_comid$comid) |>
  ggplot() + 
  geom_sf(aes(color=river)) + 
  theme(legend.key.height = unit(12, "point"))

# combining both floodplain and instream rearing acreages/WUAs for comparison
dsm_habitat_combined <- mainstems |> 
  mutate(length_ft = st_length(geometry) |> units::set_units("ft") |> units::drop_units()) |>
  group_by(river) |> 
  summarize(length_ft = sum(length_ft)) |>
  st_drop_geometry() |>
  inner_join(full_join(dsm_habitat_instream, dsm_habitat_floodplain, by=join_by(river, flow_cfs)), by=join_by(river)) |>
  group_by(river) |>
  arrange(river, flow_cfs) |>
  mutate(FR_juv_wua = zoo::na.approx(FR_juv_wua, flow_cfs, na.rm=F),
         FR_floodplain_acres_suitable = zoo::na.approx(FR_floodplain_acres_suitable, flow_cfs, na.rm=F)) |>
  transmute(river, flow_cfs, 
            instream_wua_per_lf = coalesce(FR_juv_wua/1000,0),
            instream_suitable_ac = coalesce(FR_juv_wua/1000,0)*length_ft/43560,
            floodplain_wua_per_lf = coalesce(FR_floodplain_acres_suitable,0)*43560/length_ft,
            floodplain_suitable_ac = coalesce(FR_floodplain_acres_suitable,0),
            combined_wua_per_lf = instream_wua_per_lf + floodplain_wua_per_lf,
            combined_suitable_ac =  instream_suitable_ac + floodplain_suitable_ac) |>
  ungroup()

dsm_habitat_wua_per_lf <- dsm_habitat_combined |>
  select(river, flow_cfs, instream_wua_per_lf, floodplain_wua_per_lf) |>
#  mutate(combined_wua_per_lf = pmax(instream_wua_per_lf, floodplain_wua_per_lf)) |>
  pivot_longer(cols=c(instream_wua_per_lf, floodplain_wua_per_lf)) |>
  mutate(name = paste("DSMhabitat", str_replace(name, "_wua_per_lf", "")),
         value = if_else(value>0, value, NA))

dsm_habitat_suitable_ac <- dsm_habitat_combined |>
  select(river, flow_cfs, instream_suitable_ac, floodplain_suitable_ac) |>
#  mutate(combined_suitable_ac = pmax(instream_suitable_ac, floodplain_suitable_ac)) |>
  pivot_longer(cols=c(instream_suitable_ac, floodplain_suitable_ac)) |>
  mutate(value = if_else(value>0, value, NA)) |>
  mutate(name = paste("DSMhabitat", str_replace(name, "_suitable_ac", "")),
         value = if_else(value>0, value, NA))
```

### One-step model, scale-dependent: WUA/LF vs flow

```{r dsmhabitat-sd-val, fig.width=6.5, fig.height=6.5, dpi=300}
pd_sd <- flowline_attributes |> 
  select(comid, any_of(sd_rec$var_info$variable)) |> 
  inner_join(select(mainstems_comid, comid, river), by=join_by(comid)) |>
  #expand_grid(flow_cfs = interp_flows) |>
  inner_join(prediction_flows, by=join_by(river==watershed), relationship="many-to-many") |>
  arrange(comid, flow_cfs) |>
  filter(comid %in% mainstems_comid$comid) |>
  drop_na() #

pd_sd_dsmhabitat_pred <- 
  sd_rec |> 
  prep(training(td_split)) |> 
  bake(pd_sd) |> 
  bind_cols(pd_sd |> transmute(comid, flow_cfs_orig=flow_cfs)) |>
  drop_na() %>% # need to use magrittr pipe for this purpose
  mutate(ihs_wua_per_lf_pred = predict(rfr_sd$fit$fit, new_data=.)[[".pred"]]) |>
  transmute(comid, flow_cfs=flow_cfs_orig, ihs_wua_per_lf_pred) |>
  mutate(wua_per_lf_pred = sinh(ihs_wua_per_lf_pred)) |>
  select(comid, 
         flow_cfs, wua_per_lf_pred)

pd_rf_by_mainstem <-
  mainstems_comid |> 
  st_drop_geometry() |>
  inner_join(pd_sd_dsmhabitat_pred, by=join_by(comid), relationship="one-to-many") |>
  mutate(length_ft = coalesce(length_ft,0),
         wua_per_lf_pred = coalesce(wua_per_lf_pred,0),
         wua_ft2_pred = wua_per_lf_pred * length_ft) 

pd_rf_by_mainstem_summary <- pd_rf_by_mainstem |>
  group_by(river, flow_cfs) |>
  summarize(tot_length_ft = sum(length_ft),
            tot_wua_ft2 = sum(wua_ft2_pred),
            tot_wua_ac = tot_wua_ft2 / 43560,
            avg_wua_ft2_per_lf = tot_wua_ft2 / tot_length_ft,
            avg_wua_ft2_per_1000ft = 1000 * avg_wua_ft2_per_lf)

pd_rf_by_mainstem |>
  ggplot(aes(x = flow_cfs)) + 
  geom_line(aes(y = wua_per_lf_pred, group = comid, color = hqt_gradient_class), alpha=0.5) + 
  geom_line(data=pd_rf_by_mainstem_summary, aes(y = avg_wua_ft2_per_lf), linewidth=1.0) + #geom_smooth(method="loess", se=F, color="black") +
  facet_wrap(~river, scales="free_y") + 
  scale_x_log10(labels=scales::comma_format(), expand=c(0,0), 
                limits=c(min(interp_flows),max(interp_flows))) + 
  scale_y_continuous() + 
  theme(legend.position="top", 
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  xlab("Flow (cfs)") + ylab("Suitable Habitat Area (ft2 per linear ft)") +
  scale_color_manual(values = palette_hqt_gradient_class)

pd_rf_by_mainstem_summary |>
  ggplot() + 
  geom_line(aes(x = flow_cfs, y = tot_wua_ac, color = river)) + #, color=species, linetype=habitat)) + 
  facet_wrap(~river, scales="free_y") + 
  scale_x_log10(labels=scales::comma_format(), expand=c(0,0), 
                limits=c(min(interp_flows),max(interp_flows))) + 
  scale_y_continuous() + 
  theme(legend.position="none", 
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  xlab("Flow (cfs)") + ylab("Total WUA (ac)")

pd_rf_by_mainstem_summary |>
  ggplot() + 
  geom_line(aes(x = flow_cfs, y = avg_wua_ft2_per_lf, color = river)) + #, color=species, linetype=habitat)) + 
  facet_wrap(~river, scales="free_y") + 
  scale_x_log10(labels=scales::comma_format(), expand=c(0,0), 
                limits=c(min(interp_flows),max(interp_flows))) + 
  scale_y_continuous() + 
  theme(legend.position="none", 
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  xlab("Flow (cfs)") + ylab("Suitable Habitat Area (ft2 per linear ft)")

pd_rf_by_mainstem_summary |> 
  #inner_join(dsm_habitat_floodplain, by=join_by(river==river, flow_cfs==flow_cfs)) |> 
  #filter(river == "American River") |> 
  ggplot() +
  geom_line(aes(x = flow_cfs, y = tot_wua_ac, color="habistat prediction")) + 
  geom_line(data=dsm_habitat_suitable_ac, aes(x = flow_cfs, y = value, color = name)) +  
  facet_wrap(~river) + #, scales="free_y") + 
  scale_x_log10(labels=scales::comma_format(), expand=c(0,0), 
                limits=c(min(interp_flows),max(interp_flows))) + 
  scale_y_log10(labels=scales::comma_format(), expand=c(0,0)) + 
  theme(legend.position="top", legend.title=element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  xlab("Flow (cfs)") + ylab("Total WUA (ac)") +
  scale_color_manual(values = palette_dsmhabitat_comparison)

pd_rf_by_mainstem_summary |> 
  #inner_join(dsm_habitat_instream, by=join_by(river==river, flow_cfs==flow_cfs)) |> 
  ggplot() +
  geom_line(aes(x = flow_cfs, y = avg_wua_ft2_per_lf, color="habistat prediction")) + 
  geom_line(data=dsm_habitat_wua_per_lf, aes(x = flow_cfs, y = value, color = name)) +  
  facet_wrap(~river) + #, scales="free_y") + 
  scale_x_log10(labels=scales::comma_format(), expand=c(0,0), 
                limits=c(min(interp_flows),max(interp_flows))) + 
  scale_y_log10(labels=scales::comma_format(), expand=c(0,0)) + 
  theme(legend.position="top", legend.title=element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  xlab("Flow (cfs)") + ylab("Suitable Habitat Area (ft2 per linear ft)") +
  scale_color_manual(values = palette_dsmhabitat_comparison)

pd_rf_by_mainstem_summary |> filter(flow_cfs>=min(interp_flows)) |>
  ggplot() +
  geom_line(aes(x = flow_cfs, y = tot_wua_ac, color="habistat prediction")) + 
  geom_line(data=dsm_habitat_suitable_ac, aes(x = flow_cfs, y = value, color = name)) +  
  facet_wrap(~river, scales="free_y") + 
  scale_x_log10(labels=scales::comma_format(), expand=c(0,0), 
                limits=c(min(interp_flows),max(interp_flows))) + 
  scale_y_continuous(labels=scales::comma_format(), expand=c(0,0), limits=c(0,NA)) + 
  theme(legend.position="top", legend.title=element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  xlab("Flow (cfs)") + ylab("Total WUA (ac)") +
  scale_color_manual(values = palette_dsmhabitat_comparison)

```

### One-step model, scale-independent: WUA/LF vs normalized flow

```{r dsmhabitat-si2-val, fig.width=6.5, fig.height=6.5, dpi=300}
pd_si2 <- flowline_attributes |> 
  select(comid, any_of(si2_rec$var_info$variable), !!norm_var) |> #erom_q_ma_cfs, nf_bfl_dry_cfs, nf_bfl_wet_cfs) |> 
  inner_join(select(mainstems_comid, comid, river), by=join_by(comid)) |>
  #expand_grid(flow_cfs = interp_flows) |>
  inner_join(prediction_flows, by=join_by(river==watershed), relationship="many-to-many") |>
  arrange(comid, flow_cfs) |>
  filter(comid %in% mainstems_comid$comid) |>
  mutate(flow_norm_cfs = flow_cfs / !!norm_var) |> #,
       #nf_bfl_dry_cfs_norm = nf_bfl_dry_cfs/erom_q_ma_cfs, 
       #nf_bfl_wet_cfs_norm = nf_bfl_wet_cfs/erom_q_ma_cfs) |>
  drop_na()

pd_si2_dsmhabitat_pred <- 
  si2_rec |> 
  prep(training(td_split)) |> 
  bake(pd_si2) |> 
  bind_cols(select(pd_si2, comid, flow_cfs_actual=flow_cfs, !!norm_var)) |>
  drop_na() %>% # need to use magrittr pipe for this purpose
  mutate(ihs_wua_per_lf_norm_pred = predict(rfr_si2$fit$fit, new_data=.)[[".pred"]]) |>
  mutate(wua_per_lf_pred = sinh(ihs_wua_per_lf_norm_pred) * !!norm_var) |> 
  select(comid, flow_norm_cfs, wua_per_lf_pred, flow_cfs=flow_cfs_actual, wua_per_lf_pred)

pd_rf_by_mainstem <-
  mainstems_comid |> 
  st_drop_geometry() |>
  inner_join(pd_si2_dsmhabitat_pred, by=join_by(comid), relationship="one-to-many") |>
  mutate(length_ft = coalesce(length_ft,0),
         wua_per_lf_pred = coalesce(wua_per_lf_pred,0),
         wua_ft2_pred = wua_per_lf_pred * length_ft) 

pd_rf_by_mainstem_summary <- pd_rf_by_mainstem |>
  group_by(river, flow_cfs) |>
  summarize(tot_length_ft = sum(length_ft),
            tot_wua_ft2 = sum(wua_ft2_pred),
            tot_wua_ac = tot_wua_ft2 / 43560,
            avg_wua_ft2_per_lf = tot_wua_ft2 / tot_length_ft,
            avg_wua_ft2_per_1000ft = 1000 * avg_wua_ft2_per_lf)

pd_rf_by_mainstem |>
  ggplot(aes(x = flow_cfs)) + 
  geom_line(aes(y = wua_per_lf_pred, group = comid, color = hqt_gradient_class), alpha=0.5) + 
  geom_line(data=pd_rf_by_mainstem_summary, aes(y = avg_wua_ft2_per_lf), linewidth=1.0) + #geom_smooth(method="loess", se=F, color="black") +
  facet_wrap(~river, scales="free_y") + 
  scale_x_log10(labels=scales::comma_format(), expand=c(0,0), 
                limits=c(min(interp_flows),max(interp_flows))) + 
  scale_y_continuous() + 
  theme(legend.position="top", 
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  xlab("Flow (cfs)") + ylab("Suitable Habitat Area (ft2 per linear ft)") +
  scale_color_manual(values = palette_hqt_gradient_class)
  
pd_rf_by_mainstem_summary |>
  ggplot() + 
  geom_line(aes(x = flow_cfs, y = tot_wua_ac, color = river)) + #, color=species, linetype=habitat)) + 
  facet_wrap(~river, scales="free_y") + 
  scale_x_log10(labels=scales::comma_format(), expand=c(0,0), 
                limits=c(min(interp_flows),max(interp_flows))) + 
  scale_y_continuous() + 
  theme(legend.position="none", 
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  xlab("Flow (cfs)") + ylab("Total WUA (ac)")

pd_rf_by_mainstem_summary |>
  ggplot() + 
  geom_line(aes(x = flow_cfs, y = avg_wua_ft2_per_lf, color = river)) + #, color=species, linetype=habitat)) + 
  facet_wrap(~river, scales="free_y") + 
  scale_x_log10(labels=scales::comma_format(), expand=c(0,0), 
                limits=c(min(interp_flows),max(interp_flows))) + 
  scale_y_continuous() + 
  theme(legend.position="none", 
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  xlab("Flow (cfs)") + ylab("Suitable Habitat Area (ft2 per linear ft)")

pd_rf_by_mainstem_summary |> 
  #inner_join(dsm_habitat_floodplain, by=join_by(river==river, flow_cfs==flow_cfs)) |> 
  #filter(river == "American River") |> 
  ggplot() +
  geom_line(aes(x = flow_cfs, y = tot_wua_ac, color="habistat prediction")) + 
  geom_line(data=dsm_habitat_suitable_ac, aes(x = flow_cfs, y = value, color = name)) +  
  facet_wrap(~river) + #, scales="free_y") + 
  scale_x_log10(labels=scales::comma_format(), expand=c(0,0), 
                limits=c(min(interp_flows),max(interp_flows))) + 
  scale_y_log10(labels=scales::comma_format(), expand=c(0,0)) + 
  theme(legend.position="top", legend.title=element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  xlab("Flow (cfs)") + ylab("Total WUA (ac)") +
  scale_color_manual(values = palette_dsmhabitat_comparison)

pd_rf_by_mainstem_summary |> 
  #inner_join(dsm_habitat_instream, by=join_by(river==river, flow_cfs==flow_cfs)) |> 
  ggplot() +
  geom_line(aes(x = flow_cfs, y = avg_wua_ft2_per_lf, color="habistat prediction")) + 
  geom_line(data=dsm_habitat_wua_per_lf, aes(x = flow_cfs, y = value, color = name)) +  
  facet_wrap(~river) + #, scales="free_y") + 
  scale_x_log10(labels=scales::comma_format(), expand=c(0,0), 
                limits=c(min(interp_flows),max(interp_flows))) + 
  scale_y_log10(labels=scales::comma_format(), expand=c(0,0)) + 
  theme(legend.position="top", legend.title=element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  xlab("Flow (cfs)") + ylab("Suitable Habitat Area (ft2 per linear ft)") +
  scale_color_manual(values = palette_dsmhabitat_comparison)

pd_rf_by_mainstem_summary |> filter(flow_cfs>=min(interp_flows)) |>
  ggplot() +
  geom_line(aes(x = flow_cfs, y = tot_wua_ac, color="habistat prediction")) + 
  geom_line(data=dsm_habitat_suitable_ac, aes(x = flow_cfs, y = value, color = name)) +  
  facet_wrap(~river, scales="free_y") + 
  scale_x_log10(labels=scales::comma_format(), expand=c(0,0), 
                limits=c(min(interp_flows),max(interp_flows))) + 
  scale_y_continuous(labels=scales::comma_format(), expand=c(0,0), limits=c(0,NA)) + 
  theme(legend.position="top", legend.title=element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  xlab("Flow (cfs)") + ylab("Total WUA (ac)") +
  scale_color_manual(values = palette_dsmhabitat_comparison)

```

### Two-step model: (%HSI vs normalized flow) \* (Total Area/LF vs flow)

```{r dsmhabitat-si-val, fig.width=6.5, fig.height=6.5, dpi=300}
pd_sd2 <- flowline_attributes |> 
  select(comid, any_of(sd2_rec$var_info$variable)) |> 
  inner_join(select(mainstems_comid, comid, river), by=join_by(comid)) |>
  #expand_grid(flow_cfs = interp_flows) |>
  inner_join(prediction_flows, by=join_by(river==watershed), relationship="many-to-many") |>
  arrange(comid, flow_cfs) |>
  filter(comid %in% mainstems_comid$comid) |>
  drop_na()

pd_sd2_dsmhabitat_pred <- 
  sd2_rec |> 
  prep(training(td_split)) |> 
  bake(pd_sd2) |> 
  bind_cols(select(pd_sd2, comid, flow_cfs_orig=flow_cfs)) |>
  drop_na() %>% # need to use magrittr pipe for this purpose
  mutate(ihs_tot_area_per_lf_pred = predict(rfr_sd2$fit$fit, new_data=.)[[".pred"]]) |>
  mutate(tot_area_per_lf_pred = sinh(ihs_tot_area_per_lf_pred) ) |> 
  select(comid, flow_cfs=flow_cfs_orig, tot_area_per_lf_pred)

pd_si <- flowline_attributes |> 
  select(comid, any_of(si_rec$var_info$variable), !!norm_var) |> #, erom_q_ma_cfs, nf_bfl_dry_cfs, nf_bfl_wet_cfs) |> 
  inner_join(select(mainstems_comid, comid, river), by=join_by(comid)) |>
  #expand_grid(flow_cfs = interp_flows) |>
  inner_join(prediction_flows, by=join_by(river==watershed), relationship="many-to-many") |>
  arrange(comid, flow_cfs) |>
  filter(comid %in% mainstems_comid$comid) |>
  mutate(flow_norm_cfs = flow_cfs / !!norm_var) |>#,
       #nf_bfl_dry_cfs_norm = nf_bfl_dry_cfs/erom_q_ma_cfs, 
       #nf_bfl_wet_cfs_norm = nf_bfl_wet_cfs/erom_q_ma_cfs) |>
  drop_na()

pd_si_dsmhabitat_pred <- 
  si_rec |> 
  prep(training(td_split)) |> 
  bake(pd_si) |> 
  bind_cols(select(pd_si, comid, flow_cfs)) |>
  drop_na() %>% # need to use magrittr pipe for this purpose
  mutate(hsi_pred = predict(rfr_si$fit$fit, new_data=.)[[".pred"]]) |>
  #inner_join(flowline_attributes |> transmute(comid, length_ft = reach_length_km * 3280.84, width_ft = bf_width_m/3.2808), by=join_by(comid)) |>   
  #mutate(wua_per_lf_pred = hsi_pred * width_ft * length_ft / length_ft ) |> 
  left_join(pd_sd2_dsmhabitat_pred, by=join_by(comid, flow_cfs)) |>
  mutate(wua_per_lf_pred = hsi_pred * tot_area_per_lf_pred) |>
  select(comid, flow_norm_cfs, flow_cfs, hsi_pred, tot_area_per_lf_pred, wua_per_lf_pred)

pd_rf_by_mainstem <-
  mainstems_comid |> 
  st_drop_geometry() |>
  inner_join(pd_si_dsmhabitat_pred, by=join_by(comid), relationship="one-to-many") |>
  mutate(length_ft = coalesce(length_ft,0),
         wua_per_lf_pred = coalesce(wua_per_lf_pred,0),
         wua_ft2_pred = wua_per_lf_pred * length_ft) 

pd_rf_by_mainstem_summary <- pd_rf_by_mainstem |>
  group_by(river, flow_cfs) |>
  summarize(tot_length_ft = sum(length_ft),
            tot_wua_ft2 = sum(wua_ft2_pred),
            tot_wua_ac = tot_wua_ft2 / 43560,
            avg_wua_ft2_per_lf = tot_wua_ft2 / tot_length_ft,
            avg_wua_ft2_per_1000ft = 1000 * avg_wua_ft2_per_lf)

pd_rf_by_mainstem |>
  ggplot(aes(x = flow_cfs)) + 
  geom_line(aes(y = wua_per_lf_pred, group = comid, color = hqt_gradient_class), alpha=0.5) + 
  geom_line(data=pd_rf_by_mainstem_summary, aes(y = avg_wua_ft2_per_lf), linewidth=1.0) + #geom_smooth(method="loess", se=F, color="black") +
  facet_wrap(~river, scales="free_y") + 
  scale_x_log10(labels=scales::comma_format(), expand=c(0,0), 
                limits=c(min(interp_flows),max(interp_flows))) + 
  scale_y_continuous() + 
  theme(legend.position="top", 
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  xlab("Flow (cfs)") + ylab("Suitable Habitat Area (ft2 per linear ft)") +
  scale_color_manual(values = palette_hqt_gradient_class)

pd_rf_by_mainstem_summary |>
  ggplot() + 
  geom_line(aes(x = flow_cfs, y = tot_wua_ac, color = river)) + #, color=species, linetype=habitat)) + 
  facet_wrap(~river, scales="free_y") + 
  scale_x_log10(labels=scales::comma_format(), expand=c(0,0), 
                limits=c(min(interp_flows),max(interp_flows))) + 
  scale_y_continuous() + 
  theme(legend.position="none", 
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  xlab("Flow (cfs)") + ylab("Total WUA (ac)")

pd_rf_by_mainstem_summary |>
  ggplot() + 
  geom_line(aes(x = flow_cfs, y = avg_wua_ft2_per_lf, color = river)) + #, color=species, linetype=habitat)) + 
  facet_wrap(~river, scales="free_y") + 
  scale_x_log10(labels=scales::comma_format(), expand=c(0,0), 
                limits=c(min(interp_flows),max(interp_flows))) + 
  scale_y_continuous() + 
  theme(legend.position="none", 
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  xlab("Flow (cfs)") + ylab("Suitable Habitat Area (ft2 per linear ft)")

pd_rf_by_mainstem_summary |> 
  #inner_join(dsm_habitat_floodplain, by=join_by(river==river, flow_cfs==flow_cfs)) |> 
  #filter(river == "American River") |> 
  ggplot() +
  geom_line(aes(x = flow_cfs, y = tot_wua_ac, color="habistat prediction")) + 
  geom_line(data=dsm_habitat_suitable_ac, aes(x = flow_cfs, y = value, color = name)) +  
  facet_wrap(~river) + #, scales="free_y") + 
  scale_x_log10(labels=scales::comma_format(), expand=c(0,0), 
                limits=c(min(interp_flows),max(interp_flows))) + 
  scale_y_log10(labels=scales::comma_format(), expand=c(0,0)) + 
  theme(legend.position="top", legend.title=element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  xlab("Flow (cfs)") + ylab("Total WUA (ac)") +
  scale_color_manual(values = palette_dsmhabitat_comparison)

pd_rf_by_mainstem_summary |> 
  #inner_join(dsm_habitat_instream, by=join_by(river==river, flow_cfs==flow_cfs)) |> 
  ggplot() +
  geom_line(aes(x = flow_cfs, y = avg_wua_ft2_per_lf, color="habistat prediction")) + 
  geom_line(data=dsm_habitat_wua_per_lf, aes(x = flow_cfs, y = value, color = name)) +  
  facet_wrap(~river) + #, scales="free_y") + 
  scale_x_log10(labels=scales::comma_format(), expand=c(0,0), 
                limits=c(min(interp_flows),max(interp_flows))) + 
  scale_y_log10(labels=scales::comma_format(), expand=c(0,0)) + 
  theme(legend.position="top", legend.title=element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  xlab("Flow (cfs)") + ylab("Suitable Habitat Area (ft2 per linear ft)") +
  scale_color_manual(values = palette_dsmhabitat_comparison)

pd_rf_by_mainstem_summary |> filter(flow_cfs>=min(interp_flows)) |>
  ggplot() +
  geom_line(aes(x = flow_cfs, y = tot_wua_ac, color="habistat prediction")) + 
  geom_line(data=dsm_habitat_suitable_ac, aes(x = flow_cfs, y = value, color = name)) +  
  facet_wrap(~river, scales="free_y") + 
  scale_x_log10(labels=scales::comma_format(), expand=c(0,0), 
                limits=c(min(interp_flows),max(interp_flows))) + 
  scale_y_continuous(labels=scales::comma_format(), expand=c(0,0), limits=c(0,NA)) + 
  theme(legend.position="top", legend.title=element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  xlab("Flow (cfs)") + ylab("Total WUA (ac)") +
  scale_color_manual(values = palette_dsmhabitat_comparison)

```

## Results Export

```{r}
wua_predicted <- 
  pd |>
  inner_join(sd_pred |> select(comid, flow_cfs, wua_per_lf_pred_sd = wua_per_lf_pred),
             by=join_by(comid, flow_cfs), relationship="one-to-one") |>
  inner_join(si2_pred |> select(comid, flow_cfs, wua_per_lf_pred_si2 = wua_per_lf_pred),
             by=join_by(comid, flow_cfs), relationship="one-to-one") |>
  inner_join(sd2si_pred |> select(comid, flow_cfs, wua_per_lf_pred_sd2si = wua_per_lf_pred),
             by=join_by(comid, flow_cfs), relationship="one-to-one") |>
  #filter(row_number() == 26143) |>
  left_join(td |> select(comid, flow_cfs, wua_per_lf_actual = wua_per_lf),
            by=join_by(comid, flow_cfs), relationship="one-to-one")
  
 # td |> filter(comid==2823726)

wua_predicted |> saveRDS(here::here("data-raw", "results", "predictions_table.Rds"))
wua_predicted |> usethis::use_data(overwrite = TRUE)

# to run the leaflet map: shiny::runApp(here::here("inst", "app"))

# hs_predictions <- predictions_export
# usethis::use_data(hs_predictions, overwrite = T)

td |> filter(str_detect(dataset,"Tuolumne River")) |>
  select(comid, flow_cfs) |> arrange(comid, flow_cfs)
```

## Yuba Comparison

Non-temporal comparison

```{r yuba-comparison, fig.width = 6.5, fig.height = 4, dpi = 300}
lyr_cbec_reaches <- tribble(
  ~reach_id, ~reach_name, ~length_lf, ~rm_ds, ~rm_us,
  "FR", "Feather River", 31680, 0.2, 6.2,
  "DGR", "Daguerre Point Dam", 28512, 6.2, 11.6,
  "HR", "Hammon", 36960, 11.6, 18.6,
  "TBR", "Timbuctoo Bend", 19536, 18.6, 22.3,
  "NR", "Narrows", 5808, 22.3, 23.4,
  "EDR", "Englebright Dam", 3696, 23.4, 24.1) |>
  mutate(reach_name = as_factor(reach_name),
         reach_id = as_factor(reach_id))
  
lyr_cbec_nontemporal <- 
  read_csv(here::here("data-raw", "source", "hydraulic_model_data", "yuba_srh2d_2020", "LYR_WUA_Discharge_2020_06_30_Nontemporal_WUA_by_Domain.csv")) |>
  select(-"Total WUA") |>
  rename(flow_cfs = Discharge) |>
  pivot_longer(cols = ends_with("WUA"),
               names_to = "lyr_reach",
               values_to = "wua_ac",
               names_transform = function(x) str_replace_all(x, " WUA", "")) |>
  inner_join(lyr_cbec_reaches, by=join_by(lyr_reach == reach_id)) |>
  mutate(wua_ft2 = wua_ac * 43560,
         wua_ft2_per_lf = wua_ft2 / length_lf)

lyr_cbec_nontemporal_summary <-
  lyr_cbec_nontemporal |> 
  group_by(flow_cfs) |> 
  summarize(across(c(wua_ac, wua_ft2, length_lf), sum)) |>
  mutate(wua_ft2_per_lf = wua_ft2 / length_lf)

yuba_wua_predicted <-
  wua_predicted |> 
  select(comid, flow_cfs, starts_with("wua_per_lf")) |>
  filter(comid %in% filter(mainstems_comid, river=="Yuba River")$comid) |>
  filter(!is.na(wua_per_lf_actual)) |>
  inner_join(flowline_attributes |> transmute(comid, length_ft = reach_length_km*3280.84), by=join_by(comid))

yuba_wua_predicted_summary <- 
  yuba_wua_predicted |> 
  group_by(flow_cfs) |>
  mutate(across(starts_with("wua_per_lf"), function(x) x * length_ft)) |>
  summarize(across(c(length_ft, starts_with("wua_per_lf")), sum)) |>
  mutate(across(starts_with("wua_per_lf"), function(x) x / length_ft))

yuba_dsmhabitat_summary <-
  dsm_habitat_wua_per_lf |> 
  filter(river=="Yuba River") |>
  mutate(name = str_replace(name, " ", "_")) |>
  pivot_wider() |>
  mutate(DSMhabitat_total = coalesce(DSMhabitat_instream,0) + coalesce(DSMhabitat_floodplain,0))

ggplot() + 
  geom_line(data = yuba_wua_predicted_summary,
            aes(x = flow_cfs, y = wua_per_lf_actual, color = "Habistat Training (LYR SRH2D)", linetype = "instream + floodplain"), linewidth=1) +
  geom_line(data = yuba_wua_predicted_summary,
            aes(x = flow_cfs, y = wua_per_lf_pred_sd, color = "Habistat Model 1", linetype = "instream + floodplain"), linewidth=1) +
  geom_line(data = yuba_wua_predicted_summary,
            aes(x = flow_cfs, y = wua_per_lf_pred_si2, color = "Habistat Model 2", linetype = "instream + floodplain"), linewidth=1) +
  geom_line(data = lyr_cbec_nontemporal_summary,
            aes(x = flow_cfs, y = wua_ft2_per_lf, color = "CBEC Study (LYR SRH2D)", linetype = "instream + floodplain"), linewidth=1) +
  geom_line(data = yuba_dsmhabitat_summary,
            aes(x = flow_cfs, y = DSMhabitat_total, color = "DSMhabitat", linetype = "instream + floodplain"), linewidth=1) +
  geom_line(data = yuba_dsmhabitat_summary,
            aes(x = flow_cfs, y = DSMhabitat_instream, color = "DSMhabitat", linetype = "instream"), linewidth=1) +
  geom_line(data = yuba_dsmhabitat_summary,
            aes(x = flow_cfs, y = DSMhabitat_floodplain, color = "DSMhabitat", linetype = "floodplain"), linewidth=1) + 
  scale_x_log10(labels = scales::label_comma(), limits = c(NA, 15000)) + 
  scale_y_continuous(limits = c(0, 60), expand = c(0, 0)) +
  scale_linetype_manual(name = "domain", 
                        values = c("instream + floodplain" = "solid", 
                                   "instream" = "dashed", 
                                   "floodplain" = "dotted")) +
  xlab("Flow (cfs)") + ylab("Suitable habitat area (ft2) per linear ft") +
  theme(panel.grid.minor = element_blank()) + annotation_logticks(sides = "b") +
  ggtitle("WUA Estimates: HabiStat vs CBEC vs DSMhabitat") +
  scale_color_manual(name = "dataset", 
                     values = c("DSMhabitat" = "#8cc2ca",
                                "CBEC Study (LYR SRH2D)" = "#6388b4",
                                "Habistat Training (LYR SRH2D)" = "#55ad79",
                                "Habistat Model 1" = "#ffae34",
                                "Habistat Model 2" = "#c3bc3f"))
  
```

```{r yuba-comparison-temporal, fig.width = 6.5, fig.height = 4, dpi = 300}
lyr_cbec_temporal <- 
  read_csv(here::here("data-raw", "source", "hydraulic_model_data", "yuba_srh2d_2020", "LYR_WUA_Discharge_2020_06_30_Temporal_WUA_by_Domain_by_WY.csv")) |>
  mutate(Date_Time = mdy(Date_Time)) |>
  select(-"Total WUA") |>
  rename(flow_cfs = Flow_cfs, date_time = Date_Time, wy_day = Day) |>
  mutate(water_year = if_else(month(date_time)>=10, year(date_time)+1, year(date_time))) |>
  pivot_longer(cols = ends_with("WUA"),
               names_to = "lyr_reach",
               values_to = "wua_ac",
               names_transform = function(x) str_replace_all(x, " WUA", "")) |>
  inner_join(lyr_cbec_reaches, by=join_by(lyr_reach == reach_id)) |>
  mutate(wua_ft2 = wua_ac * 43560,
         wua_ft2_per_lf = wua_ft2 / length_lf)

lyr_cbec_temporal_summary <-
  lyr_cbec_temporal |> 
  group_by(water_year, date_time, flow_cfs) |> 
  summarize(across(c(wua_ac, wua_ft2, length_lf), sum)) |>
  mutate(wua_ft2_per_lf = wua_ft2 / length_lf)

yuba_wua_predicted_temporal <- 
  readRDS(here::here("data-raw", "results", "durhsi_applied_to_comid.Rds")) |>
  filter(dataset == "Lower Yuba River") |>
  filter(water_year %in% unique(lyr_cbec_temporal$water_year)) |>
  select(comid, water_year, flow_cfs = q, wua, durwua) |>
  inner_join(flowline_attributes |> transmute(comid, length_ft = reach_length_km*3280.84), by=join_by(comid))

yuba_wua_predicted_temporal_summary <- 
  yuba_wua_predicted_temporal |> 
  group_by(flow_cfs, water_year) |>
  summarize(wua = sum(wua * length_ft) / sum(length_ft), 
            durwua = sum(durwua * length_ft) / sum(length_ft))

source(here::here("data-raw", "scripts", "inundation-duration-functions.R"))
yuba_wua_predicted_summary_bfcremoved <- 
  yuba_wua_predicted_temporal_summary |> group_by(flow_cfs, wua) |> summarize() |> ungroup() |>
  eliminate_baseflow(600, fsa_q=flow_cfs, fsa_wua=wua)

yuba_wua_predicted_temporal_ts <-
  readRDS(here::here("data-raw", "results", "durhsi_applied_to_comid_ts.Rds")) |>
  ungroup() |>
    filter(river == "Lower Yuba River") |>
  filter(water_year %in% unique(lyr_cbec_temporal$water_year)) |>
  select(comid, water_year, flow_cfs = q, durwua, date) |>
  inner_join(flowline_attributes |> transmute(comid, length_ft = reach_length_km*3280.84), by=join_by(comid))

yuba_wua_predicted_temporal_ts_summary <- 
  yuba_wua_predicted_temporal_ts |> 
  group_by(flow_cfs, water_year, date) |>
  summarize(durwua = sum(durwua * length_ft) / sum(length_ft))

ggplot() + 
  geom_point(data = lyr_cbec_temporal_summary,
            aes(x = flow_cfs, y = wua_ft2_per_lf, color = "CBEC Study (LYR SRH2D)"), size = 0.7, stroke = 0) +
#    geom_point(data = yuba_wua_predicted_temporal_ts_summary,
#            aes(x = flow_cfs, y = durwua, color = "Habistat Model 1"), size = 0.7, stroke = 0) +
  geom_line(data = lyr_cbec_nontemporal_summary,
            aes(x = flow_cfs, y = wua_ft2_per_lf, color = "CBEC Study (LYR SRH2D)", linetype = "Non-temporal", linewidth = "Non-temporal")) +
  geom_line(data = yuba_wua_predicted_temporal_summary,
            aes(x = flow_cfs, y = durwua, group = water_year, color = "Habistat Model 1", linetype = "Temporal", linewidth = "Temporal")) +
  geom_line(data = yuba_wua_predicted_temporal_summary,
            aes(x = flow_cfs, y = wua, color = "Habistat Model 1", linetype = "Non-temporal", linewidth = "Non-temporal")) +
  geom_line(data = yuba_wua_predicted_summary_bfcremoved,
            aes(x = flow_cfs, y = wua, color = "Habistat Model 1", linetype = "Non-temporal, no baseflow", linewidth = "Non-temporal, no baseflow")) +
  scale_x_log10(labels = scales::label_comma(), limits = c(NA, 15000)) + 
  scale_y_continuous(limits = c(0, 60), expand = c(0, 0)) +
  xlab("Flow (cfs)") + ylab("Suitable habitat area (ft2) per linear ft") +
  theme(panel.grid.minor = element_blank()) + annotation_logticks(sides = "b") +
  scale_linewidth_manual(name = "model type", values = c("Non-temporal" = 1, "Temporal" = 0.7, "Non-temporal, no baseflow" = 1)) +
  scale_linetype_manual(name = "model type", values = c("Non-temporal" = "solid", "Temporal" = "dotted", "Non-temporal, no baseflow" = "dashed")) + 
  ggtitle("Temporal WUA Estimates: HabiStat vs CBEC")  +
  scale_color_manual(name = "dataset", 
                     values = c("CBEC Study (LYR SRH2D)" = "#6388b4",
                                "Habistat Model 1" = "#ffae34"))
```

