---
title: "Spawning Gravel Carrying Capacity"
author: "[Skyler Lewis](mailto:slewis@flowwest.com)"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
    toc_depth: 3
    number_sections: false
    math_method:
      engine: webtex
      url: https://latex.codecogs.com/svg.image?
---

```{r setup}
library(tidyverse)
library(sf)
library(habistat)
library(patchwork)
theme_set(theme_minimal())
glimpse_plot <- function(x, ...) {
  plot(x, ...)
  invisible(x)
}
knitr::opts_chunk$set(fig.width=6.5, fig.height=4, dpi=300)
```

Map current and historical habitat reaches of the Yuba and its major tributaries

```{r}
habistat::cv_mainstems |> 
  filter(river_group == "Yuba River") |>
  filter(!is.na(habitat)) |>
  glimpse_plot()
```

Map full lengths of Yuba and its major tributaries, and join in spawning context layer

```{r}
spawning_context <- 
  readRDS(here::here("data-raw", "results", "spawning_context.Rds"))

selected_streams <- 
  habistat::flowline_geom_proj |> 
  inner_join(habistat::flowline_attr |> select(comid, watershed_level_3, gnis_name), by=join_by(comid)) |>
  left_join(habistat::cv_mainstems |> st_drop_geometry() |> select(comid, habitat), by=join_by(comid)) |>
  filter(watershed_level_3 == "Yuba River") |> 
  filter(gnis_name %in% c("Yuba River", "Middle Yuba River", "North Yuba River", "South Yuba River")) |> 
  mutate(river_name = if_else(str_detect(coalesce(habitat,""), "rearing"), "Lower Yuba River", gnis_name) |>
         factor(levels = c("North Yuba River", "Middle Yuba River", "South Yuba River", "Yuba River", "Lower Yuba River"))) |>
  left_join(spawning_context, by=join_by(comid)) |>
  mutate(across(starts_with("spawning"), function(x) coalesce(x, FALSE))) |>
  mutate(current_spawning_reach = coalesce(str_detect(habitat, "spawning"), FALSE)) |>
  glimpse_plot(max.plot = Inf)
```

## Actuals

Import estimates from the Yuba relicensing study

```{r}
spawning_matrix_data <- read_csv(here::here("data-raw", "source", "spawning_data", "01_YSF-Habitat-Matrices-Report_table8_data.csv"), col_types = "cnnnnnnnnnnn") |>
  mutate(gnis_name = case_when(str_detect(reach, "North Yuba") ~ "North Yuba River",
                               str_detect(reach, "Middle Yuba") ~ "Middle Yuba River",
                               str_detect(reach, "South Yuba") ~ "South Yuba River",
                               TRUE ~ "Yuba River")) |>
  mutate(river_name = case_when(str_detect(reach, "North Yuba") ~ "North Yuba River",
                               str_detect(reach, "Middle Yuba") ~ "Middle Yuba River",
                               str_detect(reach, "South Yuba") ~ "South Yuba River",
                               str_detect(reach, "Lower Yuba") ~ "Lower Yuba River",
                               TRUE ~ "Yuba River")) |>
  glimpse()
```

Summarize

```{r}
spawning_matrix_summary <-
  spawning_matrix_data |>
  group_by(river_name, year) |>
  summarize(across(c(starts_with("length_"), starts_with("gravel_"), starts_with("redds_")), function(x) sum(x, na.rm=T)), .groups="drop") |>
  group_by(river_name) |>
  summarize(across(c(starts_with("length_"), starts_with("gravel_"), starts_with("redds_")), mean)) |>
  pivot_longer(cols = -river_name) |>
  separate(name, sep = "_", into = c("name", "method"), extra = "merge") |>
  pivot_wider() |>
  transmute(river_name = river_name |> factor(levels = c("North Yuba River", "Middle Yuba River", "South Yuba River", "Yuba River", "Lower Yuba River")),
            method,
            actual_length_mi = length, # units are miles
            actual_gravel_ac = gravel * 1000 / 43560)

spawning_matrix_summary |> knitr::kable()

```

## Predictions

Look at predictions

```{r}
lower_yuba_river <-
  habistat::cv_mainstems |>
  filter(river_cvpia == "Yuba River" & str_detect(habitat, "rearing"))

spawning_predictions_comid <- 
  habistat::wua_predicted |>
  inner_join(selected_streams |> select(comid, current_spawning_reach)) |>
  filter(habitat == "spawning") |>
  filter(model_name == "SD") |>
  filter(comid %in% filter(spawning_context, spawning_filter_all)$comid) |>
  inner_join(habistat::flowline_attr |> select(comid, gnis_name)) |>
  filter(gnis_name %in% c("Yuba River", "Middle Yuba River", "North Yuba River", "South Yuba River")) |>
  mutate(river_name = if_else(comid %in% lower_yuba_river$comid, "Lower Yuba River", gnis_name) |>
         factor(levels = c("North Yuba River", "Middle Yuba River", "South Yuba River", "Yuba River", "Lower Yuba River"))) 

total_reach_lengths <-
  habistat::flowline_attr |>
  mutate(river_name = if_else(comid %in% lower_yuba_river$comid, "Lower Yuba River", gnis_name) |>
         factor(levels = c("North Yuba River", "Middle Yuba River", "South Yuba River", "Yuba River", "Lower Yuba River"))) |>
  group_by(river_name) |> 
  summarize(river_length_ft = sum(reach_length_ft), .groups="drop") |>
  drop_na()

spawning_predictions <- 
  spawning_predictions_comid |>
  group_by(river_name, flow_cfs) |>
  mutate(total_wua_ft2 = wua_per_lf_pred * reach_length_ft) |>
  summarize(total_length_ft = sum(reach_length_ft),
            total_wua_ft2 = sum(total_wua_ft2),
            wua_ft2_per_lf = total_wua_ft2 / total_length_ft) |>
  left_join(total_reach_lengths, by=join_by(river_name)) |>
  mutate(wua_ft2_per_tot_lf = total_wua_ft2 / river_length_ft)

plt_wua <- spawning_predictions |>
  ggplot() + geom_line(aes(x = flow_cfs, y = wua_ft2_per_lf, color = river_name)) + 
  scale_x_log10(breaks = scales::breaks_log(10)) + annotation_logticks(sides="b") +
  ylab("Spawning Habitat\n(ft2) per habitat LF") + guides(color = "none") +
  scale_y_continuous(sec.axis = sec_axis(name = "Redds \nper 1000 linear ft", transform = ~./94*1000)) +
  scale_color_brewer(name = "Spawning/Reach", aesthetics = c("color", "fill"), palette="Paired")

plt_wua2 <- spawning_predictions |>
  ggplot() + geom_line(aes(x = flow_cfs, y = wua_ft2_per_tot_lf, color = river_name)) + 
  scale_x_log10(breaks = scales::breaks_log(10)) + annotation_logticks(sides="b") +
  ylab("Spawning Habitat\n(ft2) per total LF") + guides(color = "none") +
  scale_y_continuous(sec.axis = sec_axis(name = "Redds \nper river mile", transform = ~./94*5280)) +
  scale_color_brewer(name = "Spawning/Reach", aesthetics = c("color", "fill"), palette="Paired")

plt_tot <- spawning_predictions |>
  ggplot() + geom_area(aes(x = flow_cfs, y = total_wua_ft2 / 43560, fill = river_name), color = "white") + 
  scale_x_log10(breaks = scales::breaks_log(10)) + annotation_logticks(sides="b") +
  ylab("Spawning Habitat\n(total acres)") + guides(color = "none", fill = "none") + 
  scale_y_continuous(sec.axis = sec_axis(name = "Redds\n(total)", transform = ~./94*43560)) + 
  scale_color_brewer(name = "Spawning/Reach", aesthetics = c("color", "fill"), palette="Paired")

plt_map <- selected_streams |>
  filter(comid %in% spawning_predictions_comid$comid) |>
  ggplot() + geom_sf(data=selected_streams, color="lightgray") + 
  geom_sf(aes(color = river_name), linewidth=1) +
  theme(panel.grid = element_blank(), axis.text = element_blank()) + 
  scale_color_brewer(name = "Spawning/Reach", aesthetics = c("color", "fill"), palette="Paired")

((plt_wua / plt_wua2 / plt_tot) + 
    plot_layout(axes = "collect") & 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks.y = element_line())) | 
  (guide_area() / plt_map) +
  plot_layout(guides = "collect") 
```

```{r}
carrying_capacity <- 
  spawning_predictions |>
  mutate(ccap_redds = total_wua_ft2 / 94) # 01_YSF-Habitat-Matrices-Report estimate: 94ft^2/redd

carrying_capacity_summary <- 
  carrying_capacity |> 
  mutate(ccap_redds = total_wua_ft2 / 94) |>
  group_by(river_name) |> 
  summarize(at_flow = flow_cfs[which.max(ccap_redds)],
            ccap_redds = max(ccap_redds), .groups="drop") 

carrying_capacity_summary |> knitr::kable()
```

Length of spawning habitat

```{r}
stream_lengths_spawning_habitat <- 
  selected_streams |>
  filter(spawning_filter_all) |>
  group_by(river_name) |>
  summarize(length_ft = sum(st_length(geometry)) |> units::set_units("ft") |> units::drop_units(),
            length_mi = length_ft / 5280) |>
  st_drop_geometry()

stream_lengths_spawning_habitat |> knitr::kable()
```

Amount of spawning gravel and number of redds

Using the max value across all flows

```{r}
stream_spawning_areas <- 
  spawning_predictions |>
  group_by(river_name) |>
  summarize(across(c(total_length_ft, total_wua_ft2, wua_ft2_per_lf), max), .groups="drop") |>
  mutate(spawn_area_ft2_1000 = total_wua_ft2 / 1000,
         spawn_area_acres = total_wua_ft2 / 43560,
         n_redds = total_wua_ft2 / 94) # 94 ft2 per redd

stream_spawning_areas |> knitr::kable()
```

## Compare against yuba relicensing estimates

Compare

```{r}
spawning_matrix_summary |>
  inner_join(stream_lengths_spawning_habitat |> 
               group_by(river_name) |> 
               summarize(predicted_length_mi = sum(length_mi), .groups="drop"), 
             by=join_by(river_name)) |>
  inner_join(stream_spawning_areas |> 
               group_by(river_name) |> 
               summarize(predicted_spawning_ac = sum(spawn_area_acres), .groups="drop"), 
             by=join_by(river_name)) |>
  mutate(ratio_length = (predicted_length_mi / actual_length_mi) |> num(digits = 2),
         ratio_gravel = (predicted_spawning_ac / actual_gravel_ac) |> num(digits = 2)) |>
  select(river_name, method, 
         actual_length_mi, predicted_length_mi, ratio_length,
         actual_gravel_ac, predicted_spawning_ac, ratio_gravel) |>
  knitr::kable()
```

```{r}
spawning_predictions |>
  ggplot() + 
  geom_line(aes(x = flow_cfs, y = (total_wua_ft2 / 43560), color = river_name, linetype = "habistat Predictions")) + 
  #geom_line(aes(x = flow_cfs, y = (total_wua_ft2 / 43560) - actual_gravel_ac, color = river_name)) + 
  geom_hline(data = spawning_matrix_summary, aes(yintercept = actual_gravel_ac, color = river_name, linetype = paste("YSF Habitat Matrices", method))) +
  scale_x_log10(breaks = scales::breaks_log(10)) + annotation_logticks(sides="b") +
  ylab("Spawning Habitat\n(total acres)") + guides(color = "none", fill = "none") + 
  scale_y_continuous(sec.axis = sec_axis(name = "Redds\n(total)", transform = ~./94*43560)) + 
  scale_color_brewer(name = "Spawning/Reach", aesthetics = c("color", "fill"), palette="Paired") +
  facet_wrap(~river_name, ncol=2)  + theme(legend.position = "top", panel.grid.minor = element_blank()) +
  ggtitle("Carrying Capacity Comparison") + 
  guides(linetype = guide_legend(nrow=2, byrow=TRUE, title = ""))
```

## Redds Survey

RM 0 is at the mouth of the river, and RM 24.3 is at the base of Englebright Dam.

```{r}
lyr_redd_survey <- read_csv(here::here("data-raw", "source", "spawning_data", "spring-run_and_fall-run_redds_for_Rene_data.csv")) |>
  janitor::clean_names() |>
  select(-redds_total) |>
  pivot_longer(starts_with("redds")) |>
  separate(name, sep = "_", into = c("var", "year")) |>
  separate(rm, sep = " - ", into = c("rm_start", "rm_end")) |>
  mutate(across(starts_with("rm_"), as.numeric)) |>
  mutate(time_range = factor(time_range, levels = c("On or Prior to Oct 15", "On or After Oct 16"))) |>
  select(-var) %>%
  bind_rows(. |>
    group_by(rm_start, rm_end, time_range) |>
    summarize(value = max(value)) |>
    mutate(year = "max"))

lyr_redd_survey |>
  group_by(year, time_range) |>
  summarize(total_redds = sum(value)) |>
  mutate(spawning_area_ac = total_redds * 94 / 43560) |>
  select(time_range, year, total_redds, spawning_area_ac) |>
  arrange(time_range, year)
  
plt_redd_survey <-
  lyr_redd_survey |>
  ggplot() + 
  geom_step(aes(x = rm_start, y = value, color = time_range)) +
  xlab("River Mile") + ylab("Numer of Redds") + 
  facet_wrap(~year, ncol=1)  +
  ggtitle("2009-2010 Lower Yuba River Redd Survey") + 
  guides(color = guide_legend(nrow=1, byrow=TRUE, title = "")) +
  scale_y_continuous(sec.axis = sec_axis(name = "Redd Area (ac) per RM", transform = ~.*94/43560)) +
  theme(panel.grid.minor = element_blank(), legend.position="top") +
  scale_x_continuous(limits = c(0, 24.3))

print(plt_redd_survey)
```

```{r}
lyr_rm <- selected_streams |>
  filter(river_name == "Lower Yuba River") |>
  inner_join(habistat::flowline_attr |> select(comid, hydro_seq, reach_length_ft), by=join_by(comid)) |>
  arrange(hydro_seq) |>
  mutate(reach_length_mi = reach_length_ft / 5280,
         river_seq = row_number(),
         cum_length_end = cumsum(reach_length_mi),
         cum_length_start = cum_length_end - reach_length_mi,
         rm_start = scales::rescale(cum_length_start, 
                                    from = c(min(cum_length_start), max(cum_length_end)), 
                                    to = c(0, 24.3)),
         rm_end = scales::rescale(cum_length_end, 
                                    from = c(min(cum_length_start), max(cum_length_end)), 
                                    to = c(0, 24.3))) 

plt_spawning_filters <-
  lyr_rm |>
  pivot_longer(cols = starts_with("spawning_")) |>
  ggplot() + 
  facet_wrap(~name, ncol=1) +
  geom_rect(aes(xmin = rm_start, xmax = rm_end, ymin = -1, ymax = 1, fill = value)) +
  theme(panel.grid.major.y = element_blank(), 
        panel.grid.minor.y = element_blank(), 
        axis.text.y = element_blank(),
        legend.position = "top") +
  xlab("River Mile") + ggtitle("Habistat Geographic Context")

print(plt_spawning_filters)
```

```{r}
plt_habistat_by_rm <-
  lyr_rm |> 
  st_drop_geometry() |> 
  select(comid, rm_start, rm_end, spawning_filter_all) |>
  expand_grid(flow_idx = c(316, 1000, 3162, 10000)) |>
  left_join(spawning_predictions_comid, by=join_by(comid, flow_idx), relationship="one-to-one") |>
  #group_by(comid, rm_start, rm_end) |> 
  #summarize(wua_per_lf_pred = max(wua_per_lf_pred)) |>
  ggplot() +
  geom_step(aes(x = rm_start, y = if_else(spawning_filter_all, wua_per_lf_pred * 5280 / 43560, 0))) + 
  facet_wrap(~flow_idx, ncol = 1) + 
  ylab("Potential Spawning Area (ac) per RM") + xlab("River Mile") + 
  ggtitle("Habistat Area Predictions")

print(plt_habistat_by_rm)
```

```{r fig.height=10, fig.width=6.5}
(plt_spawning_filters / plt_habistat_by_rm / plt_redd_survey) + plot_layout(axes = "collect")
```





