---
title: "Spawning Gravel Carrying Capacity"
author: "[Skyler Lewis](mailto:slewis@flowwest.com)"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
    toc_depth: 3
    number_sections: false
    math_method:
      engine: webtex
      url: https://latex.codecogs.com/svg.image?
---

```{r setup}
library(tidyverse)
library(sf)
library(habistat)
library(patchwork)
theme_set(theme_minimal())
```

CVPIA stream "Yuba River" is just the Lower Yuba River

```{r}
habistat::cv_mainstems |> filter(str_detect(river_cvpia, "Yuba")) |> plot()
```

Take all Yuba River mainstems instead

```{r}
selected_streams <- 
habistat::flowline_geom_proj |> 
  inner_join(habistat::flowline_attr |> select(comid, watershed_level_3, gnis_name, river_cvpia)) |>
  filter(watershed_level_3 == "Yuba River") |> 
  filter(gnis_name %in% c("Yuba River", "Middle Yuba River", "North Yuba River", "South Yuba River")) 

selected_streams |> plot()
```

Filter for spawning context (likely spawning gravels)

```{r}
spawning_context <- readRDS(here::here("data-raw", "results", "spawning_context.Rds")) |>
  glimpse()

selected_streams |>
  filter(comid %in% filter(spawning_context, spawning_filter_all)$comid) |>
  plot()

```

**These are weird at the margins because of the flow range filter. Need to use unfiltered version in `wua_predicted` -- filter only for map display not for calculations.

```{r}
spawning_predictions_comid <- 
habistat::wua_predicted |>
  filter(comid %in% selected_streams$comid) |>
  filter(habitat == "spawning") |>
  filter(comid %in% filter(spawning_context, spawning_filter_all)$comid) |>
  inner_join(habistat::flowline_attr |> select(comid, gnis_name)) |>
  filter(gnis_name %in% c("Yuba River", "Middle Yuba River", "North Yuba River", "South Yuba River"))

spawning_predictions <- 
  spawning_predictions_comid |>
  group_by(gnis_name, flow_cfs) |>
  mutate(total_wua_ft2 = wua_per_lf_pred * reach_length_ft) |>
  summarize(total_wua_ft2 = sum(total_wua_ft2)) |>
  inner_join(habistat::flowline_attr |> 
               group_by(gnis_name) |> 
               summarize(reach_length_ft = sum(reach_length_ft))) |>
  mutate(wua_ft2_per_lf = total_wua_ft2 / reach_length_ft)

spawning_predictions <- 
  spawning_predictions_comid |>
  group_by(gnis_name, flow_cfs) |>
  mutate(total_wua_ft2 = wua_per_lf_pred * reach_length_ft) |>
  summarize(total_length_ft = sum(reach_length_ft),
            total_wua_ft2 = sum(total_wua_ft2),
            wua_ft2_per_lf = total_wua_ft2 / total_length_ft) 

plt_wua <- spawning_predictions |>
  ggplot() + geom_line(aes(x = flow_cfs, y = wua_ft2_per_lf, color = gnis_name)) + 
  scale_x_log10()
plt_tot <- spawning_predictions |>
  ggplot() + geom_line(aes(x = flow_cfs, y = total_wua_ft2 / 43560, color = gnis_name)) + 
  scale_x_log10()
(plt_wua / plt_tot) + plot_layout(axes = "collect", guides = "collect")
```


```{r}
carrying_capacity <- 
spawning_predictions |>
  mutate(ccap_redds = total_wua_ft2 / 94) 
# 01_YSF-Habitat-Matrices-Report estimate: 94ft^2/redd

plt_ccap <- carrying_capacity |>
  ggplot() + #geom_line(aes(x = flow_cfs, y = ccap_redds, color = gnis_name)) + 
  scale_x_log10() + geom_hline(data = carrying_capacity |> 
                                 group_by(gnis_name) |> 
                                 summarize(ccap_redds = max(ccap_redds)), 
                               aes(yintercept = ccap_redds, color = gnis_name)) 

(plt_wua / plt_tot / plt_ccap) + plot_layout(axes = "collect", guides = "collect")
```

Length of spawning habitat

Calculated just using the elevation/gradient filter

```{r}
stream_lengths <- 
selected_streams |>
  group_by(gnis_name) |>
  summarize(length_ft = sum(st_length(geometry)) |> units::set_units("ft") |> units::drop_units(),
            length_mi = length_ft / 5280) |>
  st_drop_geometry()

stream_lengths |> knitr::kable()
```

Calculated using the sediment transport filters too

```{r}
stream_lengths_filtered <- 
  selected_streams |>
  filter(comid %in% filter(spawning_context, spawning_filter_all)$comid) |>
  group_by(gnis_name) |>
  summarize(length_ft = sum(st_length(geometry)) |> units::set_units("ft") |> units::drop_units(),
            length_mi = length_ft / 5280) |>
  st_drop_geometry()

stream_lengths_filtered |> knitr::kable()
```

Amount of spawning gravel and number of redds

Using the max value across all flows

```{r}
stream_spawning_areas <- 
  spawning_predictions |>
  group_by(gnis_name) |>
  summarize(across(c(total_length_ft, total_wua_ft2, wua_ft2_per_lf), max)) |>
  mutate(spawn_area_ft2_1000 = total_wua_ft2 / 1000,
         spawn_area_acres = total_wua_ft2 / 43560,
         n_redds = total_wua_ft2 / 94) # 94 ft2 per redd

stream_spawning_areas |> knitr::kable()
```

Compare against estimates in the Yuba relicensing study

```{r}
read_csv(here::here("data-raw", "source", "spawning_data", "01_YSF-Habitat-Matrices-Report_table8_data.csv"), col_types = "cnnnnnnnnnnn") |>
  mutate(gnis_name = case_when(str_detect(reach, "North Yuba") ~ "North Yuba River",
                               str_detect(reach, "Middle Yuba") ~ "Middle Yuba River",
                               str_detect(reach, "South Yuba") ~ "South Yuba River",
                               TRUE ~ "Yuba River")) |>
  group_by(gnis_name, year) |>
  summarize(across(c(starts_with("length_"), starts_with("gravel_"), starts_with("redds_")), function(x) sum(x, na.rm=T))) |>
  select(gnis_name, year, ends_with("_UT")) |>
  group_by(gnis_name) |>
  summarize(across(c(starts_with("length_"), starts_with("gravel_"), starts_with("redds_")), mean)) |>
  transmute(gnis_name,
            actual_length_mi = length_UT, # units are miles
            actual_gravel_ac = gravel_UT * 1000 / 43560) |> # convert 1000 ft2 to ft2 to acres) |>
  inner_join(stream_lengths_filtered |> transmute(gnis_name, predicted_length_mi = length_mi)) |>
  inner_join(stream_spawning_areas |> select(gnis_name, predicted_spawning_ac = spawn_area_acres)) |>
  mutate(ratio_length = (actual_length_mi / predicted_length_mi) |> num(digits = 2),
         ratio_gravel = (actual_gravel_ac / predicted_spawning_ac) |> num(digits = 2)) |>
  knitr::kable()
```
