---
title: "Predictor Data Preparation and Consolidation"
author: "[Skyler Lewis](mailto:slewis@flowwest.com)"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
    toc_depth: 3
    number_sections: false
    math_method:
      engine: webtex
      url: https://latex.codecogs.com/svg.image?
---

```{r setup}
library(tidyverse)
library(sf)
library(habistat)
theme_set(theme_minimal())
```

## Basic Flow Crosswalk via Drainage Area and Precipitation

See watershed_delineation.R for dependencies

```{r multipliers}
make_flow_xw <- function(group_var) {
  habistat::flowline_attr |>
  select({{group_var}}, 
         comid = comid,
         da_reach = da_area_sq_km_tot,
         pc_reach = da_ppt_mean_mm) |>
  drop_na({{group_var}}) |>
  group_by({{group_var}}) |>
  mutate(outlet_comid = comid[which.max(da_reach)]) |>
  mutate(da_outlet = da_reach[which(comid == outlet_comid)],
         pc_outlet = pc_reach[which(comid == outlet_comid)]) |>
  mutate(multiplier = (da_reach / da_outlet) * (pc_reach / pc_outlet)) |>
  arrange({{group_var}}, -multiplier)
}

cv_watersheds_flow_xw <- make_flow_xw(watershed_level_3)
cv_watersheds_flow_xw |> usethis::use_data(overwrite=T)

cv_mainstems_flow_xw <- make_flow_xw(river_cvpia)
cv_mainstems_flow_xw |> usethis::use_data(overwrite=T)

habistat::flowline_geom_proj |>
  inner_join(cv_watersheds_flow_xw, by=join_by(comid)) |>
  filter(comid %in% habistat::wua_predicted$comid) |>
  ggplot() + 
  geom_sf(aes(color = multiplier)) + 
  geom_sf(data = habistat::flowline_geom_proj |> 
            filter(comid %in% cv_watersheds_flow_xw$outlet_comid) |> 
            st_line_sample(sample=1), 
          aes(color = 1)) + 
  scale_color_viridis_c(direction = -1) +
  ggtitle("Flow Multipliers by Watershed")

habistat::flowline_geom_proj |>
  inner_join(cv_mainstems_flow_xw, by=join_by(comid)) |>
  filter(comid %in% habistat::wua_predicted$comid) |>
  ggplot() + 
  geom_sf(aes(color = multiplier)) + 
  geom_sf(data = habistat::flowline_geom_proj |> 
            filter(comid %in% cv_mainstems_flow_xw$outlet_comid) |> 
            st_line_sample(sample=1), 
          aes(color = 1)) + 
  scale_color_viridis_c(direction = -1) + 
  ggtitle("Flow Multipliers by Mainstem")
```

### Applying the crosswalk to aggregate model results

```{r}
wua_predicted_cv_watersheds <- 
  habistat::wua_predicted |>
  inner_join(cv_watersheds_flow_xw, 
             by=join_by(watershed_level_3, comid)) |>
  expand_grid(scaled = c(FALSE, TRUE)) |>
  mutate(flow_cfs = if_else(scaled, flow_cfs * multiplier, flow_cfs)) |>
  arrange(habitat, model_name, comid, flow_cfs) |>
  group_by(habitat, model_name, comid) |>
  mutate(wua_per_lf_pred = if_else(scaled, NA, wua_per_lf_pred)) |>
  mutate(wua_per_lf_pred = zoo::na.approx(wua_per_lf_pred, x = flow_cfs, na.rm=F, rule=2)) |>
  filter(scaled) |>
  select(-scaled) |>
  ungroup() |>
  group_by(habitat, model_name, watershed_level_3, flow_idx) |>
  summarize(wua_per_lf_pred = sum(wua_per_lf_pred * reach_length_ft) / sum(reach_length_ft),
            wua_acres_pred = sum(wua_per_lf_pred * reach_length_ft) / 43560, .groups="drop") |>
  inner_join(habistat::wua_predicted |> 
               group_by(flow_idx) |> 
               summarize(flow_cfs = first(flow_cfs), .groups="drop"),
             by = join_by(flow_idx))

wua_predicted_cv_watersheds |> usethis::use_data(overwrite=T)

wua_predicted_cv_mainstems <- 
  habistat::wua_predicted |> 
  inner_join(cv_mainstems_flow_xw, 
             by=join_by(river_cvpia, comid)) |>
  expand_grid(scaled = c(FALSE, TRUE)) |>
  mutate(flow_cfs = if_else(scaled, flow_cfs * multiplier, flow_cfs)) |>
  arrange(habitat, model_name, comid, flow_cfs) |>
  group_by(habitat, model_name, comid) |>
  mutate(wua_per_lf_pred = if_else(scaled, NA, wua_per_lf_pred)) |>
  mutate(wua_per_lf_pred = zoo::na.approx(wua_per_lf_pred, x = flow_cfs, na.rm=F, rule=2)) |>
  filter(scaled) |>
  select(-scaled) |>
  ungroup() |>
  group_by(habitat, model_name, river_cvpia, flow_idx) |>
  summarize(wua_per_lf_pred = sum(wua_per_lf_pred * reach_length_ft) / sum(reach_length_ft),
            wua_acres_pred = sum(wua_per_lf_pred * reach_length_ft) / 43560, .groups="drop") |>
  inner_join(habistat::wua_predicted |> 
               group_by(flow_idx) |> 
               summarize(flow_cfs = first(flow_cfs), .groups="drop"),
             by = join_by(flow_idx))

wua_predicted_cv_mainstems |> usethis::use_data(overwrite=T)
```

Plotting the output

```{r pred-watersheds-new, fig.width=8.5, fig.height=8.5, dpi=300}
wua_predicted_cv_watersheds |>
  ggplot(aes(x = flow_cfs)) + 
  facet_wrap(~watershed_level_3, scales="free") + 
  geom_line(aes(y = wua_acres_pred, color = paste(habitat, model_name))) +
  scale_x_log10() +
  theme(legend.position = "top", panel.grid.minor = element_blank()) + 
  scale_color_brewer(palette = "Paired") +
  ylab("Predicted Total Habitat (acres)") + xlab("Flow at Outlet (cfs)")
```

Versus the "naive" method

```{r pred-watersheds-old, fig.width=8.5, fig.height=8.5, dpi=300}
wua_predicted |>
  filter(!is.na(watershed_level_3)) |>
  group_by(habitat, model_name, watershed_level_3, flow_idx, flow_cfs) |>
  summarize(wua_per_lf_pred = sum(wua_per_lf_pred * reach_length_ft) / sum(reach_length_ft),
            wua_acres_pred = sum(wua_per_lf_pred * reach_length_ft) / 43560, .groups="drop") |>
  ggplot(aes(x = flow_cfs)) + 
  facet_wrap(~watershed_level_3, scales="free") + 
  geom_line(aes(y = wua_acres_pred, color = paste(habitat, model_name))) +
  scale_x_log10() +
  theme(legend.position = "top", panel.grid.minor = element_blank()) + 
  scale_color_brewer(palette = "Paired") +
  ylab("Predicted Total Habitat (acres)") + xlab("Flow (cfs)")
```

## Applying this method to the duration analyis

* Associate each streamgage with a COMID, drainage area, and DA precipitation
* When doing COMID scale duration analysis, use these values to scale the x axis of the inundation curve directly
* When doing watershed scale duration analysis
  - pick a streamgage
  - for each selected comid, 
    - create a flow-to-suitable-area curve along downscaled flows using complete-approx method (above)
    - make a flow-scaled version of the gage's inundation curve
    - apply the flow-scaled inundation curve to the flow-scaled habitat curve
  - consolidate and sum the result curves

### Streamgage drainage area attributes

```{r}
streamgage_da_attr <-
  habistat::cv_mainstems |>
  nest(stem_ln_geom = c(comid, habitat, geometry), .by = c(river_group, river_cvpia)) |>
  inner_join(habistat::streamgage_attr, by=join_by(river_group, river_cvpia)) |>
  inner_join(x=habistat::streamgage_geom |> nest(gage_pt_geom = c(geometry)), 
             y=_, 
             by=join_by(station_id)) |>
  mutate(gage_comid = map2_int(gage_pt_geom, stem_ln_geom,
                               function(gage_pt, stem_ln) stem_ln$comid[[st_nearest_feature(gage_pt, stem_ln)]])) |>
  select(station_id, gage_comid) |>
  inner_join(habistat::flowline_attr |>
              select(gage_comid = comid,
                     da_gage = da_area_sq_km,
                     pc_gage = da_ppt_mean_mm), 
             by=join_by(gage_comid)) |>
  glimpse()

streamgage_da_attr |> usethis::use_data(overwrite = T)
```
### Example application to COMID-level flow-to-suitable-area curve

```{r}
selected_comid <- 8020536  
selected_gage <- "dcv"
selected_habitat <- "rearing"
selected_model <- "SD"
selected_run <- "fall"
selected_wy_type <- "Dry"

fsa <-
  habistat::wua_predicted |>
  filter((model_name == selected_model) & 
           (habitat == selected_habitat) &
           (comid == selected_comid)) |>
  glimpse()

selected_comid_attr <- 
  habistat::flowline_attr |>
  filter(comid == selected_comid) |>
  as.list()

drc <-
  habistat::streamgage_duration_rating_curves |>
  filter((station_id == selected_gage) & 
           (habitat == selected_habitat) & 
           (run == selected_run) & 
           (wy_group == selected_wy_type)) |>
  unnest(data) 

drc_scaled <- drc |>
  inner_join(streamgage_da_attr, by=join_by(station_id)) |>
  mutate(model_q = model_q * (selected_comid_attr$da_area_sq_km_tot / da_gage) * (selected_comid_attr$da_ppt_mean_mm / pc_gage)) |>
  glimpse()

habistat::duration_apply_dhsi_to_fsa_curve(
        fsa = fsa,
        fsa_q = flow_cfs,
        fsa_wua = wua_per_lf_pred,
        drc = drc_scaled,
        drc_q = model_q,
        drc_dhsi = durhsi_rearing_vf) |>
        glimpse()
```

### Example application to watershed-level flow-to-suitable-area curve

```{r}
# CONTINUE HERE
```

```{r}
knitr::knit_exit()
```

## Streamgage Crosswalks

See duration_predictions.Rmd for dependencies

```{r}
# requires first running the first part of duration_predictions.Rmd
# in order to download the streamgage time series CSVs to temporary local folder
cdec_filenames <- readRDS(here::here("data-raw", "temp", "cdec_filenames.Rds"))

read_cdec_daily_ts <- function(filename) {
  read_csv(filename, col_types = cols()) |>
    janitor::clean_names() |>
    mutate(value = as.numeric(value)) |>
    filter((value >= 0) & !is.nan(value)) |>
    mutate(obs_date = date(obs_date)) |>
    group_by(obs_date) |>
    summarize(flow_cfs = mean(value, na.rm=T))
}

fit_streamgage_power_series <- function(from_station, to_station) {
  
  ts1 <- 
    filter(cdec_filenames, station_id == from_station)$filename[[1]] |>
    read_cdec_daily_ts() |>
    rename(from_flow = flow_cfs)
  
  ts2 <- 
    filter(cdec_filenames, station_id == to_station)$filename[[1]] |>
    read_cdec_daily_ts() |>
    rename(to_flow = flow_cfs)
  
  ts <- 
    inner_join(ts1, ts2, by=join_by(obs_date)) |>
    filter(from_flow > 0 & to_flow > 0)
  
  if (nrow(ts) >= 365) {
      
    model_fit <- 
      lm(data = ts,
         formula = log(to_flow) ~ log(from_flow))
    
    model_fit_summary <- summary(model_fit)
    
    return(list(
      coeff = exp(model_fit_summary$coefficients[[1, 1]]),
      power = model_fit_summary$coefficients[[2, 1]],
      rsq = model_fit_summary$r.squared,
      n_days = nrow(ts)))
  
  } else {
      
    return(list(
      coeff = NA,
      power = NA,
      rsq = NA,
      n_days = nrow(ts)))

  }
}

if(!file.exists(here::here("data-raw", "results", "streamgage_xw_models.Rds"))) {
  
  streamgage_xw_models <-
    expand_grid(from_station = habistat::streamgage_attr$station_id,
                to_station = habistat::streamgage_attr$station_id) |>
    mutate(result = map2(from_station, to_station, fit_streamgage_power_series)) |>
    unnest_wider(result)
  
  streamgage_xw_models |> saveRDS(here::here("data-raw", "results", "streamgage_xw_models.Rds")) 
  
} else {
  
  streamgage_xw_models <- readRDS(here::here("data-raw", "results", "streamgage_xw_models.Rds"))
  
}
```






